### 2018-09-29 Manuscript/Manuscript.tex
initialize repo with an empty manuscript and some code from before

\documentclass[11pt,letterpaper]{article}

%\usepackage{fontspec}
%\usepackage[utf8]{inputenc}
\usepackage{textcomp,marvosym}
\usepackage{amsmath,amssymb}
\usepackage[normalem]{ulem}
\usepackage[left]{lineno}
\usepackage{changepage}
\usepackage{rotating}
\usepackage{color}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage[hidelinks]{hyperref}
\urlstyle{same}
\usepackage{threeparttable}
\doublespacing

\raggedright
\textwidth = 6.5 in
\textheight = 8.25 in
\oddsidemargin = 0.0 in
\evensidemargin = 0.0 in
\topmargin = 0.0 in
\headheight = 0.0 in
\headsep = 0.5 in
\parskip = 0.1 in
\parindent = 0.2in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}

% Remove brackets from numbering in List of References
%\makeatletter
%\renewcommand{\@biblabel}[1]{\quad#1.}
%\makeatother

% Self defined commands
\newcommand{\degC}{$^{\circ}$C\xspace}
\newcommand{\dC}{$\delta^{13}$C\xspace}
\newcommand{\dO}{$\delta^{18}$O\xspace}
\newcommand{\SrSr}{$^{87}$Sr/$^{86}$Sr\xspace}
\newcommand{\permil}{\textperthousand\xspace}
\newcommand{\dsil}{$d$\xspace}
\newcommand{\UPb}{$^{206}$Pb/$^{238}$U\xspace}
%

\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\lhead{Park et al., in preparation for XXX}
\rhead{\thepage}

\begin{document}

\begin{flushleft}
{\Large \textbf{XXX}}
\\
Yuem Park\textsuperscript{1},
Nicholas L. Swanson-Hysell\textsuperscript{1}
\\
\bigskip
\textsuperscript{1} Department of Earth and Planetary Science, University of California, Berkeley, CA, USA
\\
\textsuperscript{2} XXX
\bigskip

\end{flushleft}

\noindent\textit{This article is in preparation for submission to XXX.}

\linenumbers

\section*{ABSTRACT \label{sec:ABSTRACT}}

XXX

\section*{INTRODUCTION \label{sec:INTRODUCTION}}

XXX

\section*{ACKNOWLEDGEMENTS \label{sec:ACKNOWLEDGEMENTS}}
\footnotesize

XXX

\singlespacing

\bibliographystyle{gsabull}
\bibliography{References}

\end{document}

### 2018-10-01 Manuscript/Manuscript.tex
minor updates

@@
 \fancyhf{}
-\lhead{Park et al., in preparation for XXX}
+\lhead{Park et al., in preparation for AGU book entitled \textit{Environmental Change and Large Igneous Provinces}}
 \rhead{\thepage}
@@
 Yuem Park\textsuperscript{1},
-Nicholas L. Swanson-Hysell\textsuperscript{1}
+Nicholas L. Swanson-Hysell\textsuperscript{1},
+Francis A. Macdonald\textsuperscript{2}
+Lorraine E. Lisiecki\textsuperscript{2}
 \\
@@
 \\
-\textsuperscript{2} XXX
+\textsuperscript{2} Department of Earth Science, University of California, Santa Barbara, CA, USA
 \bigskip
@@
 
-\noindent\textit{This article is in preparation for submission to XXX.}
+\noindent\textit{This article is in preparation for submission to \textit{Environmental Change and Large Igneous Provinces}.}
 

### 2018-10-14 Manuscript/Manuscript.tex
Additions to discussion and methods

@@
 \begin{flushleft}
-{\Large \textbf{XXX}}
-\\
+{\Large \textbf{Evaluating the relationship between large igneous province area and Earth's long-term climate state}}
+
 Yuem Park\textsuperscript{1},
 Nicholas L. Swanson-Hysell\textsuperscript{1},
-Francis A. Macdonald\textsuperscript{2}
+Francis A. Macdonald\textsuperscript{2},
 Lorraine E. Lisiecki\textsuperscript{2}
-\\
+
 \bigskip
 \textsuperscript{1} Department of Earth and Planetary Science, University of California, Berkeley, CA, USA
-\\
 \textsuperscript{2} Department of Earth Science, University of California, Santa Barbara, CA, USA
@@
 
-XXX
+The proposed environmental impacts of large igneous province (LIP) eruptions are myriad (as summarized in \citealp{Ernst2017a}).
+One aspect of LIP emplacement that has been hypothesized to relate to long-term global climate is the effect such emplacement could have on increasing global weatherability and driving cooling.
+Global weatherability is the sum of factors aside from climate itself that contribute to overall global weathering.
+On a planet with high weatherability, the CO$_2$ input from volcanism can be removed via silicate weathering at a lower atmospheric CO$_2$ concentration than on a less weatherable planet.
+The relatively high concentrations of Ca and Mg (that ultimately sequester carbon through precipitation as carbonate) and the high weathering rates of mafic lithologies make it so that basaltic regions consume more CO$_2$ than regions where bedrock composition is closer to bulk continental crust \citep{Dessert2003a}.
+Data from basaltic watersheds show that chemical weathering rates are greatest in regions with high runoff, and as a result CO$_2$ consumption in basaltic regions is most pronounced in the tropical rain belt \citep{Dessert2003a, Hartmann2014a}.
+
+Given these factors, the emplacement of LIPs in the tropics has been associated with specific episodes of climatic cooling on Earth.
+In the Neoproterozoic, the emplacement of the Franklin LIP in the tropics ca.
+720~Ma, in concert with elevated runoff rates associated with continental break-up, has been implicated as a major contributor to the cooling that initiated the Sturtian snowball Earth \citep{Donnadieu2004b, Cox2016a}.
+In the Cenozoic, the movement of the Deccan LIP into the tropical rain belt has been implicated in drawing down CO$_2$ levels in the lead-up to Eocene glaciation of Antarctica \citep{Kent2008a}.
+
+This chapter seeks to address two interconnected questions: 1)  how uniquely large are the peaks in low-latitude LIP area that have been proposed to be associated with climatic cooling?; and 2) how strong is the overall relationship between tropical LIP area and glaciation?
+
+\section*{METHODS}
+
+Outlines of LIPs through the Phanerozoic (Fig.
+\ref{fig:LIP_map}) were taken from the compilation of \citet{Ernst2017a}, and emplacement ages were taken from the literature (Fig.
+\ref{fig:LIP_map}; Table \ref{tab:LIPs}).
+The compilation of \citet{Ernst2017a} seeks to reconstruct the original areal extent of LIPs with the caveat that there can be significant uncertainty with doing so, particularly for older more deeply eroded LIPs.
+These polygons encapsulate all of the rocks associated with a given LIP including dikes and sills (Fig.
+\ref{fig:LIP_map}).
+For some LIPs, this approach may lead to an overestimate of original areal extent given that subsurface intrusions could extend over a broader area than the surface volcanics.
+However, such intrusions likely provide the best estimate available of original extent for ancient LIPs.
+The extent of presently exposed volcanics associated with the LIPs were taken from a number of resources including the PLATES compilation (Table \ref{tab:LIPs}; \citealp{Coffin2006a}) and more recent compilation efforts associated with the Large Igneous Provinces Commission.
+
+After LIPs are emplaced they will progressively erode.
+In order to account for this decrease in area through with time, \citet{Godderis2017a} took the approach of fitting estimates of the original surface extent and estimates of the current extent for 5 LIPs with an exponential decay function.
+They used the resulting exponential decay constants to develop a first-order model to estimate changing LIP area through time.
+We extend this approach to 19 basaltic LIPs for which we have estimates of the original areal extent of the province and the current areal extent of rocks associated with it (Fig.
+\ref{fig:LIP_preservation}).
+While there are significant uncertainties associated with the area estimates, this compilation suggests that an exponential decay function is an appropriate approximate model for the progressive reduction in LIP area.
+The best-fit exponential function results in a LIP area half-life of 29~Myr.
+However, since we explicitly account for LIP burial separately in our models (see below), we prefer to exclude the 6 of 19 LIPs which have been partially or completely buried in our estimation of a representative LIP area half-life, which yields a slightly higher best-fit half-life of 36~Myr (Fig.
+\ref{fig:LIP_preservation}).
+There is variability in the individual half-lives for LIPs such that they range from shorter half-lives of $\sim$20~Myr up to $\sim$120~Myr for the Kalkarindji and Deccan Traps.
+In the analysis of LIP area through time, we implement decay scenarios informed by this analysis.
+Scenario `decay 1' uses the best-fit half-life of 36 Myr, while `decay 2' uses the slower decay with a half-life of $\sim$120~Myr.
+Given that the post-emplacement weathering and erosional history of each LIP should be dependent on the tectonic and climatic setting that each LIP experiences during and after emplacement, this approach is simplistic but provides a framework for analysis.
+
+Of the tectonic factors that could alter exposed LIP area, the most consequential is near-immediate burial of LIP volcanics that are co-located with a rift basin by sediments.
+There are numerous examples in the record where there is partial or complete burial of a LIP associated with rifting and thermal subsidence (Table \ref{tab:LIPs}).
+For example, the Afar LIP is both associated with the Ethiopian Traps which form plateau flood basalts, and successful rifting in the region of the Red Sea that has resulted in burial.
+To account for the rapid decrease in exposed surface area that would result from burial by sediments in a rift basin, we impose two different burial scenarios for LIPs that are co-located with rifting.
+`Burial 1' imposes instantaneous burial of 50$\%$ of the LIP area while `burial 2' imposes instantaneous burial of the entire LIP as an end-member scenario.
+The LIP area analysis uses all of the available combinations of these distinct decay and burial scenarios.
+Our LIP reconstruction differs from that of \citet{Johansson2018a} in several respects.
+In contrast to the decay and decay+burial scenarios implemented on estimates of original LIP extent in this study, \citet{Johansson2018a} uses a static extent for each LIP throughout the reconstruction with some of the polygons corresponding to the present-day surface extent and some representing the original extent that includes currently buried portions of the LIP (e.g.
+the North American Mid-continent Rift).
+Furthermore, the LIP reconstruction used in this study includes pre-Phanerozoic LIPs.
+However, given the imposed exponential decay since emplacement, the inclusion of these LIPs do not significantly impact the calculated LIP areas throughout the Phanerozoic.
+
+The original extent LIP polygons were assigned a plate ID corresponding to a tectonic unit on Earth using the polygons of \citet{Torsvik2016a} for the Phanerozoic.
+The LIP polygons and tectonic units were reconstructed from 520~Ma to the present utilizing the paleogeographic model of \citet{Torsvik2016a} in the spin axis reference frame (anchor plate ID of 1).
+This paleogeographic model was updated to include a revision to Ordovician Laurentia \citep{Swanson-Hysell2017a} and the Paleozoic of Asia \citep{Domeier2018a}.
+Reconstructions and area calculations within latitude bands utilized the pyGPlates function library and custom Python scripts documented within a Jupyter notebook that reproduces the analysis and development of the associated visualizations.
+The total LIP area and the LIP area reconstructed within the tropical rain belt (15\textdegree~S to 15\textdegree~N) were calculated for the various decay and burial scenarios at a resolution of 5~Myr (Fig.
+\ref{fig:LIP_area}).
+
+To determine whether there is a relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic Eon (Fig.
+\ref{fig:LIP_area}).
+The goal in doing so is to evaluate the hypothesis that there is a correlation between LIP area in the tropics and Earth's long-term climate state.
+The latitudinal extent of land ice is reflective of Earth's overall climate and allows for the delineation of glacial and non-glacial climate states.
+However, the land ice record is an imperfect tracker, as it is insensitive to changes in temperatures during non-glacial intervals and is influenced by additional factors such as the physical geography of the continents during glacial intervals.
+To seek to deal with these issues, we take two approaches for comparison.
+The first to to look at the correlation between LIP area and the extent of ice away from the pole (Fig.
+\ref{fig:LIP_correlation}).
+The second is to consider the overall correlation between intervals of high LIP area in the tropics (defined as LIP area in the tropics \textgreater30\% of the maximum) and intervals of glacial climate (defined as ice extent \textgreater30\% of the maximum).
+This latter approach places less emphasis on the specific magnitudes of the compiled ice extent and LIP area records.
+
+We note that another approach that has been implemented in the literature (e.g.
+\citealp{Johansson2018a}) is to compare LIP area records to proxy compilations of pCO$_2$.
+Such proxy compilations are problematic as they are poorly calibrated in deep time, prone to alteration, and reliant on a myriad of assumptions.
+We thus prefer to use the latitudinal extent of land ice to reflect Earth's overall climate state.
+
+\section*{RESULTS}
+
+OLD TEXT FROM THE SCIENCE MANUSCRIPT THAT NEEDS TO BE UPDATED IN ACCORDANCE WITH NEW RESULTS
+In the “decay” scenario, we observe two primary peaks and one minor peak in calculated LIP area within the tropics (Fig.
+S4B).
+The first primary peak starting ca.
+377 Ma is associated with the emplacement of the Kola-Dnieper and the second primary peak starting ca.
+251 Ma is associated with the emplacement of the Central Atlantic magmatic province (CAMP).
+A Cenozoic peak is associated with both the ca.
+30 Ma emplacement of the Afro-Arabian (Afar) LIP as well as the effect of the earlier drift of the Deccan LIP into the tropics.
+In the “burial” scenario, the peak associated with the CAMP is transient, but the other peaks remain (Fig.
+S4).
+
+\section*{DISCUSSION}
+
+The results from this analysis indicate that when the entire LIP database is considered, there is not a significant relationship between total LIP area or LIP area in the tropics and the extent of continental ice sheets.
+While this result need not imply that there is no increase in global weatherability from the emplacement of LIPs, it does suggest that LIP area is not the fundamental control on Earth's long-term climate state.
+
+In the original models that proposed the ``Fire and Ice'' hypothesis as an explanation for the onset of Sturtian glaciation, chemical weathering was modeled as an Arhennius relationship where it was solely a function of temperature and run-off \citep{Donnadieu2004a}.
+However, such an approach neglects the effects that can result through soil shielding and regolith development in low-relief regions.
+Recent progress on understanding the relationships between landscapes, topography and chemical weathering reveals that these effects are quite important (e.g.
+\cite{Maher2014a}).
+As a result, more recent modeling of chemical weathering that has incorporated such processes highlights the importance of high-relief regions relative to low-relief ones \citep{Godderis2017b}.
+Without active uplift, soil shielding from regolith development on low-relief LIPs could significantly decrease the local weatherability of a LIP and mute its impact on global weatherability.
+In contrast, processes that lead to continued exhumation of mafic lithologies and the creation of steep topography, particularly in tropical regions, may exert a strong control on global weatherability and long-term climate.
+This interpretation underlies the hypothesis that arc-continent collisons in the tropics during the Ordovician \citep{Swanson-Hysell2017a} and the Cenozoic \citep{Jagoutz2016a} played a significant role in transitions into glacial climate states at those times.
+
+A complication with the interpretation of soil shielding and limited weathering of LIPs is the rapid area decay rate inferred from the comparison of current LIP extent to estimated original extent.
+A couple considerations are relevant with respect to this analysis: 1) the current extent of LIP exposure is reduced in part by volcanics being covered by unconsolidated sediments (i.e.
+regolith development itself) in a number of the provinces; 2) the initial LIP extents are typically poorly constrained and are likely overestimates which could be resulting in inflated interpreted decay rates.
+Future efforts to develop better-constrained estimates of original LIP extent will improve analyses such as that in this contribution.
+
+We have focused this analysis on the Phanerozoic record given that well-constrained paleogeographic models are available for the past $\sim$500 Myr.
+Furthermore, the approach of seeking to evaluate correlation between LIP area and glaciation is complicated for Neoproterozoic pan-glacial events wherein ice-albedo runaway leads to persistent global glaciation absent of continued forcing through normal carbon cycle processes until the build-up of sufficient pCO$_2$ to drive deglaciation.
+Nevertheless, the  hypothesis of tropical LIP area associated with the ca.
+720 Ma Franklin LIP increasing global weatherability and contributing to the onset of the Sturtian Snowball Earth glaciation is a major motivating driver behind conducting this analysis.
+How does the tropical LIP area associated with the Franklin LIP compare to that observed in the Phanerozoic?
+DISCUSS FRANKLIN AND UMKONDO AREAS AND COMPARE TO THE PHANEROZOIC RECORD.
+
+\section*{CONCLUSIONS}
 
 \section*{ACKNOWLEDGEMENTS \label{sec:ACKNOWLEDGEMENTS}}
+
+Richard Ernst provided GIS compilations of LIP extent that were essential to the analysis.
+
+\section*{FIGURES}
+
+\begin{figure}[h!]
+\begin{center}
+	\includegraphics[width=\textwidth]{Manuscript/Figures/LIP_Map.pdf}
+	\caption{\textbf{Map of current extent of lithologies associated with LIPs that erupted between 520 Ma and the present (shown in orange) as well as the estimates of the initial LIP area used in the area analysis (shown in blue and slightly modified from \citealp{Ernst2017a}).}}
+	\label{fig:LIP_map}
+\end{center}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
+	\includegraphics[width=0.65\textwidth]{Manuscript/Figures/LIP_Preservation.pdf}
+	\caption{\textbf{LIP preservation through time.
+The ratio of estimates of the present-day area to that of the original area are shown for 17 basaltic LIPs.
+The best-fit exponential function to these data is shown which has a half-life of 36 Myr.}}
+	\label{fig:LIP_preservation}
+\end{center}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
+	\includegraphics[width=0.9\textwidth]{Manuscript/Figures/LIP_Areas.pdf}
+	\caption{\textbf{A) paleolatitude of emplacement and the initial area estimate of each LIP in the compilation.
+    B) total LIP area through time for the different post-emplacement models.
+    C) tropical LIP area through time for the different post-emplacement models.
+    D) heat map for one of the models showing the latitudinal distribution of LIP area.
+    E) latitudinal extent away from the poles of land ice through time.}}
+	\label{fig:LIP_area}
+\end{center}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
+	\includegraphics[width=0.85\textwidth]{Manuscript/Figures/overlap_correlation_cropped.pdf}
+	\caption{\textbf{The overlap and correlation values for various LIP post-emplacement scenarios are shown with circles (d1 : LIP decay with half-life of 36 Myr; d2: LIP decay with half-life of 120 Myr; b1: 50$\%$ burial of LIPs associated with rifting; b2: 100$\%$ burial of LIPs associated with rifting).
+These values are compared to histograms that are range of values that arise when comparing the LIP area record to glacial intervals that have been shifted randomly in time 100,000 times.
+The only scenario for which there is a positive correlation coefficient related to ice extent and LIP area is that of a relatively slow decay rate (d2) and complete burial of LIPs associated with rifting (b2).
+This positive correlation is weak (\textit{r}=0.10) and is not significant as randomly timed glacial intervals correlate better than the record 28$\%$ of the time.
+With an associated p-value of 0.28, the null hypothesis that glacial intervals not related to LIP area in the tropics cannot be rejected.}}
+	\label{fig:LIP_correlation}
+\end{center}
+\end{figure}
+
+\clearpage
+\newpage
 \footnotesize
-
-XXX
 

### 2018-10-15 Manuscript/Manuscript.tex
Updates from ShareLaTeX

@@
 \textsuperscript{1} Department of Earth and Planetary Science, University of California, Berkeley, CA, USA
+
 \textsuperscript{2} Department of Earth Science, University of California, Santa Barbara, CA, USA
@@
 
-Given these factors, the emplacement of LIPs in the tropics has been associated with specific episodes of climatic cooling on Earth.
-In the Neoproterozoic, the emplacement of the Franklin LIP in the tropics ca.
+Given these factors, the emplacement of LIPs in the tropics has been hypothesized to be associated with specific episodes of climatic cooling on Earth.
+In the Neoproterozoic Era, the emplacement of the Franklin LIP in the tropics ca.
 720~Ma, in concert with elevated runoff rates associated with continental break-up, has been implicated as a major contributor to the cooling that initiated the Sturtian snowball Earth \citep{Donnadieu2004b, Cox2016a}.
@@
 
-This chapter seeks to address two interconnected questions: 1)  how uniquely large are the peaks in low-latitude LIP area that have been proposed to be associated with climatic cooling?; and 2) how strong is the overall relationship between tropical LIP area and glaciation?
+This chapter seeks to address two interconnected questions: 1) how uniquely large are the peaks in low-latitude LIP area that have been proposed to be associated with climatic cooling?
+and 2) how strong is the overall relationship between tropical LIP area and glaciation?
 
@@
 For some LIPs, this approach may lead to an overestimate of original areal extent given that subsurface intrusions could extend over a broader area than the surface volcanics.
-However, such intrusions likely provide the best estimate available of original extent for ancient LIPs.
+The polygons also assume complete areal coverage by LIPs as they connect wide-spread remnants--also creating the potential for them being over-estimates.
+However, this approach likely provide the best estimates available of original extent for ancient LIPs.
 The extent of presently exposed volcanics associated with the LIPs were taken from a number of resources including the PLATES compilation (Table \ref{tab:LIPs}; \citealp{Coffin2006a}) and more recent compilation efforts associated with the Large Igneous Provinces Commission.
@@
 Given that the post-emplacement weathering and erosional history of each LIP should be dependent on the tectonic and climatic setting that each LIP experiences during and after emplacement, this approach is simplistic but provides a framework for analysis.
+The LIP reconstruction used in this study includes pre-Phanerozoic LIPs.
+However, given the imposed exponential decay since emplacement, the inclusion of these LIPs do not significantly impact the calculated LIP areas throughout the Phanerozoic.
 
-Of the tectonic factors that could alter exposed LIP area, the most consequential is near-immediate burial of LIP volcanics that are co-located with a rift basin by sediments.
+Of the tectonic factors that could alter exposed LIP area, the most consequential is near-immediate burial by sediments of LIP volcanics that are co-located with a rift basin.
 There are numerous examples in the record where there is partial or complete burial of a LIP associated with rifting and thermal subsidence (Table \ref{tab:LIPs}).
-For example, the Afar LIP is both associated with the Ethiopian Traps which form plateau flood basalts, and successful rifting in the region of the Red Sea that has resulted in burial.
+For example, the Afar LIP is both associated with the Ethiopian Traps which form plateau flood basalts, and successful rifting in the region of the Red Sea that has resulted in burial (Fig \ref{fig:LIP_map}).
 To account for the rapid decrease in exposed surface area that would result from burial by sediments in a rift basin, we impose two different burial scenarios for LIPs that are co-located with rifting.
@@
 the North American Mid-continent Rift).
-Furthermore, the LIP reconstruction used in this study includes pre-Phanerozoic LIPs.
-However, given the imposed exponential decay since emplacement, the inclusion of these LIPs do not significantly impact the calculated LIP areas throughout the Phanerozoic.
 
@@
 
-To determine whether there is a relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic Eon (Fig.
+To evaluate the relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic Eon (modified from \citealp{Royer2004a}; Fig.
 \ref{fig:LIP_area}).
@@
 The latitudinal extent of land ice is reflective of Earth's overall climate and allows for the delineation of glacial and non-glacial climate states.
-However, the land ice record is an imperfect tracker, as it is insensitive to changes in temperatures during non-glacial intervals and is influenced by additional factors such as the physical geography of the continents during glacial intervals.
-To seek to deal with these issues, we take two approaches for comparison.
+The land ice record is an imperfect tracke, as it is insensitive to changes in temperatures during non-glacial intervals and is influenced by additional factors such as the physical geography of the continents during glacial intervals.
+Nevertheless, it forms a physical record of Earth's climate state through time (Fig.
+\ref{fig:LIP_area}).
+We take two approaches for comparison between the LIP area reconstruction and the record of ice extent (Fig.
+\ref{fig:LIP_area}).
 The first to to look at the correlation between LIP area and the extent of ice away from the pole (Fig.
 \ref{fig:LIP_correlation}).
-The second is to consider the overall correlation between intervals of high LIP area in the tropics (defined as LIP area in the tropics \textgreater30\% of the maximum) and intervals of glacial climate (defined as ice extent \textgreater30\% of the maximum).
-This latter approach places less emphasis on the specific magnitudes of the compiled ice extent and LIP area records.
+The second is to consider the overlap between intervals of high LIP area in the tropics (defined as LIP area in the tropics \textgreater30\% of the maximum in a given post-emplacement model) and intervals of glacial climate (defined as ice extent \textgreater10\textdegree $\;$from the poles).
+This overlap approach places less emphasis on the specific magnitudes of the compiled ice extent and LIP area records.
 
-We note that another approach that has been implemented in the literature (e.g.
-\citealp{Johansson2018a}) is to compare LIP area records to proxy compilations of pCO$_2$.
-Such proxy compilations are problematic as they are poorly calibrated in deep time, prone to alteration, and reliant on a myriad of assumptions.
+Another approach would be to compare LIP area records to proxy compilations of pCO$_2$ (as done by \citealp{Johansson2018a}).
+However, such proxy compilations are problematic as they are poorly calibrated in deep time, prone to alteration, and reliant on a myriad of assumptions.
 We thus prefer to use the latitudinal extent of land ice to reflect Earth's overall climate state.
@@
 
+I THINK HERE IN THE RESULTS IS WHERE THE MONTE CARLO APPROACH CAN BE ELUCIDATED.
+WE CAN SUCCINCTLY MAKE THE POINT THAT DIRECT CROSS-CORRELATION ISN'T APPROPRIATE WITHOUT SUCH NULL HYPOTHESIS TESTING.
+WE CAN PRESENT THAT THERE IS A WEEK POSITIVE CORRELATION WITH d2+b2 AND THEN DESCRIBE HOW THE RANDOM GENERATE GLACIAL INTERVAL APPROACH SHOWS THAT IT IS NOT SIGNIFICANT.
+THIS TEXT FROM THE FIG 4 CAPTION IS A START:
+\textit{The only scenario for which there is a positive correlation coefficient related to ice extent and LIP area is that of a relatively slow decay rate (d2) and complete burial of LIPs associated with rifting (b2).
+This positive correlation is weak (\textit{r}=0.10) and is not significant as randomly timed glacial intervals correlate better than the record 28$\%$ of the time.
+With an associated p-value of 0.28, the null hypothesis that glacial intervals not related to LIP area in the tropics cannot be rejected.}
+
 \section*{DISCUSSION}
 
-The results from this analysis indicate that when the entire LIP database is considered, there is not a significant relationship between total LIP area or LIP area in the tropics and the extent of continental ice sheets.
+The results from this analysis indicate that when the entire LIP database is considered, there is not a significant relationship between total LIP area or LIP area in the tropics and the extent of continental ice sheets (Fig.
+\ref{fig:LIP_correlation}).
 While this result need not imply that there is no increase in global weatherability from the emplacement of LIPs, it does suggest that LIP area is not the fundamental control on Earth's long-term climate state.
@@
 Recent progress on understanding the relationships between landscapes, topography and chemical weathering reveals that these effects are quite important (e.g.
-\cite{Maher2014a}).
+\citealp{Maher2014a}).
 As a result, more recent modeling of chemical weathering that has incorporated such processes highlights the importance of high-relief regions relative to low-relief ones \citep{Godderis2017b}.

### 2018-10-15 Manuscript/Manuscript.tex

@@
 
+\section*{TABLES}
+
+\begin{table}[h!]
+\scriptsize
+\caption{place holder}
+\noindent\makebox[\textwidth]{
+\begin{tabular}{c c c c c}
+\hline
+\textbf{Fraction} & \textbf{$^{206}$Pb/$^{238}$U} & & \textbf{$^{207}$Pb/$^{235}$U} & \\
+& \textbf{Age [Ma]} & $\pm$2$\sigma$ & \textbf{Age [Ma]} & $\pm$2$\sigma$ \\
+\hline
+z1 & 823.2201669 & 1.067507326 & 823.4322744 & 2.609783978 \\
+z2 & 822.401975 & 0.714471249 & 819.0720505 & 1.260737 \\
+z4 & 836.2094848& 1.478314118 & 828.1995093 & 6.104877613 \\
+z5 & 834.109136 & 1.288954576 & 825.9918302 & 5.720788585 \\
+z16 & 831.1612867 & 0.741238812 & 832.226412 & 1.64433912 \\
+z17 & 854.9767495 & 0.72532716 & 855.8348689 & 1.16873047 \\
+z18 & 812.4046191 & 0.665110984 & 818.3985038 & 1.082055988 \\
+z19 & 832.8963617 & 0.694437022 & 833.1583004 & 1.060745556 \\
+\hline
+\end{tabular}
+}
+\begin{tablenotes}[para,flushleft]
+place holder.
+\end{tablenotes}
+\label{tab:LIPs}
+\end{table}
+
 \section*{FIGURES}
@@
 	\includegraphics[width=\textwidth]{Manuscript/Figures/LIP_Map.pdf}
-	\caption{\textbf{Map of current extent of lithologies associated with LIPs that erupted between 520 Ma and the present (shown in orange) as well as the estimates of the initial LIP area used in the area analysis (shown in blue and slightly modified from \citealp{Ernst2017a}).}}
+	\caption{Map of current extent of volcanic lithologies associated with LIPs that erupted between 520~Ma and the present as well as the estimates of the initial LIP area used in the area analysis (modified slightly from \citealp{Ernst2017a}).}
 	\label{fig:LIP_map}
@@
 	\includegraphics[width=0.65\textwidth]{Manuscript/Figures/LIP_Preservation.pdf}
-	\caption{\textbf{LIP preservation through time.
-The ratio of estimates of the present-day area to that of the original area are shown for 17 basaltic LIPs.
-The best-fit exponential function to these data is shown which has a half-life of 36 Myr.}}
+	\caption{LIP preservation through time.
+The ratio of estimates of the present-day area to that of the original area are shown for 19 basaltic LIPs.
+An exponential fit is made to the 13 basaltic LIPs that have not been buried after emplacement, which yields a half-life of $\sim$36~Myr.}
 	\label{fig:LIP_preservation}
@@
 	\includegraphics[width=0.9\textwidth]{Manuscript/Figures/LIP_Areas.pdf}
-	\caption{\textbf{A) paleolatitude of emplacement and the initial area estimate of each LIP in the compilation.
-    B) total LIP area through time for the different post-emplacement models.
-    C) tropical LIP area through time for the different post-emplacement models.
-    D) heat map for one of the models showing the latitudinal distribution of LIP area.
-    E) latitudinal extent away from the poles of land ice through time.}}
+	\caption{\textbf{A)} Paleolatitude of emplacement (centroid) and the initial area estimate of each LIP in the compilation.
+\textbf{B)} Total LIP area through time for the different post-emplacement models.
+\textbf{C)} Tropical LIP area through time for the different post-emplacement models.
+The arrows indicate tropical LIP area for the ca.
+720~Ma Franklin LIP and the ca.
+1109~Ma Umkondo LIP.
+\textbf{D)} Heat map for the preferred post-emplacement model showing the latitudinal distribution of LIP area.
+\textbf{E)} Latitudinal extent of land ice away from the poles through time.}
 	\label{fig:LIP_area}
@@
 	\includegraphics[width=0.85\textwidth]{Manuscript/Figures/overlap_correlation_cropped.pdf}
-	\caption{\textbf{The overlap and correlation values for various LIP post-emplacement scenarios are shown with circles (d1 : LIP decay with half-life of 36 Myr; d2: LIP decay with half-life of 120 Myr; b1: 50$\%$ burial of LIPs associated with rifting; b2: 100$\%$ burial of LIPs associated with rifting).
-These values are compared to histograms that are range of values that arise when comparing the LIP area record to glacial intervals that have been shifted randomly in time 100,000 times.
+	\caption{The overlap and correlation values for various LIP post-emplacement scenarios are shown with circles (d1 : LIP decay with half-life of 36~Myr; d2: LIP decay with half-life of~120 Myr; b1: 50\% burial of LIPs associated with rifting; b2: 100\% burial of LIPs associated with rifting).
+These values are compared to histograms that show the range of values that arise when comparing the LIP area record to glacial intervals that have been shifted randomly in time 100,000 times.
 The only scenario for which there is a positive correlation coefficient related to ice extent and LIP area is that of a relatively slow decay rate (d2) and complete burial of LIPs associated with rifting (b2).
-This positive correlation is weak (\textit{r}=0.10) and is not significant as randomly timed glacial intervals correlate better than the record 28$\%$ of the time.
-With an associated p-value of 0.28, the null hypothesis that glacial intervals not related to LIP area in the tropics cannot be rejected.}}
+This positive correlation is weak (\textit{r}=0.10) and is not significant as randomly timed glacial intervals correlate better than the record 28\% of the time.
+With an associated p-value of 0.28, the null hypothesis that glacial intervals not related to LIP area in the tropics cannot be rejected.}
 	\label{fig:LIP_correlation}

### 2018-10-15 Manuscript/Manuscript.tex

@@
 The total LIP area and the LIP area reconstructed within the tropical rain belt (15\textdegree~S to 15\textdegree~N) were calculated for the various decay and burial scenarios at a resolution of 5~Myr (Fig.
-\ref{fig:LIP_area}).
+\ref{fig:LIP_area}B and C).
 
 To evaluate the relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic Eon (modified from \citealp{Royer2004a}; Fig.
-\ref{fig:LIP_area}).
+\ref{fig:LIP_area}E).
 The goal in doing so is to evaluate the hypothesis that there is a correlation between LIP area in the tropics and Earth's long-term climate state.
 The latitudinal extent of land ice is reflective of Earth's overall climate and allows for the delineation of glacial and non-glacial climate states.
-The land ice record is an imperfect tracke, as it is insensitive to changes in temperatures during non-glacial intervals and is influenced by additional factors such as the physical geography of the continents during glacial intervals.
-Nevertheless, it forms a physical record of Earth's climate state through time (Fig.
-\ref{fig:LIP_area}).
-We take two approaches for comparison between the LIP area reconstruction and the record of ice extent (Fig.
-\ref{fig:LIP_area}).
-The first to to look at the correlation between LIP area and the extent of ice away from the pole (Fig.
+The land ice record is an imperfect tracker, as it is insensitive to changes in temperatures during non-glacial intervals and is influenced by additional factors such as the physical geography of the continents during glacial intervals.
+Nevertheless, it forms a physical record of Earth's climate state through time.
+We take two approaches for comparison between the LIP area reconstruction and the record of ice extent.
+The first is to look at the correlation between LIP area and the extent of ice away from the pole (Fig.
 \ref{fig:LIP_correlation}).
-The second is to consider the overlap between intervals of high LIP area in the tropics (defined as LIP area in the tropics \textgreater30\% of the maximum in a given post-emplacement model) and intervals of glacial climate (defined as ice extent \textgreater10\textdegree $\;$from the poles).
+The second is to consider the overlap between intervals of high LIP area in the tropics (defined as LIP area in the tropics \textgreater30\% of the maximum in a given post-emplacement model) and intervals of glacial climate (defined as ice extent \textgreater10\textdegree\xspace from the poles).
 This overlap approach places less emphasis on the specific magnitudes of the compiled ice extent and LIP area records.
@@
 
-OLD TEXT FROM THE SCIENCE MANUSCRIPT THAT NEEDS TO BE UPDATED IN ACCORDANCE WITH NEW RESULTS
-In the “decay” scenario, we observe two primary peaks and one minor peak in calculated LIP area within the tropics (Fig.
-S4B).
-The first primary peak starting ca.
-377 Ma is associated with the emplacement of the Kola-Dnieper and the second primary peak starting ca.
-251 Ma is associated with the emplacement of the Central Atlantic magmatic province (CAMP).
-A Cenozoic peak is associated with both the ca.
-30 Ma emplacement of the Afro-Arabian (Afar) LIP as well as the effect of the earlier drift of the Deccan LIP into the tropics.
-In the “burial” scenario, the peak associated with the CAMP is transient, but the other peaks remain (Fig.
-S4).
+In the `decay 1' scenario, we observe three primary peaks and one minor peak in the calculated LIP area within the tropics (Fig.
+\ref{fig:LIP_area}C).
+The first primary peak ca.
+510~Ma is associated with the emplacement of the Kalkarindji LIP, the second primary peak ca.
+380~Ma is associated with the emplacement of the Kola-Dnieper LIP, the third primary peak ca.
+200~Ma is associated with the emplacement of the CAMP, and the minor Cenozoic peak is associated with both the ca.
+30~Ma emplacement of the Afar LIP as well as the earlier drift of the ca.
+66~Ma Deccan LIP into the tropics (Fig.
+\ref{fig:LIP_area}A).
+When we account for burial (`decay 1 + burial 1' and `decay 1 + burial 2' scenarios), only the latter two of these four peaks are affected -- the ca.
+200~Ma peak is attenuated/removed due to the partial/complete removal of the CAMP, and the Cenozoic peak is attenuated due to the partial/complete removal of the Afar LIP.
+However, after accounting for burial, a minor area of LIPs remain in the tropics from ca.
+130~Ma onwards, due to the EQUAMP, Caribbean-Colombian, and Deccan LIPs.
+Using the longer decay half-life of 120~Myr (the `decay 2 + burial 2' scenario) increases the area of LIPs in the tropics at any given time step, and has the effect of extending the duration of each peak.
+
+The only scenario which results in a positive Pearson correlation coefficient (0.10) between LIP area in the tropics (Fig.
+\ref{fig:LIP_area}C) and the ice extent record (Fig.
+\ref{fig:LIP_area}E) is the scenario with the slow decay rate and complete burial of LIPs associated with rifting (the `decay 2 + burial 2' scenario; Fig.
+\ref{fig:LIP_correlation}).
+All other scenarios yield a weak negative correlation coefficient.
+Qualitatively, the weak positive correlation of the `decay 2 + burial 2' scenario relative to the other scenarios can be primarily attributed to the complete removal of the CAMP, which was emplaced during an extended interval of ice-free conditions, as well as the effect of the longer decay half-life extending the duration of the earlier two peaks, such that they overlap more with the Late Ordivician and Permo-Carboniferous glacial intervals.
+
+To assess the statistical significance of the raw Pearson correlation coefficients,
 
@@
 WE CAN SUCCINCTLY MAKE THE POINT THAT DIRECT CROSS-CORRELATION ISN'T APPROPRIATE WITHOUT SUCH NULL HYPOTHESIS TESTING.
-WE CAN PRESENT THAT THERE IS A WEEK POSITIVE CORRELATION WITH d2+b2 AND THEN DESCRIBE HOW THE RANDOM GENERATE GLACIAL INTERVAL APPROACH SHOWS THAT IT IS NOT SIGNIFICANT.
+WE CAN PRESENT THAT THERE IS A WEAK POSITIVE CORRELATION WITH d2+b2 AND THEN DESCRIBE HOW THE RANDOM GENERATE GLACIAL INTERVAL APPROACH SHOWS THAT IT IS NOT SIGNIFICANT.
 THIS TEXT FROM THE FIG 4 CAPTION IS A START:
@@
 	\caption{\textbf{A)} Paleolatitude of emplacement (centroid) and the initial area estimate of each LIP in the compilation.
+Car.-Col.
+= Caribbean-Colombian.
 \textbf{B)} Total LIP area through time for the different post-emplacement models.
+Only the `no decay' model excludes pre-Phanerozoic LIPs.
 \textbf{C)} Tropical LIP area through time for the different post-emplacement models.

### 2018-10-15 Manuscript/Manuscript.tex
Updates from ShareLaTeX

@@
 
-To assess the statistical significance of the raw Pearson correlation coefficients,
-
-I THINK HERE IN THE RESULTS IS WHERE THE MONTE CARLO APPROACH CAN BE ELUCIDATED.
-WE CAN SUCCINCTLY MAKE THE POINT THAT DIRECT CROSS-CORRELATION ISN'T APPROPRIATE WITHOUT SUCH NULL HYPOTHESIS TESTING.
-WE CAN PRESENT THAT THERE IS A WEAK POSITIVE CORRELATION WITH d2+b2 AND THEN DESCRIBE HOW THE RANDOM GENERATE GLACIAL INTERVAL APPROACH SHOWS THAT IT IS NOT SIGNIFICANT.
-THIS TEXT FROM THE FIG 4 CAPTION IS A START:
-\textit{The only scenario for which there is a positive correlation coefficient related to ice extent and LIP area is that of a relatively slow decay rate (d2) and complete burial of LIPs associated with rifting (b2).
-This positive correlation is weak (\textit{r}=0.10) and is not significant as randomly timed glacial intervals correlate better than the record 28$\%$ of the time.
-With an associated p-value of 0.28, the null hypothesis that glacial intervals not related to LIP area in the tropics cannot be rejected.}
+To assess the statistical significance of the raw Pearson correlation coefficients, we simulated the four glacial episodes (Fig.
+\ref{fig:LIP_area}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and \% overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these simulations.
+For all scenarios except for the `decay 2 + burial 2' scenario, more than half of the randomly timed glacial intervals correlate better with the LIP area in the tropics (Fig.
+\ref{fig:LIP_correlation}).
+For the `decay 2 + burial 2' scenario, 28\% of the randomly timed glacial intervals correlate better with the LIP area in the tropics.
+With an associated p-value of 0.28, the null hypothesis that glacial intervals do not correlate to LIP area in the tropics cannot be rejected.
 
@@
 	\caption{The overlap and correlation values for various LIP post-emplacement scenarios are shown with circles (d1 : LIP decay with half-life of 36~Myr; d2: LIP decay with half-life of~120 Myr; b1: 50\% burial of LIPs associated with rifting; b2: 100\% burial of LIPs associated with rifting).
-These values are compared to histograms that show the range of values that arise when comparing the LIP area record to glacial intervals that have been shifted randomly in time 100,000 times.
-The only scenario for which there is a positive correlation coefficient related to ice extent and LIP area is that of a relatively slow decay rate (d2) and complete burial of LIPs associated with rifting (b2).
-This positive correlation is weak (\textit{r}=0.10) and is not significant as randomly timed glacial intervals correlate better than the record 28\% of the time.
-With an associated p-value of 0.28, the null hypothesis that glacial intervals not related to LIP area in the tropics cannot be rejected.}
+These values are compared to histograms that show the range of values that arise when comparing the LIP area record to glacial intervals that have been shifted randomly in time 100,000 times.}
 	\label{fig:LIP_correlation}

### 2018-10-16 Manuscript/Figures/LIP_Table.tex
add LIP table tex (but it needs to be fixed)

\begin{tabular}{lrlrrlrrl}
\toprule
                name &    age &                     age ref. &  original area Mkm2 &  present area Mkm2 &                                 present area ref. &  present/original &  half-life &   buried \\
\midrule
      Columbia River &   16.0 &   Kasbohm and Schoene (2018) &                0.68 &               0.38 &                           Buchan and Ernst (2004) &              0.56 &       19.1 &       no \\
                Afar &   30.0 &  Courtillot and Renne (2003) &                2.05 &               0.63 &                              Coffin et al. (2006) &              0.31 &       17.8 &  partial \\
                NAIP &   62.0 &         Larsen et al. (2016) &                1.07 &               0.29 &     Buchan and Ernst (2004), Coffin et al. (2006) &              0.27 &       32.8 &  partial \\
              Deccan &   66.0 &               Schoene (2015) &                0.83 &               0.56 &                              Coffin et al. (2006) &              0.68 &      118.6 &       no \\
          Seychelles &   66.0 &               Schoene (2015) &                0.46 &               0.00 &                              Coffin et al. (2006) &              0.00 &        0.0 &      yes \\
          Madagascar &   90.0 &    Cucciniello et al. (2010) &                0.63 &               0.03 &                              Coffin et al. (2006) &              0.05 &       20.8 &       no \\
 Caribbean-Colombian &   94.0 &         Loewen et al. (2013) &                0.71 &               0.13 &                              Coffin et al. (2006) &              0.18 &       38.0 &       no \\
               HALIP &   95.0 &      Kingsbury et al. (2018) &                3.60 &               0.15 &                      Moosdorf and Hartmann (2015) &              0.04 &       20.5 &       no \\
              EQUAMP &  131.0 &       Hollanda et al. (2016) &                0.66 &               0.01 &                            Hollanda et al. (2016) &              0.01 &       19.7 &       no \\
               Comei &  132.0 &            Zhu et al. (2009) &                0.11 &                NaN &                                               n/a &               NaN &        NaN &       no \\
             Bunbury &  132.0 &            Zhu et al. (2009) &                0.03 &               0.00 &                              Thorne et al. (2014) &              0.05 &       30.5 &       no \\
     Parana-Etendeka &  135.0 &      Florisbal et al. (2014) &                3.12 &               0.40 &  Coffin et al. (2016), Teixeira and Albrez (2017) &              0.13 &       45.9 &  partial \\
                Trap &  140.0 &          Ernst et al. (2001) &                0.03 &               0.00 &                           Ernst and Buchan (2001) &              0.00 &        0.0 &       no \\
 NW Australia Margin &  160.0 &            COME BACK TO THIS &                0.62 &               0.00 &                              Coffin et al. (2006) &              0.00 &        0.0 &      yes \\
               Karoo &  183.0 &        Burgess et al. (2015) &                3.21 &               0.15 &                               de Kock compilation &              0.05 &       42.3 &       no \\
              Ferrar &  183.0 &        Burgess et al. (2015) &                0.18 &                NaN &                                               n/a &               NaN &        NaN &       no \\
                CAMP &  201.0 &      Blackburn et al. (2013) &               11.46 &               0.23 &                   Marzoli and Parisio compilation &              0.02 &       35.6 &  partial \\
             Siberia &  252.0 &   Burgess and Bowring (2015) &                3.46 &               0.47 &                              Coffin et al. (2006) &              0.14 &       88.8 &       no \\
            Emeishan &  259.0 &           Zhou et al. (2002) &                0.71 &               0.06 &                              Coffin et al. (2006) &              0.09 &       74.6 &       no \\
    Panjal-Qiangtang &  283.0 &           Zhai et al. (2013) &                0.11 &                NaN &                                               n/a &               NaN &        NaN &       no \\
               Tarim &  290.0 &             Xu et al. (2014) &                0.35 &                NaN &                                               n/a &               NaN &        NaN &       no \\
            Magdalen &  360.0 &         Murphy et al. (1999) &                0.42 &                NaN &                                               n/a &               NaN &        NaN &       no \\
              Vilyui &  374.0 &          Renne et al. (2010) &                1.14 &                NaN &                                               n/a &               NaN &        NaN &       no \\
        Kola-Dnieper &  380.0 &    Arzamastsev and Wu (2014) &                5.90 &                NaN &                                               n/a &               NaN &        NaN &       no \\
            Suordakh &  450.0 &       Khudoley et al. (2013) &                0.02 &                NaN &                                               n/a &               NaN &        NaN &       no \\
         Kalkarindji &  511.0 &        Jourdan et al. (2014) &                3.54 &               0.17 &                              Thorne et al. (2014) &              0.05 &      118.2 &       no \\
            Franklin &  720.0 &       Denyszyn et al. (2009) &                2.62 &               0.04 &                           Buchan and Ernst (2004) &              0.02 &      127.6 &       no \\
\bottomrule
\end{tabular}

### 2018-10-16 Manuscript/Manuscript.tex
Updates from ShareLaTeX

@@
 \usepackage[left]{lineno}
+\usepackage{booktabs}
 \usepackage{changepage}
@@
 One aspect of LIP emplacement that has been hypothesized to relate to long-term global climate is the effect such emplacement could have on increasing global weatherability and driving cooling.
-Global weatherability is the sum of factors aside from climate itself that contribute to overall global weathering.
+Global weatherability is the sum of factors aside from climate itself, such as the latitudinal distribution of continents and mountain belts, that contribute to overall global weathering.
 On a planet with high weatherability, the CO$_2$ input from volcanism can be removed via silicate weathering at a lower atmospheric CO$_2$ concentration than on a less weatherable planet.
@@
 
-In the original models that proposed the ``Fire and Ice'' hypothesis as an explanation for the onset of Sturtian glaciation, chemical weathering was modeled as an Arhennius relationship where it was solely a function of temperature and run-off \citep{Donnadieu2004a}.
+In the original models that proposed the `Fire and Ice' hypothesis as an explanation for the onset of the Sturtian `Snowball Earth' Glaciation, chemical weathering was modeled as an Arrhenius relationship where it was solely a function of temperature and run-off \citep{Donnadieu2004a}.
 However, such an approach neglects the effects that can result through soil shielding and regolith development in low-relief regions.
-Recent progress on understanding the relationships between landscapes, topography and chemical weathering reveals that these effects are quite important (e.g.
+Recent progress on understanding the relationships between landscapes, topography, and chemical weathering reveals that these effects are quite important (e.g.
 \citealp{Maher2014a}).
@@
 A couple considerations are relevant with respect to this analysis: 1) the current extent of LIP exposure is reduced in part by volcanics being covered by unconsolidated sediments (i.e.
-regolith development itself) in a number of the provinces; 2) the initial LIP extents are typically poorly constrained and are likely overestimates which could be resulting in inflated interpreted decay rates.
+regolith development itself) in a number of the provinces; 2) the current extent of LIP exposure may be incomplete for some of the provinces; 3) the initial LIP extents are typically poorly constrained and are likely overestimates which could be resulting in inflated interpreted decay rates.
 Future efforts to develop better-constrained estimates of original LIP extent will improve analyses such as that in this contribution.
 
-We have focused this analysis on the Phanerozoic record given that well-constrained paleogeographic models are available for the past $\sim$500 Myr.
-Furthermore, the approach of seeking to evaluate correlation between LIP area and glaciation is complicated for Neoproterozoic pan-glacial events wherein ice-albedo runaway leads to persistent global glaciation absent of continued forcing through normal carbon cycle processes until the build-up of sufficient pCO$_2$ to drive deglaciation.
+We have focused this analysis on the Phanerozoic record given that well-constrained paleogeographic models are available for the past $\sim$500~Myr.
+Furthermore, the approach of seeking to evaluate correlation between LIP area and glaciation is complicated for Neoproterozoic Snowball Earth events, since ice-albedo runaway leads to persistent global glaciation absent of continued forcing through normal carbon cycle processes until the build-up of sufficient pCO$_2$ to drive deglaciation.
 Nevertheless, the  hypothesis of tropical LIP area associated with the ca.
-720 Ma Franklin LIP increasing global weatherability and contributing to the onset of the Sturtian Snowball Earth glaciation is a major motivating driver behind conducting this analysis.
+720~Ma Franklin LIP increasing global weatherability and contributing to the onset of the Sturtian Snowball Earth Glaciation is a major motivating driver behind conducting this analysis.
 How does the tropical LIP area associated with the Franklin LIP compare to that observed in the Phanerozoic?
-DISCUSS FRANKLIN AND UMKONDO AREAS AND COMPARE TO THE PHANEROZOIC RECORD.
+Using the paleomagnetic pole of \citet{Denyszyn2009a}, we reconstruct the paleolatitude of the Franklin LIP at the time of emplacement, and find that $\sim$99.7\% (or $\sim$2.6~Mkm$^{2}$) of the LIP erupted within 15\textdegree\xspace of the equator.
+This Franklin LIP tropical area at the time of emplacement is approximately equivalent to the minor Cenozoic peak, and is significantly smaller than the other Phanerozoic peaks (Fig.
+\ref{fig:LIP_area}C).
+The ca.
+1109~Ma Umkondo is another Precambrian LIP that is constrained to have erupted in the tropics, although it is not known to be associated with any glaciation.
+We reconstruct the paleolatitude of the Umkondo LIP at the time of emplacement using the paleomagnetic pole of \citet{Swanson-Hysell2015b}, and find that effectively all of the LIP (or $\sim$2.0~Mkm$^{2}$) erupted within the tropics, an area that is slightly smaller than that for the Franklin LIP (Fig.
+\ref{fig:LIP_area}C).
+Together, these results indicate that the tropical area of the Franklin LIP, when compared that of Phanerozoic as well as other Precambrian LIPs, is not particularly notable, and therefore that the increase in global weatherability associated with the emplacement of the Franklin LIP in the tropics is also not uniquely large, especially when considering the severity of the glaciation that followed it.
+This conclusion is consistent with the results of our Phanerozoic analysis, which suggests that LIP area and its associated impact on global weatherability is not the fundamental control on Earth's long-term climate state.
 
@@
 \noindent\makebox[\textwidth]{
-\begin{tabular}{c c c c c}
-\hline
-\textbf{Fraction} & \textbf{$^{206}$Pb/$^{238}$U} & & \textbf{$^{207}$Pb/$^{235}$U} & \\
-& \textbf{Age [Ma]} & $\pm$2$\sigma$ & \textbf{Age [Ma]} & $\pm$2$\sigma$ \\
-\hline
-z1 & 823.2201669 & 1.067507326 & 823.4322744 & 2.609783978 \\
-z2 & 822.401975 & 0.714471249 & 819.0720505 & 1.260737 \\
-z4 & 836.2094848& 1.478314118 & 828.1995093 & 6.104877613 \\
-z5 & 834.109136 & 1.288954576 & 825.9918302 & 5.720788585 \\
-z16 & 831.1612867 & 0.741238812 & 832.226412 & 1.64433912 \\
-z17 & 854.9767495 & 0.72532716 & 855.8348689 & 1.16873047 \\
-z18 & 812.4046191 & 0.665110984 & 818.3985038 & 1.082055988 \\
-z19 & 832.8963617 & 0.694437022 & 833.1583004 & 1.060745556 \\
-\hline
+\begin{tabular}{lrlrrlrrl}
+\toprule
+                name &    age &                     age ref.
+&  original area Mkm2 &  present area Mkm2 &                                 present area ref.
+&  present/original &  half-life &   buried \\
+\midrule
+      Columbia River &   16.0 &   Kasbohm and Schoene (2018) &                0.68 &               0.38 &                           Buchan and Ernst (2004) &              0.56 &       19.1 &       no \\
+                Afar &   30.0 &  Courtillot and Renne (2003) &                2.05 &               0.63 &                              Coffin et al.
+(2006) &              0.31 &       17.8 &  partial \\
+                NAIP &   62.0 &         Larsen et al.
+(2016) &                1.07 &               0.29 &     Buchan and Ernst (2004), Coffin et al.
+(2006) &              0.27 &       32.8 &  partial \\
+              Deccan &   66.0 &               Schoene (2015) &                0.83 &               0.56 &                              Coffin et al.
+(2006) &              0.68 &      118.6 &       no \\
+          Seychelles &   66.0 &               Schoene (2015) &                0.46 &               0.00 &                              Coffin et al.
+(2006) &              0.00 &        0.0 &      yes \\
+          Madagascar &   90.0 &    Cucciniello et al.
+(2010) &                0.63 &               0.03 &                              Coffin et al.
+(2006) &              0.05 &       20.8 &       no \\
+ Caribbean-Colombian &   94.0 &         Loewen et al.
+(2013) &                0.71 &               0.13 &                              Coffin et al.
+(2006) &              0.18 &       38.0 &       no \\
+               HALIP &   95.0 &      Kingsbury et al.
+(2018) &                3.60 &               0.15 &                      Moosdorf and Hartmann (2015) &              0.04 &       20.5 &       no \\
+              EQUAMP &  131.0 &       Hollanda et al.
+(2016) &                0.66 &               0.01 &                            Hollanda et al.
+(2016) &              0.01 &       19.7 &       no \\
+               Comei &  132.0 &            Zhu et al.
+(2009) &                0.11 &                NaN &                                               n/a &               NaN &        NaN &       no \\
+             Bunbury &  132.0 &            Zhu et al.
+(2009) &                0.03 &               0.00 &                              Thorne et al.
+(2014) &              0.05 &       30.5 &       no \\
+     Parana-Etendeka &  135.0 &      Florisbal et al.
+(2014) &                3.12 &               0.40 &  Coffin et al.
+(2016), Teixeira and Albrez (2017) &              0.13 &       45.9 &  partial \\
+                Trap &  140.0 &          Ernst et al.
+(2001) &                0.03 &               0.00 &                           Ernst and Buchan (2001) &              0.00 &        0.0 &       no \\
+ NW Australia Margin &  160.0 &            COME BACK TO THIS &                0.62 &               0.00 &                              Coffin et al.
+(2006) &              0.00 &        0.0 &      yes \\
+               Karoo &  183.0 &        Burgess et al.
+(2015) &                3.21 &               0.15 &                               de Kock compilation &              0.05 &       42.3 &       no \\
+              Ferrar &  183.0 &        Burgess et al.
+(2015) &                0.18 &                NaN &                                               n/a &               NaN &        NaN &       no \\
+                CAMP &  201.0 &      Blackburn et al.
+(2013) &               11.46 &               0.23 &                   Marzoli and Parisio compilation &              0.02 &       35.6 &  partial \\
+             Siberia &  252.0 &   Burgess and Bowring (2015) &                3.46 &               0.47 &                              Coffin et al.
+(2006) &              0.14 &       88.8 &       no \\
+            Emeishan &  259.0 &           Zhou et al.
+(2002) &                0.71 &               0.06 &                              Coffin et al.
+(2006) &              0.09 &       74.6 &       no \\
+    Panjal-Qiangtang &  283.0 &           Zhai et al.
+(2013) &                0.11 &                NaN &                                               n/a &               NaN &        NaN &       no \\
+               Tarim &  290.0 &             Xu et al.
+(2014) &                0.35 &                NaN &                                               n/a &               NaN &        NaN &       no \\
+            Magdalen &  360.0 &         Murphy et al.
+(1999) &                0.42 &                NaN &                                               n/a &               NaN &        NaN &       no \\
+              Vilyui &  374.0 &          Renne et al.
+(2010) &                1.14 &                NaN &                                               n/a &               NaN &        NaN &       no \\
+        Kola-Dnieper &  380.0 &    Arzamastsev and Wu (2014) &                5.90 &                NaN &                                               n/a &               NaN &        NaN &       no \\
+            Suordakh &  450.0 &       Khudoley et al.
+(2013) &                0.02 &                NaN &                                               n/a &               NaN &        NaN &       no \\
+         Kalkarindji &  511.0 &        Jourdan et al.
+(2014) &                3.54 &               0.17 &                              Thorne et al.
+(2014) &              0.05 &      118.2 &       no \\
+            Franklin &  720.0 &       Denyszyn et al.
+(2009) &                2.62 &               0.04 &                           Buchan and Ernst (2004) &              0.02 &      127.6 &       no \\
+\bottomrule
 \end{tabular}

### 2018-10-16 Manuscript/Manuscript.tex

@@
 \usepackage{setspace}
-\usepackage{}
+\usepackage{array}
 \usepackage{fancyhdr}
@@
 
-XXX
+One of the hypothesized effects of the emplacement of large igneous provinces (LIPs) is planetary cooling on million-year timescales associated with enhanced silicate weathering.
+A systematic analysis of LIP area through the Phanerozoic reveals that there is not a significant correlation between total LIP area, nor LIP area in the tropics, and the extent of continental ice sheets.
+The largest peaks in tropical LIP area are at times of non-glacial climate.
+Changes in planetary weatherability associated with LIPs are therefore not the fundamental control on whether Earth is in a glacial or non-glacial climate although they could provide a secondary modulating effect in conjunction with other processes.
 
@@
 However, this approach likely provide the best estimates available of original extent for ancient LIPs.
-The extent of presently exposed volcanics associated with the LIPs were taken from a number of resources including the PLATES compilation (Table \ref{tab:LIPs}; \citealp{Coffin2006a}) and more recent compilation efforts associated with the Large Igneous Provinces Commission.
+The extents of presently-exposed volcanics associated with LIPs were taken from a number of resources including the PLATES compilation (Table \ref{tab:LIPs}; \citealp{Coffin2006a}) and more recent compilation efforts associated with the Large Igneous Provinces Commission.
 
-After LIPs are emplaced they will progressively erode.
+After LIPs are emplaced, they will progressively erode.
 In order to account for this decrease in area through with time, \citet{Godderis2017a} took the approach of fitting estimates of the original surface extent and estimates of the current extent for 5 LIPs with an exponential decay function.
@@
 \ref{fig:LIP_preservation}).
-There is variability in the individual half-lives for LIPs such that they range from shorter half-lives of $\sim$20~Myr up to $\sim$120~Myr for the Kalkarindji and Deccan Traps.
+There is variability in the half-lives for LIPs, if they are each individual fit with an exponential decay function, such that they range from shorter half-lives of $\sim$20~Myr up to $\sim$120~Myr for the Kalkarindji and Deccan Traps.
 In the analysis of LIP area through time, we implement decay scenarios informed by this analysis.
@@
 The LIP reconstruction used in this study includes pre-Phanerozoic LIPs.
-However, given the imposed exponential decay since emplacement, the inclusion of these LIPs do not significantly impact the calculated LIP areas throughout the Phanerozoic.
+However, given the imposed exponential decay since emplacement, the inclusion of these LIPs does not significantly impact the calculated LIP areas throughout the Phanerozoic Eon which is the focus of this analysis.
 
@@
 In contrast to the decay and decay+burial scenarios implemented on estimates of original LIP extent in this study, \citet{Johansson2018a} uses a static extent for each LIP throughout the reconstruction with some of the polygons corresponding to the present-day surface extent and some representing the original extent that includes currently buried portions of the LIP (e.g.
-the North American Mid-continent Rift).
+the North American Midcontinent Rift).
 
@@
 When we account for burial (`decay 1 + burial 1' and `decay 1 + burial 2' scenarios), only the latter two of these four peaks are affected -- the ca.
-200~Ma peak is attenuated/removed due to the partial/complete removal of the CAMP, and the Cenozoic peak is attenuated due to the partial/complete removal of the Afar LIP.
+200~Ma peak is attenuated/removed due to the partial/complete burial of the CAMP, and the Cenozoic peak is attenuated due to the partial/complete removal of the Afar LIP.
 However, after accounting for burial, a minor area of LIPs remain in the tropics from ca.
@@
 \ref{fig:LIP_correlation}).
-All other scenarios yield a weak negative correlation coefficient.
-Qualitatively, the weak positive correlation of the `decay 2 + burial 2' scenario relative to the other scenarios can be primarily attributed to the complete removal of the CAMP, which was emplaced during an extended interval of ice-free conditions, as well as the effect of the longer decay half-life extending the duration of the earlier two peaks, such that they overlap more with the Late Ordivician and Permo-Carboniferous glacial intervals.
+All other scenarios yield a near zero or weak negative correlation coefficient.
+Qualitatively, the weak positive correlation of the `decay 2 + burial 2' scenario relative to the other scenarios can be primarily attributed to the complete removal of the CAMP, which was emplaced during an extended interval of ice-free conditions, as well as the effect of the longer decay half-life extending the duration of the earlier two peaks, such that they overlap more with the Late Ordovician and Permo-Carboniferous glacial intervals.
 
 To assess the statistical significance of the raw Pearson correlation coefficients, we simulated the four glacial episodes (Fig.
-\ref{fig:LIP_area}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and \% overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these simulations.
-For all scenarios except for the `decay 2 + burial 2' scenario, more than half of the randomly timed glacial intervals correlate better with the LIP area in the tropics (Fig.
+\ref{fig:LIP_area}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and \% overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these 100,000 simulations.
+For the `decay 2 + burial 2' scenario, 28\% of the randomly timed glacial intervals correlate better with the LIP area in the tropics.
+With an associated p-value of 0.28, the null hypothesis that glacial intervals do not correlate to LIP area in the tropics cannot be rejected (Fig.
 \ref{fig:LIP_correlation}).
-For the `decay 2 + burial 2' scenario, 28\% of the randomly timed glacial intervals correlate better with the LIP area in the tropics.
-With an associated p-value of 0.28, the null hypothesis that glacial intervals do not correlate to LIP area in the tropics cannot be rejected.
+Taking this approach, none of the positive or negative correlations that emerge between the LIP area scenarios are statistically significant.
 
@@
 
-In the original models that proposed the `Fire and Ice' hypothesis as an explanation for the onset of the Sturtian `Snowball Earth' Glaciation, chemical weathering was modeled as an Arrhenius relationship where it was solely a function of temperature and run-off \citep{Donnadieu2004a}.
+In the original models that proposed the `Fire and Ice' hypothesis as an explanation for the onset of the Sturtian `Snowball Earth' glaciation, chemical weathering was modeled as an Arrhenius relationship where it was solely a function of temperature and run-off \citep{Donnadieu2004a}.
 However, such an approach neglects the effects that can result through soil shielding and regolith development in low-relief regions.
@@
 In contrast, processes that lead to continued exhumation of mafic lithologies and the creation of steep topography, particularly in tropical regions, may exert a strong control on global weatherability and long-term climate.
-This interpretation underlies the hypothesis that arc-continent collisons in the tropics during the Ordovician \citep{Swanson-Hysell2017a} and the Cenozoic \citep{Jagoutz2016a} played a significant role in transitions into glacial climate states at those times.
+This interpretation underlies the hypothesis that arc-continent collisions in the tropics during the Ordovician \citep{Swanson-Hysell2017a} and the Cenozoic \citep{Jagoutz2016a} played a significant role in transitions into glacial climate states at those times.
 
@@
 We have focused this analysis on the Phanerozoic record given that well-constrained paleogeographic models are available for the past $\sim$500~Myr.
-Furthermore, the approach of seeking to evaluate correlation between LIP area and glaciation is complicated for Neoproterozoic Snowball Earth events, since ice-albedo runaway leads to persistent global glaciation absent of continued forcing through normal carbon cycle processes until the build-up of sufficient pCO$_2$ to drive deglaciation.
+Furthermore, the approach of seeking to evaluate correlation between LIP area and glaciation is complicated for Neoproterozoic Snowball Earth events, since ice-albedo runaway leads to persistent global glaciation absent of continued forcing through normal carbon cycle processes until sufficient pCO$_2$ builds up in the absence of silicate weathering to drive deglaciation.
 Nevertheless, the hypothesis of tropical LIP area associated with the ca.
@@
 \ref{fig:LIP_area}C).
-Together, these results indicate that the tropical area of the Franklin LIP, when compared that of Phanerozoic as well as other Precambrian LIPs, is not particularly notable, and therefore that the increase in global weatherability associated with the emplacement of the Franklin LIP in the tropics is also not uniquely large, especially when considering the severity of the glaciation that followed it.
+Together, these results indicate that the tropical area of the Franklin LIP, when compared that of Phanerozoic as well as other Precambrian LIPs, did not have a singularly large are in the tropics.
+Given that similar peaks in tropical LIP area are not associated with the onset of glacial periods, additional processes must have been at play in drawing down atmospheric CO$_2$ in the lead-up to the Sturtian Snowball Earth.
+One such process could have been an increase in global weatherability associated with the accretion of arc terranes within terranes of the present-day Arabian-Nubian shield.
+An increase in global weatherability associated with the emplacement of the Franklin LIP in the tropics could have been a contributing factor on top of other such processes.
+The temporal overlap between Franklin LIP eruptions \citep{Denyszyn2009a} and the initiation of Sturtian glaciation \citep{Macdonald2010a, MacLennan2018a} remains compelling.
+This overlap continues to suggest that other aspects of LIP emplacement, such as the injection of sulfur aerosols in the stratosphere \citep{Macdonald2017a}, could have played a role in the initiation of low-latitude glaciation.
+injection is also not uniquely large, especially when considering the severity of the glaciation that followed it.
 This conclusion is consistent with the results of our Phanerozoic analysis, which suggests that LIP area and its associated impact on global weatherability is not the fundamental control on Earth's long-term climate state.
-
-\section*{CONCLUSIONS}
 
@@
 
+\clearpage
+\newpage
+
 \section*{TABLES}
@@
 \scriptsize
-\caption{place holder}
-\noindent\makebox[\textwidth]{
-\begin{tabular}{lrlrrlrrl}
+\caption{Phanerozoic (and the Franklin) large igneous provinces.}
+\resizebox{\textwidth}{!}{
+\begin{tabular}{lc>{\raggedright}p{3.5cm}cc>{\raggedright}p{3.5cm}ccl}
 \toprule
-                name &    age &                     age ref.
-&  original area Mkm2 &  present area Mkm2 &                                 present area ref.
-&  present/original &  half-life &   buried \\
+name & age & age ref. & original & present & present area ref. & present/ & half-life & buried? \\
+ & (Ma) & & area & area & & original & (Myr) & \\
+ & & & (Mkm$^{2}$) & (Mkm$^{2}$) & & & & \\
 \midrule
-      Columbia River &   16.0 &   Kasbohm and Schoene (2018) &                0.68 &               0.38 &                           Buchan and Ernst (2004) &              0.56 &       19.1 &       no \\
-                Afar &   30.0 &  Courtillot and Renne (2003) &                2.05 &               0.63 &                              Coffin et al.
-(2006) &              0.31 &       17.8 &  partial \\
-                NAIP &   62.0 &         Larsen et al.
-(2016) &                1.07 &               0.29 &     Buchan and Ernst (2004), Coffin et al.
-(2006) &              0.27 &       32.8 &  partial \\
-              Deccan &   66.0 &               Schoene (2015) &                0.83 &               0.56 &                              Coffin et al.
-(2006) &              0.68 &      118.6 &       no \\
-          Seychelles &   66.0 &               Schoene (2015) &                0.46 &               0.00 &                              Coffin et al.
-(2006) &              0.00 &        0.0 &      yes \\
-          Madagascar &   90.0 &    Cucciniello et al.
-(2010) &                0.63 &               0.03 &                              Coffin et al.
-(2006) &              0.05 &       20.8 &       no \\
- Caribbean-Colombian &   94.0 &         Loewen et al.
-(2013) &                0.71 &               0.13 &                              Coffin et al.
-(2006) &              0.18 &       38.0 &       no \\
-               HALIP &   95.0 &      Kingsbury et al.
-(2018) &                3.60 &               0.15 &                      Moosdorf and Hartmann (2015) &              0.04 &       20.5 &       no \\
-              EQUAMP &  131.0 &       Hollanda et al.
-(2016) &                0.66 &               0.01 &                            Hollanda et al.
-(2016) &              0.01 &       19.7 &       no \\
-               Comei &  132.0 &            Zhu et al.
-(2009) &                0.11 &                NaN &                                               n/a &               NaN &        NaN &       no \\
-             Bunbury &  132.0 &            Zhu et al.
-(2009) &                0.03 &               0.00 &                              Thorne et al.
-(2014) &              0.05 &       30.5 &       no \\
-     Parana-Etendeka &  135.0 &      Florisbal et al.
-(2014) &                3.12 &               0.40 &  Coffin et al.
-(2016), Teixeira and Albrez (2017) &              0.13 &       45.9 &  partial \\
-                Trap &  140.0 &          Ernst et al.
-(2001) &                0.03 &               0.00 &                           Ernst and Buchan (2001) &              0.00 &        0.0 &       no \\
- NW Australia Margin &  160.0 &            COME BACK TO THIS &                0.62 &               0.00 &                              Coffin et al.
-(2006) &              0.00 &        0.0 &      yes \\
-               Karoo &  183.0 &        Burgess et al.
-(2015) &                3.21 &               0.15 &                               de Kock compilation &              0.05 &       42.3 &       no \\
-              Ferrar &  183.0 &        Burgess et al.
-(2015) &                0.18 &                NaN &                                               n/a &               NaN &        NaN &       no \\
-                CAMP &  201.0 &      Blackburn et al.
-(2013) &               11.46 &               0.23 &                   Marzoli and Parisio compilation &              0.02 &       35.6 &  partial \\
-             Siberia &  252.0 &   Burgess and Bowring (2015) &                3.46 &               0.47 &                              Coffin et al.
-(2006) &              0.14 &       88.8 &       no \\
-            Emeishan &  259.0 &           Zhou et al.
-(2002) &                0.71 &               0.06 &                              Coffin et al.
-(2006) &              0.09 &       74.6 &       no \\
-    Panjal-Qiangtang &  283.0 &           Zhai et al.
-(2013) &                0.11 &                NaN &                                               n/a &               NaN &        NaN &       no \\
-               Tarim &  290.0 &             Xu et al.
-(2014) &                0.35 &                NaN &                                               n/a &               NaN &        NaN &       no \\
-            Magdalen &  360.0 &         Murphy et al.
-(1999) &                0.42 &                NaN &                                               n/a &               NaN &        NaN &       no \\
-              Vilyui &  374.0 &          Renne et al.
-(2010) &                1.14 &                NaN &                                               n/a &               NaN &        NaN &       no \\
-        Kola-Dnieper &  380.0 &    Arzamastsev and Wu (2014) &                5.90 &                NaN &                                               n/a &               NaN &        NaN &       no \\
-            Suordakh &  450.0 &       Khudoley et al.
-(2013) &                0.02 &                NaN &                                               n/a &               NaN &        NaN &       no \\
-         Kalkarindji &  511.0 &        Jourdan et al.
-(2014) &                3.54 &               0.17 &                              Thorne et al.
-(2014) &              0.05 &      118.2 &       no \\
-            Franklin &  720.0 &       Denyszyn et al.
-(2009) &                2.62 &               0.04 &                           Buchan and Ernst (2004) &              0.02 &      127.6 &       no \\
+Columbia River & 16 & Kasbohm and Schoene (2018) & 0.68 & 0.38 & Buchan and Ernst (2004) & 0.56 & 19.2 & no \\
+Afar & 30 & Courtillot and Renne (2003) & 2.05 & 0.63 & Coffin et al. (2006) & 0.31 & 17.7 & partial \\
+NAIP & 62 & Larsen et al. (2016) & 1.07 & 0.29 & Buchan and Ernst (2004), Coffin et al. (2006) & 0.27 & 33.0 & partial \\
+Deccan & 66 & Schoene (2015) & 0.83 & 0.56 & Coffin et al. (2006) & 0.68 & 116.6 & no \\
+Seychelles & 66 & Schoene (2015) & 0.46 & 0.00 & Coffin et al. (2006) & 0.00 & 0.0 & yes \\
+Madagascar & 90 & Cucciniello et al. (2010) & 0.63 & 0.03 & Coffin et al. (2006) & 0.05 & 20.8 & no \\
+Caribbean-Colombian & 94 & Loewen et al. (2013) & 0.71 & 0.13 & Coffin et al. (2006) & 0.18 & 37.6 & no \\
+HALIP & 95 & Kingsbury et al. (2018) & 3.60 & 0.15 & Moosdorf and Hartmann (2015) & 0.04 & 20.7 & no \\
+EQUAMP & 131 & Hollanda et al. (2016) & 0.66 & 0.01 & Hollanda et al. (2016) & 0.01 & 20.5 & no \\
+Comei & 132 & Zhu et al. (2009) & 0.11 & - & - & - & - & no \\
+Bunbury & 132 & Zhu et al. (2009) & 0.03 & 0.00 & Thorne et al. (2014) & 0.05 & 30.9 & no \\
+Parana-Etendeka & 135 & Florisbal et al. (2014) & 3.12 & 0.40 & Coffin et al. (2016), Teixeira and Albrez (2017) & 0.13 & 45.7 & partial \\
+Trap & 140 & Ernst et al. (2001) & 0.03 & 0.00 & Ernst and Buchan (2001) & 0.00 & 0.0 & no \\
+NW Australia Margin & 160 & COME BACK TO THIS & 0.62 & 0.00 & Coffin et al. (2006) & 0.00 & 0.0 & yes \\
+Karoo & 183 & Burgess et al. (2015) & 3.21 & 0.15 & de Kock compilation & 0.05 & 41.3 & no \\
+Ferrar & 183 & Burgess et al. (2015) & 0.18 & - & - & - & - & no \\
+CAMP & 201 & Blackburn et al. (2013) & 11.46 & 0.23 & Marzoli and Parisio compilation & 0.02 & 35.7 & partial \\
+Siberia & 252 & Burgess and Bowring (2015) & 3.46 & 0.47 & Coffin et al. (2006) & 0.14 & 87.5 & no \\
+Emeishan & 259 & Zhou et al. (2002) & 0.71 & 0.06 & Coffin et al. (2006) & 0.09 & 72.9 & no \\
+Panjal-Qiangtang & 283 & Zhai et al. (2013) & 0.11 & - & - & - & - & no \\
+Tarim & 290 & Xu et al. (2014) & 0.35 & - & - & - & - & no \\
+Magdalen & 360 & Murphy et al. (1999) & 0.42 & - & - & - & - & no \\
+Vilyui & 374 & Renne et al. (2010) & 1.14 & - & - & - & - & no \\
+Kola-Dnieper & 380 & Arzamastsev and Wu (2014) & 5.90 & - & - & - & - & no \\
+Suordakh & 450 & Khudoley et al. (2013) & 0.02 & - & - & - & - & no \\
+Kalkarindji & 511 & Jourdan et al. (2014) & 3.54 & 0.17 & Thorne et al. (2014) & 0.05 & 116.3 & no \\
+Franklin & 720 & Denyszyn et al. (2009) & 2.62 & 0.04 & Buchan and Ernst (2004) & 0.02 & 121.8 & no \\
 \bottomrule
@@
 }
-\begin{tablenotes}[para,flushleft]
-place holder.
-\end{tablenotes}
 \label{tab:LIPs}
 \end{table}
+
+\clearpage
+\newpage
 
@@
 \begin{center}
-	\includegraphics[width=0.65\textwidth]{Manuscript/Figures/LIP_Preservation.pdf}
+	\includegraphics[width=0.6\textwidth]{Manuscript/Figures/LIP_Preservation.pdf}
 	\caption{LIP preservation through time.

### 2018-10-16 Manuscript/Figures/LIP_Table.tex
calculate individual halflives BEFORE rounding, and update figure labels
add many references for the LIPs table

@@
 \toprule
-                name &    age &                     age ref.
-&  original area Mkm2 &  present area Mkm2 &                                 present area ref.
-&  present/original &  half-life &   buried \\
+                name &    age &                      age ref. &  original area Mkm2 &  present area Mkm2 &                              present area ref. &  present/original &  half-life &   buried \\
 \midrule
-      Columbia River &   16.0 &   Kasbohm and Schoene (2018) &                0.68 &               0.38 &                           Buchan and Ernst (2004) &              0.56 &       19.1 &       no \\
-                Afar &   30.0 &  Courtillot and Renne (2003) &                2.05 &               0.63 &                              Coffin et al.
-(2006) &              0.31 &       17.8 &  partial \\
-                NAIP &   62.0 &         Larsen et al.
-(2016) &                1.07 &               0.29 &     Buchan and Ernst (2004), Coffin et al.
-(2006) &              0.27 &       32.8 &  partial \\
-              Deccan &   66.0 &               Schoene (2015) &                0.83 &               0.56 &                              Coffin et al.
-(2006) &              0.68 &      118.6 &       no \\
-          Seychelles &   66.0 &               Schoene (2015) &                0.46 &               0.00 &                              Coffin et al.
-(2006) &              0.00 &        0.0 &      yes \\
-          Madagascar &   90.0 &    Cucciniello et al.
-(2010) &                0.63 &               0.03 &                              Coffin et al.
-(2006) &              0.05 &       20.8 &       no \\
- Caribbean-Colombian &   94.0 &         Loewen et al.
-(2013) &                0.71 &               0.13 &                              Coffin et al.
-(2006) &              0.18 &       38.0 &       no \\
-               HALIP &   95.0 &      Kingsbury et al.
-(2018) &                3.60 &               0.15 &                      Moosdorf and Hartmann (2015) &              0.04 &       20.5 &       no \\
-              EQUAMP &  131.0 &       Hollanda et al.
-(2016) &                0.66 &               0.01 &                            Hollanda et al.
-(2016) &              0.01 &       19.7 &       no \\
-               Comei &  132.0 &            Zhu et al.
-(2009) &                0.11 &                NaN &                                               n/a &               NaN &        NaN &       no \\
-             Bunbury &  132.0 &            Zhu et al.
-(2009) &                0.03 &               0.00 &                              Thorne et al.
-(2014) &              0.05 &       30.5 &       no \\
-     Parana-Etendeka &  135.0 &      Florisbal et al.
-(2014) &                3.12 &               0.40 &  Coffin et al.
-(2016), Teixeira and Albrez (2017) &              0.13 &       45.9 &  partial \\
-                Trap &  140.0 &          Ernst et al.
-(2001) &                0.03 &               0.00 &                           Ernst and Buchan (2001) &              0.00 &        0.0 &       no \\
- NW Australia Margin &  160.0 &            COME BACK TO THIS &                0.62 &               0.00 &                              Coffin et al.
-(2006) &              0.00 &        0.0 &      yes \\
-               Karoo &  183.0 &        Burgess et al.
-(2015) &                3.21 &               0.15 &                               de Kock compilation &              0.05 &       42.3 &       no \\
-              Ferrar &  183.0 &        Burgess et al.
-(2015) &                0.18 &                NaN &                                               n/a &               NaN &        NaN &       no \\
-                CAMP &  201.0 &      Blackburn et al.
-(2013) &               11.46 &               0.23 &                   Marzoli and Parisio compilation &              0.02 &       35.6 &  partial \\
-             Siberia &  252.0 &   Burgess and Bowring (2015) &                3.46 &               0.47 &                              Coffin et al.
-(2006) &              0.14 &       88.8 &       no \\
-            Emeishan &  259.0 &           Zhou et al.
-(2002) &                0.71 &               0.06 &                              Coffin et al.
-(2006) &              0.09 &       74.6 &       no \\
-    Panjal-Qiangtang &  283.0 &           Zhai et al.
-(2013) &                0.11 &                NaN &                                               n/a &               NaN &        NaN &       no \\
-               Tarim &  290.0 &             Xu et al.
-(2014) &                0.35 &                NaN &                                               n/a &               NaN &        NaN &       no \\
-            Magdalen &  360.0 &         Murphy et al.
-(1999) &                0.42 &                NaN &                                               n/a &               NaN &        NaN &       no \\
-              Vilyui &  374.0 &          Renne et al.
-(2010) &                1.14 &                NaN &                                               n/a &               NaN &        NaN &       no \\
+      Columbia River &   16.0 &    Kasbohm and Schoene (2018) &                0.68 &               0.38 &                        Buchan and Ernst (2004) &              0.56 &       19.2 &       no \\
+                Afar &   30.0 &   Courtillot and Renne (2003) &                2.05 &               0.63 &                           Coffin et al. (2006) &              0.31 &       17.7 &  partial \\
+                NAIP &   62.0 &          Larsen et al. (2015) &                1.07 &               0.29 &  Buchan and Ernst (2004), Coffin et al. (2006) &              0.27 &       33.0 &  partial \\
+              Deccan &   66.0 &         Schoene et al. (2014) &                0.83 &               0.56 &                           Coffin et al. (2006) &              0.68 &      116.6 &       no \\
+          Seychelles &   66.0 &         Schoene et al. (2014) &                0.46 &               0.00 &                           Coffin et al. (2006) &              0.00 &        0.0 &      yes \\
+          Madagascar &   90.0 &            Cucciniello (2010) &                0.63 &               0.03 &                           Coffin et al. (2006) &              0.05 &       20.8 &       no \\
+ Caribbean-Colombian &   94.0 &          Loewen et al. (2013) &                0.71 &               0.13 &                           Coffin et al. (2006) &              0.18 &       37.6 &       no \\
+               HALIP &   95.0 &       Kingsbury et al. (2018) &                3.60 &               0.15 &                   Moosdorf and Hartmann (2015) &              0.04 &       20.7 &       no \\
+              EQUAMP &  131.0 &        Hollanda et al. (2016) &                0.66 &               0.01 &                         Hollanda et al. (2016) &              0.01 &       20.5 &       no \\
+               Comei &  132.0 &             Zhu et al. (2009) &                0.11 &                NaN &                                            n/a &               NaN &        NaN &       no \\
+             Bunbury &  132.0 &             Zhu et al. (2009) &                0.03 &               0.00 &                           Thorne et al. (2014) &              0.05 &       30.9 &       no \\
+     Parana-Etendeka &  135.0 &       Florisbal et al. (2014) &                3.12 &               0.40 &                           Coffin et al. (2006) &              0.13 &       45.7 &  partial \\
+                Trap &  140.0 &       Ernst and Buchan (2001) &                0.03 &               0.00 &                        Ernst and Buchan (2001) &              0.00 &        0.0 &       no \\
+ NW Australia Margin &  160.0 &    Pirajno and Hoatson (2012) &                0.62 &               0.00 &                           Coffin et al. (2006) &              0.00 &        0.0 &      yes \\
+               Karoo &  183.0 &         Burgess et al. (2015) &                3.21 &               0.15 &                            de Kock compilation &              0.05 &       41.3 &       no \\
+              Ferrar &  183.0 &         Burgess et al. (2015) &                0.18 &                NaN &                                            n/a &               NaN &        NaN &       no \\
+                CAMP &  201.0 &       Blackburn et al. (2013) &               11.46 &               0.23 &                Marzoli and Parisio compilation &              0.02 &       35.7 &  partial \\
+             Siberia &  252.0 &    Burgess and Bowring (2015) &                3.46 &               0.47 &                           Coffin et al. (2006) &              0.14 &       87.5 &       no \\
+            Emeishan &  259.0 &            Zhou et al. (2002) &                0.71 &               0.06 &                           Coffin et al. (2006) &              0.09 &       72.9 &       no \\
+    Panjal-Qiangtang &  283.0 &            Zhai et al. (2013) &                0.11 &                NaN &                                            n/a &               NaN &        NaN &       no \\
+               Tarim &  290.0 &              Xu et al. (2014) &                0.35 &                NaN &                                            n/a &               NaN &        NaN &       no \\
+            Magdalen &  360.0 &  Brendan Murphy et al. (1999) &                0.42 &                NaN &                                            n/a &               NaN &        NaN &       no \\
+              Vilyui &  374.0 &           Ricci et al. (2013) &                1.14 &                NaN &                                            n/a &               NaN &        NaN &       no \\
         Kola-Dnieper &  380.0 &     Arzamastsev and Wu (2014) &                5.90 &                NaN &                                            n/a &               NaN &        NaN &       no \\
-            Suordakh &  450.0 &       Khudoley et al.
-(2013) &                0.02 &                NaN &                                               n/a &               NaN &        NaN &       no \\
-         Kalkarindji &  511.0 &        Jourdan et al.
-(2014) &                3.54 &               0.17 &                              Thorne et al.
-(2014) &              0.05 &      118.2 &       no \\
-            Franklin &  720.0 &       Denyszyn et al.
-(2009) &                2.62 &               0.04 &                           Buchan and Ernst (2004) &              0.02 &      127.6 &       no \\
+            Suordakh &  450.0 &        Khudoley et al. (2013) &                0.02 &                NaN &                                            n/a &               NaN &        NaN &       no \\
+         Kalkarindji &  511.0 &         Jourdan et al. (2014) &                3.54 &               0.17 &                           Thorne et al. (2014) &              0.05 &      116.3 &       no \\
+            Franklin &  720.0 &        Denyszyn et al. (2009) &                2.62 &               0.04 &                        Buchan and Ernst (2004) &              0.02 &      121.8 &       no \\
 \bottomrule

### 2018-10-16 Manuscript/Manuscript.tex
Updates from ShareLaTeX

@@
 720~Ma, in concert with elevated runoff rates associated with continental break-up, has been implicated as a major contributor to the cooling that initiated the Sturtian snowball Earth \citep{Donnadieu2004b, Cox2016a}.
-In the Cenozoic, the movement of the Deccan LIP into the tropical rain belt has been implicated in drawing down CO$_2$ levels in the lead-up to Eocene glaciation of Antarctica \citep{Kent2008a}.
+In the Cenozoic, the movement of the Deccan LIP into the tropical rain belt, in concert with the low-latitude emplacement of the Ethiopian Traps, has been implicated in drawing down CO$_2$ levels in the lead-up to Oligocene glaciation of Antarctica \citep{Kent2008a,Kent2013a}.
 
@@
 For some LIPs, this approach may lead to an overestimate of original areal extent given that subsurface intrusions could extend over a broader area than the surface volcanics.
-The polygons also assume complete areal coverage by LIPs as they connect wide-spread remnants--also creating the potential for them being over-estimates.
-However, this approach likely provide the best estimates available of original extent for ancient LIPs.
+The polygons also assume complete areal coverage by LIPs as they connect wide-spread remnants -- also creating the potential for them being over-estimates.
+However, this approach likely provides the best estimates available of original extent for ancient LIPs.
 The extents of presently-exposed volcanics associated with LIPs were taken from a number of resources including the PLATES compilation (Table \ref{tab:LIPs}; \citealp{Coffin2006a}) and more recent compilation efforts associated with the Large Igneous Provinces Commission.
@@
 In the analysis of LIP area through time, we implement decay scenarios informed by this analysis.
-Scenario `decay 1' uses the best-fit half-life of 36 Myr, while `decay 2' uses the slower decay with a half-life of $\sim$120~Myr.
+Scenario `t$_{1/2}$ = 36 Myr' uses the best-fit half-life of 36 Myr, while `t$_{1/2}$ = 120 Myr' uses the slower decay.
 Given that the post-emplacement weathering and erosional history of each LIP should be dependent on the tectonic and climatic setting that each LIP experiences during and after emplacement, this approach is simplistic but provides a framework for analysis.
@@
 To account for the rapid decrease in exposed surface area that would result from burial by sediments in a rift basin, we impose two different burial scenarios for LIPs that are co-located with rifting.
-`Burial 1' imposes instantaneous burial of 50$\%$ of the LIP area while `burial 2' imposes instantaneous burial of the entire LIP as an end-member scenario.
+The `50\% burial' scenario imposes instantaneous burial of 50$\%$ of the LIP area while the `100\% burial' imposes instantaneous burial of the entire LIP as an end-member scenario.
 The LIP area analysis uses all of the available combinations of these distinct decay and burial scenarios.
@@
 
-In the `decay 1' scenario, we observe three primary peaks and one minor peak in the calculated LIP area within the tropics (Fig.
+In the `t$_{1/2}$ = 36 Myr' scenario, we observe three primary peaks and one minor peak in the calculated LIP area within the tropics (Fig.
 \ref{fig:LIP_area}C).
@@
 \ref{fig:LIP_area}A).
-When we account for burial (`decay 1 + burial 1' and `decay 1 + burial 2' scenarios), only the latter two of these four peaks are affected -- the ca.
+When we account for burial (`t$_{1/2}$ = 36 Myr + 50\% burial' and `t$_{1/2}$ = 36 Myr + 100\% burial' scenarios), only the latter two of these four peaks are affected -- the ca.
 200~Ma peak is attenuated/removed due to the partial/complete burial of the CAMP, and the Cenozoic peak is attenuated due to the partial/complete removal of the Afar LIP.
@@
 130~Ma onwards, due to the EQUAMP, Caribbean-Colombian, and Deccan LIPs.
-Using the longer decay half-life of 120~Myr (the `decay 2 + burial 2' scenario) increases the area of LIPs in the tropics at any given time step, and has the effect of extending the duration of each peak.
+Using the longer decay half-life of 120~Myr (the `t$_{1/2}$ = 120 Myr + 100\% burial' scenario) increases the area of LIPs in the tropics at any given time step, and has the effect of extending the duration of each peak.
 
@@
 \ref{fig:LIP_area}C) and the ice extent record (Fig.
-\ref{fig:LIP_area}E) is the scenario with the slow decay rate and complete burial of LIPs associated with rifting (the `decay 2 + burial 2' scenario; Fig.
+\ref{fig:LIP_area}E) is the scenario with the slow decay rate and complete burial of LIPs associated with rifting (the `t$_{1/2}$ = 120 Myr + 100\% burial' scenario; Fig.
 \ref{fig:LIP_correlation}).
 All other scenarios yield a near zero or weak negative correlation coefficient.
-Qualitatively, the weak positive correlation of the `decay 2 + burial 2' scenario relative to the other scenarios can be primarily attributed to the complete removal of the CAMP, which was emplaced during an extended interval of ice-free conditions, as well as the effect of the longer decay half-life extending the duration of the earlier two peaks, such that they overlap more with the Late Ordovician and Permo-Carboniferous glacial intervals.
+Qualitatively, the weak positive correlation of the `t$_{1/2}$ = 120 Myr + 100\% burial' scenario relative to the other scenarios can be primarily attributed to the complete removal of the CAMP, which was emplaced during an extended interval of ice-free conditions, as well as the effect of the longer decay half-life extending the duration of the earlier two peaks, such that they overlap more with the Late Ordovician and Permo-Carboniferous glacial intervals.
 
@@
 \ref{fig:LIP_area}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and \% overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these 100,000 simulations.
-For the `decay 2 + burial 2' scenario, 28\% of the randomly timed glacial intervals correlate better with the LIP area in the tropics.
+For the `t$_{1/2}$ = 120 Myr + 100\% burial' scenario, 28\% of the randomly timed glacial intervals correlate better with the LIP area in the tropics.
 With an associated p-value of 0.28, the null hypothesis that glacial intervals do not correlate to LIP area in the tropics cannot be rejected (Fig.

### 2018-10-16 Manuscript/Manuscript.tex

@@
 
-The proposed environmental impacts of large igneous province (LIP) eruptions are myriad (as summarized in \citealp{Ernst2017a}).
+The proposed environmental impacts of large igneous province (LIP) eruptions are myriad (as summarized in \citealp{Ernst2017a} and discussed throughout this volume).
 One aspect of LIP emplacement that has been hypothesized to relate to long-term global climate is the effect such emplacement could have on increasing global weatherability and driving cooling.
-Global weatherability is the sum of factors aside from climate itself, such as the latitudinal distribution of continents and mountain belts, that contribute to overall global weathering.
+Global weatherability is the sum of factors aside from climate itself, such as the latitudinal distribution of continents and mountain belts, that contribute to overall global weathering and associated CO$_2$ sinks.
 On a planet with high weatherability, the CO$_2$ input from volcanism can be removed via silicate weathering at a lower atmospheric CO$_2$ concentration than on a less weatherable planet.
 The relatively high concentrations of Ca and Mg (that ultimately sequester carbon through precipitation as carbonate) and the high weathering rates of mafic lithologies make it so that basaltic regions consume more CO$_2$ than regions where bedrock composition is closer to bulk continental crust \citep{Dessert2003a}.
-Data from basaltic watersheds show that chemical weathering rates are greatest in regions with high runoff, and as a result CO$_2$ consumption in basaltic regions is most pronounced in the tropical rain belt \citep{Dessert2003a, Hartmann2014a}.
+Data from basaltic watersheds show that chemical weathering rates are greatest in regions with high runoff, and as a result CO$_2$ consumption in basaltic regions is most pronounced in the tropical rain belt \citep{Dessert2003a, Hartmann2009a, Hartmann2014a}.
 
@@
 720~Ma, in concert with elevated runoff rates associated with continental break-up, has been implicated as a major contributor to the cooling that initiated the Sturtian snowball Earth \citep{Donnadieu2004b, Cox2016a}.
-In the Cenozoic, the movement of the Deccan LIP into the tropical rain belt, in concert with the low-latitude emplacement of the Ethiopian Traps, has been implicated in drawing down CO$_2$ levels in the lead-up to Oligocene glaciation of Antarctica \citep{Kent2008a,Kent2013a}.
+In the Cenozoic, the movement of the Deccan LIP into the tropical rain belt, in concert with the low-latitude emplacement of the Ethiopian Traps, has been implicated in drawing down CO$_2$ levels in the lead-up to Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
 
@@
 After LIPs are emplaced, they will progressively erode.
-In order to account for this decrease in area through with time, \citet{Godderis2017a} took the approach of fitting estimates of the original surface extent and estimates of the current extent for 5 LIPs with an exponential decay function.
+In order to account for this decrease in area with time, \citet{Godderis2017a} took the approach of fitting estimates of the original surface extent and estimates of the current extent for 5 LIPs with an exponential decay function.
 They used the resulting exponential decay constants to develop a first-order model to estimate changing LIP area through time.
@@
 In contrast to the decay and decay+burial scenarios implemented on estimates of original LIP extent in this study, \citet{Johansson2018a} uses a static extent for each LIP throughout the reconstruction with some of the polygons corresponding to the present-day surface extent and some representing the original extent that includes currently buried portions of the LIP (e.g.
-the North American Midcontinent Rift).
+the Keweenawan Midcontinent Rift for the which the implemented extent is from geophysical data and largely correponds to buried subsurface exposures).
 
@@
 \ref{fig:LIP_area}B and C).
+We use 15\textdegree~S to 15\textdegree~N as the definition of the tropical rain belt as those latitudes approximately correspond with precipitation greater than 1.0 m/yr in modern climatalogical data \citep{Kalnay1996a}.
+The consistent paleolatitudes of evaporite deposits in the subtropics demonstrate the stability of the large-scale atmospheric circulation that
 
@@
 \resizebox{\textwidth}{!}{
-\begin{tabular}{lc>{\raggedright}p{3.5cm}cc>{\raggedright}p{3.5cm}ccl}
+\begin{tabular}{lc>{\raggedright}p{4cm}cc>{\raggedright}p{4cm}ccl}
 \toprule
-name & age & age ref. & original & present & present area ref. & present/ & half-life & buried? \\
- & (Ma) & & area & area & & original & (Myr) & \\
- & & & (Mkm$^{2}$) & (Mkm$^{2}$) & & & & \\
+\textbf{name} & \textbf{age} & \textbf{age ref.} & \textbf{original} & \textbf{present} & \textbf{present area ref.} & \textbf{present/} & \textbf{half-life*} & \textbf{buried?} \\
+ & \textbf{(Ma)} & & \textbf{area} & \textbf{area} & & \textbf{original} & \textbf{(Myr)} & \\
+ & & & \textbf{(Mkm$^{2}$)} & \textbf{(Mkm$^{2}$)} & & & & \\
 \midrule
-Columbia River & 16 & Kasbohm and Schoene (2018) & 0.68 & 0.38 & Buchan and Ernst (2004) & 0.56 & 19.2 & no \\
-Afar & 30 & Courtillot and Renne (2003) & 2.05 & 0.63 & Coffin et al. (2006) & 0.31 & 17.7 & partial \\
-NAIP & 62 & Larsen et al. (2016) & 1.07 & 0.29 & Buchan and Ernst (2004), Coffin et al. (2006) & 0.27 & 33.0 & partial \\
-Deccan & 66 & Schoene (2015) & 0.83 & 0.56 & Coffin et al. (2006) & 0.68 & 116.6 & no \\
-Seychelles & 66 & Schoene (2015) & 0.46 & 0.00 & Coffin et al. (2006) & 0.00 & 0.0 & yes \\
-Madagascar & 90 & Cucciniello et al. (2010) & 0.63 & 0.03 & Coffin et al. (2006) & 0.05 & 20.8 & no \\
-Caribbean-Colombian & 94 & Loewen et al. (2013) & 0.71 & 0.13 & Coffin et al. (2006) & 0.18 & 37.6 & no \\
-HALIP & 95 & Kingsbury et al. (2018) & 3.60 & 0.15 & Moosdorf and Hartmann (2015) & 0.04 & 20.7 & no \\
-EQUAMP & 131 & Hollanda et al. (2016) & 0.66 & 0.01 & Hollanda et al. (2016) & 0.01 & 20.5 & no \\
-Comei & 132 & Zhu et al. (2009) & 0.11 & - & - & - & - & no \\
-Bunbury & 132 & Zhu et al. (2009) & 0.03 & 0.00 & Thorne et al. (2014) & 0.05 & 30.9 & no \\
-Parana-Etendeka & 135 & Florisbal et al. (2014) & 3.12 & 0.40 & Coffin et al. (2016), Teixeira and Albrez (2017) & 0.13 & 45.7 & partial \\
-Trap & 140 & Ernst et al. (2001) & 0.03 & 0.00 & Ernst and Buchan (2001) & 0.00 & 0.0 & no \\
-NW Australia Margin & 160 & COME BACK TO THIS & 0.62 & 0.00 & Coffin et al. (2006) & 0.00 & 0.0 & yes \\
-Karoo & 183 & Burgess et al. (2015) & 3.21 & 0.15 & de Kock compilation & 0.05 & 41.3 & no \\
-Ferrar & 183 & Burgess et al. (2015) & 0.18 & - & - & - & - & no \\
-CAMP & 201 & Blackburn et al. (2013) & 11.46 & 0.23 & Marzoli and Parisio compilation & 0.02 & 35.7 & partial \\
-Siberia & 252 & Burgess and Bowring (2015) & 3.46 & 0.47 & Coffin et al. (2006) & 0.14 & 87.5 & no \\
-Emeishan & 259 & Zhou et al. (2002) & 0.71 & 0.06 & Coffin et al. (2006) & 0.09 & 72.9 & no \\
-Panjal-Qiangtang & 283 & Zhai et al. (2013) & 0.11 & - & - & - & - & no \\
-Tarim & 290 & Xu et al. (2014) & 0.35 & - & - & - & - & no \\
-Magdalen & 360 & Murphy et al. (1999) & 0.42 & - & - & - & - & no \\
-Vilyui & 374 & Renne et al. (2010) & 1.14 & - & - & - & - & no \\
-Kola-Dnieper & 380 & Arzamastsev and Wu (2014) & 5.90 & - & - & - & - & no \\
-Suordakh & 450 & Khudoley et al. (2013) & 0.02 & - & - & - & - & no \\
-Kalkarindji & 511 & Jourdan et al. (2014) & 3.54 & 0.17 & Thorne et al. (2014) & 0.05 & 116.3 & no \\
-Franklin & 720 & Denyszyn et al. (2009) & 2.62 & 0.04 & Buchan and Ernst (2004) & 0.02 & 121.8 & no \\
+Columbia River & 16 & \cite{Kasbohm2018a} & 0.68 & 0.38 & \cite{Buchan2004a} & 0.56 & 19.2 & no \\
+Afar & 30 & \cite{Courtillot2003a} & 2.05 & 0.63 & \cite{Coffin2006a} & 0.31 & 17.7 & partial \\
+NAIP & 62 & \cite{Larsen2015a} & 1.07 & 0.29 & \cite{Buchan2004a, Coffin2006a} & 0.27 & 33.0 & partial \\
+Deccan & 66 & \cite{Schoene2014a} & 0.83 & 0.56 & \cite{Coffin2006a} & 0.68 & 116.6 & no \\
+Seychelles & 66 & \cite{Schoene2014a} & 0.46 & 0.00 & \cite{Coffin2006a} & 0.00 & 0.0 & yes \\
+Madagascar & 90 & \cite{Cucciniello2010a} & 0.63 & 0.03 & \cite{Coffin2006a} & 0.05 & 20.8 & no \\
+Caribbean-Colombian & 94 & \cite{Loewen2013a} & 0.71 & 0.13 & \cite{Coffin2006a} & 0.18 & 37.6 & no \\
+HALIP & 95 & \cite{Kingsbury2018a} & 3.60 & 0.15 & \cite{Moosdorf2015a} & 0.04 & 20.7 & no \\
+EQUAMP & 131 & \cite{Hollanda2016a} & 0.66 & 0.01 & \cite{Hollanda2016a} & 0.01 & 20.5 & no \\
+Comei & 132 & \cite{Zhu2009a} & 0.11 & - & - & - & - & no \\
+Bunbury & 132 & \cite{Zhu2009a} & 0.03 & 0.00 & \cite{Thorne2014a} & 0.05 & 30.9 & no \\
+Parana-Etendeka & 135 & \cite{Florisbal2014a} & 3.12 & 0.40 & \cite{Coffin2006a} & 0.13 & 45.7 & partial \\
+Trap & 140 & \cite{Ernst2001a} & 0.03 & 0.00 & \cite{Ernst2001a} & 0.00 & 0.0 & no \\
+NW Australia Margin & 160 & \cite{Pirajno2012a} & 0.62 & 0.00 & \cite{Coffin2006a} & 0.00 & 0.0 & yes \\
+Karoo & 183 & \cite{Burgess2015a} & 3.21 & 0.15 & de Kock compilation & 0.05 & 41.3 & no \\
+Ferrar & 183 & \cite{Burgess2015a} & 0.18 & - & - & - & - & no \\
+CAMP & 201 & \cite{Blackburn2013a} & 11.46 & 0.23 & Marzoli and Parisio compilation & 0.02 & 35.7 & partial \\
+Siberia & 252 & \cite{Burgess2015b} & 3.46 & 0.47 & \cite{Coffin2006a} & 0.14 & 87.5 & no \\
+Emeishan & 259 & \cite{Zhou2002a} & 0.71 & 0.06 & \cite{Coffin2006a} & 0.09 & 72.9 & no \\
+Panjal-Qiangtang & 283 & \cite{Zhai2013a} & 0.11 & - & - & - & - & no \\
+Tarim & 290 & \cite{Xu2014a} & 0.35 & - & - & - & - & no \\
+Magdalen & 360 & \cite{Brendan_Murphy1999a} & 0.42 & - & - & - & - & no \\
+Vilyui & 374 & \cite{Ricci2013a} & 1.14 & - & - & - & - & no \\
+Kola-Dnieper & 380 & \cite{Arzamastsev2014a} & 5.90 & - & - & - & - & no \\
+Suordakh & 450 & \cite{Khudoley2013a} & 0.02 & - & - & - & - & no \\
+Kalkarindji & 511 & \cite{Jourdan2014a} & 3.54 & 0.17 & \cite{Thorne2014a} & 0.05 & 116.3 & no \\
+Franklin & 720 & \cite{Denyszyn2009a} & 2.62 & 0.04 & \cite{Buchan2004a} & 0.02 & 121.8 & no \\
 \bottomrule
@@
 }
+\begin{tablenotes}[para,flushleft]
+\textbf{*}assuming exponential decay with form $N(t) = 2^{t/t_{1/2}}$
+\end{tablenotes}
 \label{tab:LIPs}

### 2018-10-16 Manuscript/Manuscript.tex

@@
 The largest peaks in tropical LIP area are at times of non-glacial climate.
-Changes in planetary weatherability associated with LIPs are therefore not the fundamental control on whether Earth is in a glacial or non-glacial climate although they could provide a secondary modulating effect in conjunction with other processes.
+These results suggest that changes in planetary weatherability associated with LIPs are not the fundamental control on whether Earth is in a glacial or non-glacial climate, although they could provide a secondary modulating effect in conjunction with other processes.
 
@@
 \ref{fig:LIP_area}B and C).
-We use 15\textdegree~S to 15\textdegree~N as the definition of the tropical rain belt as those latitudes approximately correspond with precipitation greater than 1.0 m/yr in modern climatalogical data \citep{Kalnay1996a}.
-The consistent paleolatitudes of evaporite deposits in the subtropics demonstrate the stability of the large-scale atmospheric circulation that
+We use 15\textdegree~S to 15\textdegree~N as the definition of the tropical rain belt as those latitudes approximately correspond with precipitation greater than 1.0~m/yr in modern climatalogical data \citep{Kalnay1996a}.
+The consistent paleolatitudes of evaporite deposits in the subtropics demonstrate the stability of the large-scale atmospheric circulation that gives rise to intense precipitation in the tropical rain belt \citep{Evans2006a}.
 
@@
 The latitudinal extent of land ice is reflective of Earth's overall climate and allows for the delineation of glacial and non-glacial climate states.
-The land ice record is an imperfect tracker, as it is insensitive to changes in temperatures during non-glacial intervals and is influenced by additional factors such as the physical geography of the continents during glacial intervals.
-Nevertheless, it forms a physical record of Earth's climate state through time.
+The land ice record is an imperfect tracker of climate, as it is insensitive to changes in temperatures during non-glacial intervals and is influenced by additional factors such as the physical geography of the continents during glacial intervals.
+Nevertheless, it forms a physical record of Earth's climate through time and delineates glacial and non-glacial climate states.
 We take two approaches for comparison between the LIP area reconstruction and the record of ice extent.
-The first is to look at the correlation between LIP area and the extent of ice away from the pole (Fig.
+The first is to analyze the correlation between LIP area and the extent of ice away from the pole (Fig.
 \ref{fig:LIP_correlation}).
-The second is to consider the overlap between intervals of high LIP area in the tropics (defined as LIP area in the tropics \textgreater30\% of the maximum in a given post-emplacement model) and intervals of glacial climate (defined as ice extent \textgreater10\textdegree\xspace from the poles).
+The second is to consider the overlap between intervals of high LIP area (defined as LIP area \textgreater30\% of the maximum in a given post-emplacement model) and intervals of glacial climate (defined as ice extent \textgreater10\textdegree\xspace from the poles).
 This overlap approach places less emphasis on the specific magnitudes of the compiled ice extent and LIP area records.
@@
 However, such proxy compilations are problematic as they are poorly calibrated in deep time, prone to alteration, and reliant on a myriad of assumptions.
-We thus prefer to use the latitudinal extent of land ice to reflect Earth's overall climate state.
+We thus prefer to use the latitudinal extent of land ice to reflect Earth's overall climate state throughout the Phanerozoic.
 
@@
 
-To assess the statistical significance of the raw Pearson correlation coefficients, we simulated the four glacial episodes (Fig.
+To assess the statistical significance of the correlation associated with the Pearson correlation coefficients, we simulated the four glacial episodes (Fig.
 \ref{fig:LIP_area}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and \% overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these 100,000 simulations.
+This approach accounts for the reality that spurious correlation can arise between auto-correlated data sets such as these; where each value is not independent, but is instead dependent on the previous state of the system.
 For the `t$_{1/2}$ = 120 Myr + 100\% burial' scenario, 28\% of the randomly timed glacial intervals correlate better with the LIP area in the tropics.
@@
 In the original models that proposed the `Fire and Ice' hypothesis as an explanation for the onset of the Sturtian `Snowball Earth' glaciation, chemical weathering was modeled as an Arrhenius relationship where it was solely a function of temperature and run-off \citep{Donnadieu2004a}.
-However, such an approach neglects the effects that can result through soil shielding and regolith development in low-relief regions.
-Recent progress on understanding the relationships between landscapes, topography, and chemical weathering reveals that these effects are quite important (e.g.
-\citealp{Maher2014a}).
+However, such an approach neglects the effects of soil shielding and regolith development in low-relief regions.
+Recent progress on understanding the relationships between landscapes, topography, and chemical weathering reveals that these effects are quite important \citeap{Gabet2009, Maher2014a}.
 As a result, more recent modeling of chemical weathering that has incorporated such processes highlights the importance of high-relief regions relative to low-relief ones \citep{Godderis2017b}.

### 2018-10-16 Manuscript/Manuscript.tex
Updates from ShareLaTeX

@@
 
-\noindent\textit{This article is in preparation for submission to \textit{Environmental Change and Large Igneous Provinces}.}
+\noindent\textit{This article is in preparation for submission to the AGU book \textit{Environmental Change and Large Igneous Provinces}.}
 
@@
 One of the hypothesized effects of the emplacement of large igneous provinces (LIPs) is planetary cooling on million-year timescales associated with enhanced silicate weathering.
-A systematic analysis of LIP area through the Phanerozoic reveals that there is not a significant correlation between total LIP area, nor LIP area in the tropics, and the extent of continental ice sheets.
+A systematic analysis of LIP area through the Phanerozoic reveals that there is no significant correlation between total LIP area, nor LIP area in the tropics, and the extent of continental ice sheets.
 The largest peaks in tropical LIP area are at times of non-glacial climate.
@@
 
-The proposed environmental impacts of large igneous province (LIP) eruptions are myriad (as summarized in \citealp{Ernst2017a} and discussed throughout this volume).
-One aspect of LIP emplacement that has been hypothesized to relate to long-term global climate is the effect such emplacement could have on increasing global weatherability and driving cooling.
-Global weatherability is the sum of factors aside from climate itself, such as the latitudinal distribution of continents and mountain belts, that contribute to overall global weathering and associated CO$_2$ sinks.
+Global weatherability is the sum of factors aside from climate itself, such as the latitudinal distribution of continents and mountain belts, that contribute to overall global weathering and associated CO$_2$ consumption.
 On a planet with high weatherability, the CO$_2$ input from volcanism can be removed via silicate weathering at a lower atmospheric CO$_2$ concentration than on a less weatherable planet.
@@
 
-Given these factors, the emplacement of LIPs in the tropics has been hypothesized to be associated with specific episodes of climatic cooling on Earth.
-In the Neoproterozoic Era, the emplacement of the Franklin LIP in the tropics ca.
-720~Ma, in concert with elevated runoff rates associated with continental break-up, has been implicated as a major contributor to the cooling that initiated the Sturtian snowball Earth \citep{Donnadieu2004b, Cox2016a}.
+One aspect of large igneous province (LIP) emplacement that has been hypothesized to relate to long-term global climate is the effect such emplacement could have on increasing global weatherability and driving cooling.
+In particular, the emplacement of LIPs in the tropics has been hypothesized to be associated with specific episodes of climatic cooling on Earth.
+In the Neoproterozoic era, the emplacement of the Franklin LIP in the tropics ca.
+720~Ma, in concert with elevated runoff rates associated with continental break-up, has been implicated as a major contributor to the cooling that initiated the Sturtian `Snowball Earth' \citep{Donnadieu2004b, Cox2016a}.
 In the Cenozoic, the movement of the Deccan LIP into the tropical rain belt, in concert with the low-latitude emplacement of the Ethiopian Traps, has been implicated in drawing down CO$_2$ levels in the lead-up to Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
@@
 \ref{fig:LIP_map}; Table \ref{tab:LIPs}).
-The compilation of \citet{Ernst2017a} seeks to reconstruct the original areal extent of LIPs with the caveat that there can be significant uncertainty with doing so, particularly for older more deeply eroded LIPs.
-These polygons encapsulate all of the rocks associated with a given LIP including dikes and sills (Fig.
+The compilation of \citet{Ernst2017a} seeks to reconstruct the original surface extent of LIPs with the caveat that there can be significant uncertainty with doing so, particularly for older more deeply eroded LIPs.
+These polygons encapsulate all of the rocks associated with a given LIP, including dikes and sills (Fig.
 \ref{fig:LIP_map}).
-For some LIPs, this approach may lead to an overestimate of original areal extent given that subsurface intrusions could extend over a broader area than the surface volcanics.
-The polygons also assume complete areal coverage by LIPs as they connect wide-spread remnants -- also creating the potential for them being over-estimates.
+For some LIPs, this approach may lead to an overestimate of original surface extent, given that subsurface intrusions could extend over a broader area than the surface volcanics.
+The polygons also assume complete surface coverage between wide-spread remnants, creating further potential for these outlines being over-estimates.
 However, this approach likely provides the best estimates available of original extent for ancient LIPs.
-The extents of presently-exposed volcanics associated with LIPs were taken from a number of resources including the PLATES compilation (Table \ref{tab:LIPs}; \citealp{Coffin2006a}) and more recent compilation efforts associated with the Large Igneous Provinces Commission.
+The extents of presently-exposed volcanics associated with LIPs were taken from a number of resources including the PLATES compilation \citep{Coffin2006a} and more recent compilation efforts associated with the Large Igneous Provinces Commission (Table \ref{tab:LIPs}).
 
 After LIPs are emplaced, they will progressively erode.
-In order to account for this decrease in area with time, \citet{Godderis2017a} took the approach of fitting estimates of the original surface extent and estimates of the current extent for 5 LIPs with an exponential decay function.
+In order to account for this decrease in area with time, \citet{Godderis2017a} took the approach of fitting an exponential decay function to estimates of the original surface extent and the current surface extent for 5 LIPs.
 They used the resulting exponential decay constants to develop a first-order model to estimate changing LIP area through time.
-We extend this approach to 19 basaltic LIPs for which we have estimates of the original areal extent of the province and the current areal extent of rocks associated with it (Fig.
+We extend this approach to 19 basaltic LIPs for which we have estimates of the original surface extent of the province and the current surface extent of rocks associated with it (Fig.
 \ref{fig:LIP_preservation}).
-While there are significant uncertainties associated with the area estimates, this compilation suggests that an exponential decay function is an appropriate approximate model for the progressive reduction in LIP area.
+While there are significant uncertainties associated with the area estimates, this compilation suggests that an exponential decay function is an appropriate first-order model for the progressive reduction in LIP area.
 The best-fit exponential function results in a LIP area half-life of 29~Myr.
-However, since we explicitly account for LIP burial separately in our models (see below), we prefer to exclude the 6 of 19 LIPs which have been partially or completely buried in our estimation of a representative LIP area half-life, which yields a slightly higher best-fit half-life of 36~Myr (Fig.
+However, since we explicitly account for LIP burial separately in our models (see below), we prefer to exclude the 6 of 19 LIPs which have been partially or completely buried in our estimation of a representative LIP area half-life.
+This latter approach yields a slightly higher best-fit half-life of 36~Myr (Fig.
 \ref{fig:LIP_preservation}).
-There is variability in the half-lives for LIPs, if they are each individual fit with an exponential decay function, such that they range from shorter half-lives of $\sim$20~Myr up to $\sim$120~Myr for the Kalkarindji and Deccan Traps.
+Although the root mean square error of the fit is low (0.14), there is variability in the half-lives for LIPs if they are fit individually with an exponential decay function, such that they range from shorter half-lives of $\sim$20~Myr up to $\sim$120~Myr for the Kalkarindji and Deccan Traps (Table \ref{tab:LIPs}).
 In the analysis of LIP area through time, we implement decay scenarios informed by this analysis.
-Scenario `t$_{1/2}$ = 36 Myr' uses the best-fit half-life of 36 Myr, while `t$_{1/2}$ = 120 Myr' uses the slower decay.
+Scenario `t$_{1/2}$ = 36~Myr' uses the best-fit half-life of 36~Myr, while `t$_{1/2}$ = 120~Myr' uses the slower decay.
 Given that the post-emplacement weathering and erosional history of each LIP should be dependent on the tectonic and climatic setting that each LIP experiences during and after emplacement, this approach is simplistic but provides a framework for analysis.
-The LIP reconstruction used in this study includes pre-Phanerozoic LIPs.
-However, given the imposed exponential decay since emplacement, the inclusion of these LIPs does not significantly impact the calculated LIP areas throughout the Phanerozoic Eon which is the focus of this analysis.
+The LIP reconstructions used in this study also include pre-Phanerozoic LIPs.
+However, given the imposed exponential decay since emplacement, the inclusion of these LIPs does not significantly impact the calculated LIP areas throughout the Phanerozoic, which is the focus of this analysis.
 
@@
 There are numerous examples in the record where there is partial or complete burial of a LIP associated with rifting and thermal subsidence (Table \ref{tab:LIPs}).
-For example, the Afar LIP is both associated with the Ethiopian Traps which form plateau flood basalts, and successful rifting in the region of the Red Sea that has resulted in burial (Fig \ref{fig:LIP_map}).
+For example, the Afar LIP is both associated with the Ethiopian Traps which form plateau flood basalts as well as successful rifting in the region of the Red Sea that has resulted in burial (Fig \ref{fig:LIP_map}).
 To account for the rapid decrease in exposed surface area that would result from burial by sediments in a rift basin, we impose two different burial scenarios for LIPs that are co-located with rifting.
@@
 The LIP area analysis uses all of the available combinations of these distinct decay and burial scenarios.
+
 Our LIP reconstruction differs from that of \citet{Johansson2018a} in several respects.
 In contrast to the decay and decay+burial scenarios implemented on estimates of original LIP extent in this study, \citet{Johansson2018a} uses a static extent for each LIP throughout the reconstruction with some of the polygons corresponding to the present-day surface extent and some representing the original extent that includes currently buried portions of the LIP (e.g.
-the Keweenawan Midcontinent Rift for the which the implemented extent is from geophysical data and largely correponds to buried subsurface exposures).
+the Keweenawan Midcontinent Rift for which the implemented extent is from geophysical data that largely correponds to buried subsurface exposures).
 
@@
 The LIP polygons and tectonic units were reconstructed from 520~Ma to the present utilizing the paleogeographic model of \citet{Torsvik2016a} in the spin axis reference frame (anchor plate ID of 1).
-This paleogeographic model was updated to include a revision to Ordovician Laurentia \citep{Swanson-Hysell2017a} and the Paleozoic of Asia \citep{Domeier2018a}.
+This paleogeographic model was updated to include a revision to Ordovician Laurentia \citep{Swanson-Hysell2017a} and Paleozoic Asia \citep{Domeier2018a}.
 Reconstructions and area calculations within latitude bands utilized the pyGPlates function library and custom Python scripts documented within a Jupyter notebook that reproduces the analysis and development of the associated visualizations.
@@
 
-To evaluate the relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic Eon (modified from \citealp{Royer2004a}; Fig.
+To evaluate the relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic (modified from \citealp{Royer2004a}; Fig.
 \ref{fig:LIP_area}E).
 The goal in doing so is to evaluate the hypothesis that there is a correlation between LIP area in the tropics and Earth's long-term climate state.
-The latitudinal extent of land ice is reflective of Earth's overall climate and allows for the delineation of glacial and non-glacial climate states.
-The land ice record is an imperfect tracker of climate, as it is insensitive to changes in temperatures during non-glacial intervals and is influenced by additional factors such as the physical geography of the continents during glacial intervals.
+The land ice record is an imperfect tracker of climate, as it is insensitive to changes in temperature during non-glacial intervals and is influenced by additional factors such as the physical geography of the continents during glacial intervals.
 Nevertheless, it forms a physical record of Earth's climate through time and delineates glacial and non-glacial climate states.
@@
 
-In the `t$_{1/2}$ = 36 Myr' scenario, we observe three primary peaks and one minor peak in the calculated LIP area within the tropics (Fig.
+In the `t$_{1/2}$ = 36~Myr' scenario, we observe three primary peaks and one minor peak in the calculated LIP area within the tropics (Fig.
 \ref{fig:LIP_area}C).
@@
 \ref{fig:LIP_area}A).
-When we account for burial (`t$_{1/2}$ = 36 Myr + 50\% burial' and `t$_{1/2}$ = 36 Myr + 100\% burial' scenarios), only the latter two of these four peaks are affected -- the ca.
+When we account for burial (`t$_{1/2}$ = 36~Myr + 50\% burial' and `t$_{1/2}$ = 36~Myr + 100\% burial' scenarios), only the latter two of these four peaks are affected -- the ca.
 200~Ma peak is attenuated/removed due to the partial/complete burial of the CAMP, and the Cenozoic peak is attenuated due to the partial/complete removal of the Afar LIP.
@@
 130~Ma onwards, due to the EQUAMP, Caribbean-Colombian, and Deccan LIPs.
-Using the longer decay half-life of 120~Myr (the `t$_{1/2}$ = 120 Myr + 100\% burial' scenario) increases the area of LIPs in the tropics at any given time step, and has the effect of extending the duration of each peak.
+Using the longer decay half-life of 120~Myr (the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario) increases the area of LIPs in the tropics at any given time step, and has the effect of extending the duration of each peak.
 
@@
 \ref{fig:LIP_area}C) and the ice extent record (Fig.
-\ref{fig:LIP_area}E) is the scenario with the slow decay rate and complete burial of LIPs associated with rifting (the `t$_{1/2}$ = 120 Myr + 100\% burial' scenario; Fig.
+\ref{fig:LIP_area}E) is the scenario with the slow decay rate and complete burial of LIPs associated with rifting (the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario; Fig.
 \ref{fig:LIP_correlation}).
 All other scenarios yield a near zero or weak negative correlation coefficient.
-Qualitatively, the weak positive correlation of the `t$_{1/2}$ = 120 Myr + 100\% burial' scenario relative to the other scenarios can be primarily attributed to the complete removal of the CAMP, which was emplaced during an extended interval of ice-free conditions, as well as the effect of the longer decay half-life extending the duration of the earlier two peaks, such that they overlap more with the Late Ordovician and Permo-Carboniferous glacial intervals.
+Qualitatively, the weak positive correlation of the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario relative to the other scenarios can be primarily attributed to the complete removal of the CAMP, which was emplaced during an extended interval of ice-free conditions, as well as the effect of the longer decay half-life extending the duration of the earlier two peaks, such that they overlap more with the Late Ordovician and Permo-Carboniferous glacial intervals.
 
-To assess the statistical significance of the correlation associated with the Pearson correlation coefficients, we simulated the four glacial episodes (Fig.
-\ref{fig:LIP_area}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and \% overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these 100,000 simulations.
-This approach accounts for the reality that spurious correlation can arise between auto-correlated data sets such as these; where each value is not independent, but is instead dependent on the previous state of the system.
-For the `t$_{1/2}$ = 120 Myr + 100\% burial' scenario, 28\% of the randomly timed glacial intervals correlate better with the LIP area in the tropics.
-With an associated p-value of 0.28, the null hypothesis that glacial intervals do not correlate to LIP area in the tropics cannot be rejected (Fig.
+To assess the statistical significance of the correlation (or lack thereof) implied by the Pearson correlation coefficients, we simulated the four glacial episodes (Fig.
+\ref{fig:LIP_area}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and \% overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these 100,000 simulations (Fig.
 \ref{fig:LIP_correlation}).
+This approach accounts for the reality that spurious correlation can arise between auto-correlated data sets such as these, where each value is not independent, but is instead dependent on the previous state of the system.
+For the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario, 28\% of the randomly timed glacial intervals correlate better with the LIP area in the tropics.
+With an associated p-value of 0.28, the null hypothesis that glacial intervals do not correlate to LIP area in the tropics cannot be rejected.
 Taking this approach, none of the positive or negative correlations that emerge between the LIP area scenarios are statistically significant.
@@
 
-The results from this analysis indicate that when the entire LIP database is considered, there is not a significant relationship between total LIP area or LIP area in the tropics and the extent of continental ice sheets (Fig.
+The results from this analysis indicate that when the entire LIP database is considered, there is no significant relationship between total LIP area or LIP area in the tropics and the extent of continental ice sheets (Fig.
 \ref{fig:LIP_correlation}).
@@
 However, such an approach neglects the effects of soil shielding and regolith development in low-relief regions.
-Recent progress on understanding the relationships between landscapes, topography, and chemical weathering reveals that these effects are quite important \citeap{Gabet2009, Maher2014a}.
+Recent progress on understanding the relationships between landscapes, topography, and chemical weathering reveals that these effects are quite important \citep{Gabet2009a, Maher2014a}.
 As a result, more recent modeling of chemical weathering that has incorporated such processes highlights the importance of high-relief regions relative to low-relief ones \citep{Godderis2017b}.
@@
 
-A complication with the interpretation of soil shielding and limited weathering of LIPs is the rapid area decay rate inferred from the comparison of current LIP extent to estimated original extent.
+A complication with the interpretation of soil shielding and limited weathering of LIPs is the rapid area decay rate inferred from the comparison of current LIP extent to estimated original extent (Fig.
+\ref{fig:LIP_preservation}).
 A couple considerations are relevant with respect to this analysis: 1) the current extent of LIP exposure is reduced in part by volcanics being covered by unconsolidated sediments (i.e.
-regolith development itself) in a number of the provinces; 2) the current extent of LIP exposure may be incomplete for some of the provinces; 3) the initial LIP extents are typically poorly constrained and are likely overestimates which could be resulting in inflated interpreted decay rates.
+regolith development itself) in a number of the provinces; 2) the current extent of LIP exposure may be incomplete and an underestimate for some of the provinces; 3) the initial LIP extents are typically poorly constrained and are likely overestimates which could be resulting in inflated interpreted decay rates; and 4) the relationship between LIP area and volume is poorly constrained.
 Future efforts to develop better-constrained estimates of original LIP extent will improve analyses such as that in this contribution.
 
-We have focused this analysis on the Phanerozoic record given that well-constrained paleogeographic models are available for the past $\sim$500~Myr.
+We have focused this analysis on the Phanerozoic record given that well-constrained paleogeographic models are available for the past $\sim$520~Myr.
 Furthermore, the approach of seeking to evaluate correlation between LIP area and glaciation is complicated for Neoproterozoic Snowball Earth events, since ice-albedo runaway leads to persistent global glaciation absent of continued forcing through normal carbon cycle processes until sufficient pCO$_2$ builds up in the absence of silicate weathering to drive deglaciation.
 Nevertheless, the hypothesis of tropical LIP area associated with the ca.
-720~Ma Franklin LIP increasing global weatherability and contributing to the onset of the Sturtian Snowball Earth Glaciation is a major motivating driver behind conducting this analysis.
+720~Ma Franklin LIP increasing global weatherability and contributing to the onset of the Sturtian Snowball Earth is a major motivating driver behind conducting this analysis.
 How does the tropical LIP area associated with the Franklin LIP compare to that observed in the Phanerozoic?
 Using the paleomagnetic pole of \citet{Denyszyn2009a}, we reconstruct the paleolatitude of the Franklin LIP at the time of emplacement, and find that $\sim$99.7\% (or $\sim$2.6~Mkm$^{2}$) of the LIP erupted within 15\textdegree\xspace of the equator.
-This Franklin LIP tropical area at the time of emplacement is approximately equivalent to the minor Cenozoic peak, and is significantly smaller than the other Phanerozoic peaks (Fig.
+This Franklin LIP tropical area at the time of emplacement is approximately equivalent to the Cenozoic peak, and is smaller than the other Phanerozoic peaks (Fig.
 \ref{fig:LIP_area}C).
 The ca.
-1109~Ma Umkondo is another Precambrian LIP that is constrained to have erupted in the tropics, although it is not known to be associated with any glaciation.
+1109~Ma Umkondo is another Precambrian LIP that is constrained to have erupted in the tropics, although it is not known to be associated with any glaciation (no glacial deposits are found within the contemporaneous Midcontinent Rift basin; \citealp{Swanson-Hysell2018a}).
 We reconstruct the paleolatitude of the Umkondo LIP at the time of emplacement using the paleomagnetic pole of \citet{Swanson-Hysell2015b}, and find that effectively all of the LIP (or $\sim$2.0~Mkm$^{2}$) erupted within the tropics, an area that is slightly smaller than that for the Franklin LIP (Fig.
 \ref{fig:LIP_area}C).
-Together, these results indicate that the tropical area of the Franklin LIP, when compared that of Phanerozoic as well as other Precambrian LIPs, did not have a singularly large are in the tropics.
+Together, these results indicate that the tropical area of the Franklin LIP, when compared to that of Phanerozoic as well as other Precambrian LIPs, did not have a singularly large area in the tropics.
 Given that similar peaks in tropical LIP area are not associated with the onset of glacial periods, additional processes must have been at play in drawing down atmospheric CO$_2$ in the lead-up to the Sturtian Snowball Earth.
-One such process could have been an increase in global weatherability associated with the accretion of arc terranes within terranes of the present-day Arabian-Nubian shield.
-An increase in global weatherability associated with the emplacement of the Franklin LIP in the tropics could have been a contributing factor on top of other such processes.
+One such process could have been an increase in global weatherability associated with arc accretion within terranes of the present-day Arabian-Nubian shield.
+An increase in global weatherability associated with the emplacement of the Franklin LIP in the tropics could have been a contributing factor in addition to other such processes.
 The temporal overlap between Franklin LIP eruptions \citep{Denyszyn2009a} and the initiation of Sturtian glaciation \citep{Macdonald2010a, MacLennan2018a} remains compelling.
 This overlap continues to suggest that other aspects of LIP emplacement, such as the injection of sulfur aerosols in the stratosphere \citep{Macdonald2017a}, could have played a role in the initiation of low-latitude glaciation.
-injection is also not uniquely large, especially when considering the severity of the glaciation that followed it.
-This conclusion is consistent with the results of our Phanerozoic analysis, which suggests that LIP area and its associated impact on global weatherability is not the fundamental control on Earth's long-term climate state.
+The latitude of such aerosol injections is important for their temporary effect on albedo which is maximized at low latitude, and their presence at high concentrations is pre-conditioned on eruption through sedimentary basins hosting evaporite deposits \citep{Macdonald2017a}.
+Overall, the conclusion that tropical LIP area associated with the Franklin is not uniquely high and therefore not a sole driver of Snowball Earth onset is consistent with the results of the Phanerozoic analysis.
+LIP area and its associated impact on global weatherability is not the fundamental control on Earth's long-term climate state.
 
@@
 Richard Ernst provided GIS compilations of LIP extent that were essential to the analysis.
+Discussions with Yves Godd\'eris were valuable in contributing to aspects of the interpretation.
+Park was supported by NSF Grant 1547434 awarded to Swanson-Hysell.
 
@@
 \scriptsize
-\caption{Phanerozoic (and the Franklin) large igneous provinces.}
+\caption{Phanerozoic large igneous provinces (and the Franklin).}
 \resizebox{\textwidth}{!}{
@@
 \toprule
-\textbf{name} & \textbf{age} & \textbf{age ref.} & \textbf{original} & \textbf{present} & \textbf{present area ref.} & \textbf{present/} & \textbf{half-life*} & \textbf{buried?} \\
+\textbf{name} & \textbf{age} & \textbf{age ref.} & \textbf{original} & \textbf{present} & \textbf{present area ref.} & \textbf{present/} & \textbf{half-life}$^{1}$ & \textbf{buried?} \\
  & \textbf{(Ma)} & & \textbf{area} & \textbf{area} & & \textbf{original} & \textbf{(Myr)} & \\
@@
 NW Australia Margin & 160 & \cite{Pirajno2012a} & 0.62 & 0.00 & \cite{Coffin2006a} & 0.00 & 0.0 & yes \\
-Karoo & 183 & \cite{Burgess2015a} & 3.21 & 0.15 & de Kock compilation & 0.05 & 41.3 & no \\
+Karoo & 183 & \cite{Burgess2015a} & 3.21 & 0.15 & de Kock compilation$^{2}$ & 0.05 & 41.3 & no \\
 Ferrar & 183 & \cite{Burgess2015a} & 0.18 & - & - & - & - & no \\
-CAMP & 201 & \cite{Blackburn2013a} & 11.46 & 0.23 & Marzoli and Parisio compilation & 0.02 & 35.7 & partial \\
+CAMP & 201 & \cite{Blackburn2013a} & 11.46 & 0.23 & Marzoli and Parisio compilation$^{3}$ & 0.02 & 35.7 & partial \\
 Siberia & 252 & \cite{Burgess2015b} & 3.46 & 0.47 & \cite{Coffin2006a} & 0.14 & 87.5 & no \\
@@
 \begin{tablenotes}[para,flushleft]
-\textbf{*}assuming exponential decay with form $N(t) = 2^{t/t_{1/2}}$
+$^{1}$assuming exponential decay with form $N(t) = 2^{t/t_{1/2}}$.
+
+$^{2}$from ArcGIS compilation produced under LIPs Project contract to M. de Kock.
+
+$^{3}$from ArcGIS CAMP compilation of A. Marzoli and L. Parisio produced under contract to LIPs Reconstruction Project.
 \end{tablenotes}
@@
 	\includegraphics[width=\textwidth]{Manuscript/Figures/LIP_Map.pdf}
-	\caption{Map of current extent of volcanic lithologies associated with LIPs that erupted between 520~Ma and the present as well as the estimates of the initial LIP area used in the area analysis (modified slightly from \citealp{Ernst2017a}).}
+	\caption{Map of current extent of volcanic lithologies associated with LIPs that erupted between 520~Ma and the present, as well as the estimates of the initial LIP area used in the area analysis (modified slightly from \citealp{Ernst2017a}).}
 	\label{fig:LIP_map}
@@
 The ratio of estimates of the present-day area to that of the original area are shown for 19 basaltic LIPs.
-An exponential fit is made to the 13 basaltic LIPs that have not been buried after emplacement, which yields a half-life of $\sim$36~Myr.}
+An exponential fit is made to the 13 basaltic LIPs that have not been buried after emplacement (Table \ref{tab:LIPs}), which yields a half-life of $\sim$36~Myr.
+RMSE = root mean square error.}
 	\label{fig:LIP_preservation}
@@
 \textbf{C)} Tropical LIP area through time for the different post-emplacement models.
-The arrows indicate tropical LIP area for the ca.
+The arrows to the right indicate reconstructed tropical LIP area at the time of emplacement for the ca.
 720~Ma Franklin LIP and the ca.
@@
 	\includegraphics[width=0.85\textwidth]{Manuscript/Figures/overlap_correlation_cropped.pdf}
-	\caption{The overlap and correlation values for various LIP post-emplacement scenarios are shown with circles (d1 : LIP decay with half-life of 36~Myr; d2: LIP decay with half-life of~120 Myr; b1: 50\% burial of LIPs associated with rifting; b2: 100\% burial of LIPs associated with rifting).
+	\caption{The \% overlap and Pearson correlation coefficients for various LIP post-emplacement scenarios are shown with circles.
 These values are compared to histograms that show the range of values that arise when comparing the LIP area record to glacial intervals that have been shifted randomly in time 100,000 times.}

### 2018-11-15 Manuscript/Manuscript.tex
Updates from ShareLaTeX

@@
 \begin{flushleft}
-{\Large \textbf{Evaluating the relationship between large igneous province area and Earth's long-term climate state}}
+{\Large \textbf{Evaluating the relationship between the area and latitude of large igneous provinces and Earth's long-term climate state}}
 
@@
 Nicholas L. Swanson-Hysell\textsuperscript{1},
-Francis A. Macdonald\textsuperscript{2},
-Lorraine E. Lisiecki\textsuperscript{2}
+Lorraine E. Lisiecki\textsuperscript{2},
+Francis A. Macdonald\textsuperscript{2}
 
@@
 On a planet with high weatherability, the CO$_2$ input from volcanism can be removed via silicate weathering at a lower atmospheric CO$_2$ concentration than on a less weatherable planet.
-The relatively high concentrations of Ca and Mg (that ultimately sequester carbon through precipitation as carbonate) and the high weathering rates of mafic lithologies make it so that basaltic regions consume more CO$_2$ than regions where bedrock composition is closer to bulk continental crust \citep{Dessert2003a}.
+Basaltic regions consume more CO$_2$ than regions where bedrock composition is closer to bulk continental crust because of the relatively high concentrations of Ca and Mg (that ultimately sequester carbon through precipitation as carbonate) and the high weathering rates of mafic lithologies \citep{Dessert2003a}.
 Data from basaltic watersheds show that chemical weathering rates are greatest in regions with high runoff, and as a result CO$_2$ consumption in basaltic regions is most pronounced in the tropical rain belt \citep{Dessert2003a, Hartmann2009a, Hartmann2014a}.
@@
 In particular, the emplacement of LIPs in the tropics has been hypothesized to be associated with specific episodes of climatic cooling on Earth.
-In the Neoproterozoic era, the emplacement of the Franklin LIP in the tropics ca.
-720~Ma, in concert with elevated runoff rates associated with continental break-up, has been implicated as a major contributor to the cooling that initiated the Sturtian `Snowball Earth' \citep{Donnadieu2004b, Cox2016a}.
+In the Neoproterozoic Era, the emplacement of the ca.
+720~Ma Franklin LIP in the tropics, in concert with elevated runoff rates associated with continental break-up, has been implicated as a major contributor to the cooling that initiated the Sturtian `Snowball Earth' \citep{Donnadieu2004b, Macdonald2010a, Cox2016a}.
 In the Cenozoic, the movement of the Deccan LIP into the tropical rain belt, in concert with the low-latitude emplacement of the Ethiopian Traps, has been implicated in drawing down CO$_2$ levels in the lead-up to Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
 
-This chapter seeks to address two interconnected questions: 1) how uniquely large are the peaks in low-latitude LIP area that have been proposed to be associated with climatic cooling?
+This chapter seeks to address two interconnected questions: 1) how unique are the peaks in low-latitude LIP area that have been proposed to be associated with climatic cooling?
 and 2) how strong is the overall relationship between tropical LIP area and glaciation?
@@
 The compilation of \citet{Ernst2017a} seeks to reconstruct the original surface extent of LIPs with the caveat that there can be significant uncertainty with doing so, particularly for older more deeply eroded LIPs.
-These polygons encapsulate all of the rocks associated with a given LIP, including dikes and sills (Fig.
+These polygons encapsulate all of the rocks associated with a given LIP, including dikes, sills, and layered intrusions in addition to subaerial volcanics (Fig.
 \ref{fig:LIP_map}).
@@
 However, this approach likely provides the best estimates available of original extent for ancient LIPs.
-The extents of presently-exposed volcanics associated with LIPs were taken from a number of resources including the PLATES compilation \citep{Coffin2006a} and more recent compilation efforts associated with the Large Igneous Provinces Commission (Table \ref{tab:LIPs}).
+The extents of presently-exposed volcanics associated with LIPs that were used for present-day area estimates, and the construction of the original extent polygons by \citet{Ernst2017a}, were taken from a number of resources including the PLATES compilation \citep{Coffin2006a} and more recent compilation efforts associated with the LIPs Reconstruction Project (\citealp{Ernst2013a}; Table \ref{tab:LIPs}).
 
@@
 
-To evaluate the relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic (modified from \citealp{Royer2004a}; Fig.
+To evaluate the relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic (Macdonald et al., in review; Fig.
 \ref{fig:LIP_area}E).
@@
 
-Another approach would be to compare LIP area records to proxy compilations of pCO$_2$ (as done by \citealp{Johansson2018a}).
+Another approach would be to compare LIP area records to proxy compilations of \textit{p}CO$_2$ (as done by \citealp{Johansson2018a}).
 However, such proxy compilations are problematic as they are poorly calibrated in deep time, prone to alteration, and reliant on a myriad of assumptions.
@@
 regolith development itself) in a number of the provinces; 2) the current extent of LIP exposure may be incomplete and an underestimate for some of the provinces; 3) the initial LIP extents are typically poorly constrained and are likely overestimates which could be resulting in inflated interpreted decay rates; and 4) the relationship between LIP area and volume is poorly constrained.
-Future efforts to develop better-constrained estimates of original LIP extent will improve analyses such as that in this contribution.
+Future efforts that improve the LIP database, such as developing better-constrained estimates of original LIP extent, constraining burial and uplift histories, and refining the timing of eruptions associated with LIPs, will improve analyses such as that in this contribution.
 
@@
 
-Richard Ernst provided GIS compilations of LIP extent that were essential to the analysis.
+Richard Ernst provided GIS compilations of LIP extent and present-day exposure that were essential to the analysis.
 Discussions with Yves Godd\'eris were valuable in contributing to aspects of the interpretation.
@@
 Bunbury & 132 & \cite{Zhu2009a} & 0.03 & 0.00 & \cite{Thorne2014a} & 0.05 & 30.9 & no \\
-Parana-Etendeka & 135 & \cite{Florisbal2014a} & 3.12 & 0.40 & \cite{Coffin2006a} & 0.13 & 45.7 & partial \\
+Parana-Etendeka & 135 & \cite{Florisbal2014a, Almeida2018} & 3.12 & 0.40 & \cite{Coffin2006a} & 0.13 & 45.7 & partial \\
 Trap & 140 & \cite{Ernst2001a} & 0.03 & 0.00 & \cite{Ernst2001a} & 0.00 & 0.0 & no \\
@@
 
-$^{2}$from ArcGIS compilation produced under LIPs Project contract to M. de Kock.
+$^{2}$from ArcGIS compilation produced by M. de Kock for the LIPs Reconstruction Project \citep{Ernst2013a}.
 
-$^{3}$from ArcGIS CAMP compilation of A. Marzoli and L. Parisio produced under contract to LIPs Reconstruction Project.
+$^{3}$from ArcGIS compilation produced by A. Marzoli and L. Parisio for the LIPs Reconstruction Project \citep{Ernst2013a}.
 \end{tablenotes}

### 2018-11-15 Manuscript/Manuscript.tex

@@
 
-To evaluate the relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic (Macdonald et al., in review; Fig.
+To evaluate the relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic (\citealp{Macdonald2018a}; Fig.
 \ref{fig:LIP_area}E).
@@
 Bunbury & 132 & \cite{Zhu2009a} & 0.03 & 0.00 & \cite{Thorne2014a} & 0.05 & 30.9 & no \\
-Parana-Etendeka & 135 & \cite{Florisbal2014a, Almeida2018} & 3.12 & 0.40 & \cite{Coffin2006a} & 0.13 & 45.7 & partial \\
+Parana-Etendeka & 135 & \cite{Florisbal2014a, Almeida2018a} & 3.12 & 0.40 & \cite{Coffin2006a} & 0.13 & 45.7 & partial \\
 Trap & 140 & \cite{Ernst2001a} & 0.03 & 0.00 & \cite{Ernst2001a} & 0.00 & 0.0 & no \\

### 2018-11-15 Manuscript/Manuscript.tex

@@
 One of the hypothesized effects of the emplacement of large igneous provinces (LIPs) is planetary cooling on million-year timescales associated with enhanced silicate weathering.
-A systematic analysis of LIP area through the Phanerozoic reveals that there is no significant correlation between total LIP area, nor LIP area in the tropics, and the extent of continental ice sheets.
+A systematic analysis of LIP area and latitude through the Phanerozoic reveals that there is no significant correlation between total LIP area, nor LIP area in the tropics, and the extent of continental ice sheets.
 The largest peaks in tropical LIP area are at times of non-glacial climate.
@@
 
-This chapter seeks to address two interconnected questions: 1) how unique are the peaks in low-latitude LIP area that have been proposed to be associated with climatic cooling?
-and 2) how strong is the overall relationship between tropical LIP area and glaciation?
+This chapter seeks to address two interconnected questions: 1) how unique are the peaks in low-latitude LIP area that have been proposed to be associated with climatic cooling?; and 2) how strong is the overall relationship between tropical LIP area and glaciation?
 
@@
 
-Outlines of LIPs through the Phanerozoic (Fig.
-\ref{fig:LIP_map}) were taken from the compilation of \citet{Ernst2017a}, and emplacement ages were taken from the literature (Fig.
-\ref{fig:LIP_map}; Table \ref{tab:LIPs}).
+Outlines of the original surface extent of LIPs through the Phanerozoic (Fig.
+\ref{fig:LIP_map}) were slightly modified from the compilation of \citet{Ernst2017a}, and emplacement ages were taken from the literature (Table \ref{tab:LIPs}).
 The compilation of \citet{Ernst2017a} seeks to reconstruct the original surface extent of LIPs with the caveat that there can be significant uncertainty with doing so, particularly for older more deeply eroded LIPs.
@@
 The best-fit exponential function results in a LIP area half-life of 29~Myr.
-However, since we explicitly account for LIP burial separately in our models (see below), we prefer to exclude the 6 of 19 LIPs which have been partially or completely buried in our estimation of a representative LIP area half-life.
+However, since we explicitly account for LIP burial separately in our models (see below), we prefer to exclude from our estimation of a representative LIP area half-life the 6 of 19 LIPs which have been partially or completely buried.
 This latter approach yields a slightly higher best-fit half-life of 36~Myr (Fig.
@@
 There are numerous examples in the record where there is partial or complete burial of a LIP associated with rifting and thermal subsidence (Table \ref{tab:LIPs}).
-For example, the Afar LIP is both associated with the Ethiopian Traps which form plateau flood basalts as well as successful rifting in the region of the Red Sea that has resulted in burial (Fig \ref{fig:LIP_map}).
+For example, the Afar LIP is both associated with the Ethiopian Traps which form plateau flood basalts as well as successful rifting in the region of the Red Sea that has resulted in burial (Fig.
+\ref{fig:LIP_map}).
 To account for the rapid decrease in exposed surface area that would result from burial by sediments in a rift basin, we impose two different burial scenarios for LIPs that are co-located with rifting.
@@
 \ref{fig:LIP_area}B and C).
-We use 15\textdegree~S to 15\textdegree~N as the definition of the tropical rain belt as those latitudes approximately correspond with precipitation greater than 1.0~m/yr in modern climatalogical data \citep{Kalnay1996a}.
+We use 15\textdegree~S to 15\textdegree~N as the definition of the tropical rain belt, as those latitudes approximately correspond with precipitation greater than 1.0~m/yr in modern climatalogical data \citep{Kalnay1996a}.
 The consistent paleolatitudes of evaporite deposits in the subtropics demonstrate the stability of the large-scale atmospheric circulation that gives rise to intense precipitation in the tropical rain belt \citep{Evans2006a}.
@@
 To assess the statistical significance of the correlation (or lack thereof) implied by the Pearson correlation coefficients, we simulated the four glacial episodes (Fig.
-\ref{fig:LIP_area}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and \% overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these 100,000 simulations (Fig.
+\ref{fig:LIP_area}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and percent overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these 100,000 simulations (Fig.
 \ref{fig:LIP_correlation}).
@@
 We have focused this analysis on the Phanerozoic record given that well-constrained paleogeographic models are available for the past $\sim$520~Myr.
-Furthermore, the approach of seeking to evaluate correlation between LIP area and glaciation is complicated for Neoproterozoic Snowball Earth events, since ice-albedo runaway leads to persistent global glaciation absent of continued forcing through normal carbon cycle processes until sufficient pCO$_2$ builds up in the absence of silicate weathering to drive deglaciation.
+Furthermore, the approach of seeking to evaluate correlation between LIP area and glaciation is complicated for Neoproterozoic Snowball Earth events, since ice-albedo runaway leads to persistent global glaciation absent of continued forcing through normal carbon cycle processes until sufficient CO$_2$ builds up in the absence of silicate weathering to drive deglaciation.
 Nevertheless, the hypothesis of tropical LIP area associated with the ca.
@@
 Given that similar peaks in tropical LIP area are not associated with the onset of glacial periods, additional processes must have been at play in drawing down atmospheric CO$_2$ in the lead-up to the Sturtian Snowball Earth.
-One such process could have been an increase in global weatherability associated with arc accretion within terranes of the present-day Arabian-Nubian shield.
+One such process could have been an increase in global weatherability associated with arc accretion within terranes of the present-day Arabian-Nubian Shield \citep{Park2018a}.
 An increase in global weatherability associated with the emplacement of the Franklin LIP in the tropics could have been a contributing factor in addition to other such processes.
+
 The temporal overlap between Franklin LIP eruptions \citep{Denyszyn2009a} and the initiation of Sturtian glaciation \citep{Macdonald2010a, MacLennan2018a} remains compelling.
@@
 The latitude of such aerosol injections is important for their temporary effect on albedo which is maximized at low latitude, and their presence at high concentrations is pre-conditioned on eruption through sedimentary basins hosting evaporite deposits \citep{Macdonald2017a}.
+
 Overall, the conclusion that tropical LIP area associated with the Franklin is not uniquely high and therefore not a sole driver of Snowball Earth onset is consistent with the results of the Phanerozoic analysis.
@@
 Columbia River & 16 & \cite{Kasbohm2018a} & 0.68 & 0.38 & \cite{Buchan2004a} & 0.56 & 19.2 & no \\
+ & & & & & & & & \\
 Afar & 30 & \cite{Courtillot2003a} & 2.05 & 0.63 & \cite{Coffin2006a} & 0.31 & 17.7 & partial \\
+ & & & & & & & & \\
 NAIP & 62 & \cite{Larsen2015a} & 1.07 & 0.29 & \cite{Buchan2004a, Coffin2006a} & 0.27 & 33.0 & partial \\
+ & & & & & & & & \\
 Deccan & 66 & \cite{Schoene2014a} & 0.83 & 0.56 & \cite{Coffin2006a} & 0.68 & 116.6 & no \\
+ & & & & & & & & \\
 Seychelles & 66 & \cite{Schoene2014a} & 0.46 & 0.00 & \cite{Coffin2006a} & 0.00 & 0.0 & yes \\
+ & & & & & & & & \\
 Madagascar & 90 & \cite{Cucciniello2010a} & 0.63 & 0.03 & \cite{Coffin2006a} & 0.05 & 20.8 & no \\
+ & & & & & & & & \\
 Caribbean-Colombian & 94 & \cite{Loewen2013a} & 0.71 & 0.13 & \cite{Coffin2006a} & 0.18 & 37.6 & no \\
+ & & & & & & & & \\
 HALIP & 95 & \cite{Kingsbury2018a} & 3.60 & 0.15 & \cite{Moosdorf2015a} & 0.04 & 20.7 & no \\
+ & & & & & & & & \\
 EQUAMP & 131 & \cite{Hollanda2016a} & 0.66 & 0.01 & \cite{Hollanda2016a} & 0.01 & 20.5 & no \\
+ & & & & & & & & \\
 Comei & 132 & \cite{Zhu2009a} & 0.11 & - & - & - & - & no \\
+ & & & & & & & & \\
 Bunbury & 132 & \cite{Zhu2009a} & 0.03 & 0.00 & \cite{Thorne2014a} & 0.05 & 30.9 & no \\
+ & & & & & & & & \\
 Parana-Etendeka & 135 & \cite{Florisbal2014a, Almeida2018a} & 3.12 & 0.40 & \cite{Coffin2006a} & 0.13 & 45.7 & partial \\
+ & & & & & & & & \\
 Trap & 140 & \cite{Ernst2001a} & 0.03 & 0.00 & \cite{Ernst2001a} & 0.00 & 0.0 & no \\
+ & & & & & & & & \\
 NW Australia Margin & 160 & \cite{Pirajno2012a} & 0.62 & 0.00 & \cite{Coffin2006a} & 0.00 & 0.0 & yes \\
+ & & & & & & & & \\
 Karoo & 183 & \cite{Burgess2015a} & 3.21 & 0.15 & de Kock compilation$^{2}$ & 0.05 & 41.3 & no \\
+ & & & & & & & & \\
 Ferrar & 183 & \cite{Burgess2015a} & 0.18 & - & - & - & - & no \\
+ & & & & & & & & \\
 CAMP & 201 & \cite{Blackburn2013a} & 11.46 & 0.23 & Marzoli and Parisio compilation$^{3}$ & 0.02 & 35.7 & partial \\
+ & & & & & & & & \\
 Siberia & 252 & \cite{Burgess2015b} & 3.46 & 0.47 & \cite{Coffin2006a} & 0.14 & 87.5 & no \\
+ & & & & & & & & \\
 Emeishan & 259 & \cite{Zhou2002a} & 0.71 & 0.06 & \cite{Coffin2006a} & 0.09 & 72.9 & no \\
+ & & & & & & & & \\
 Panjal-Qiangtang & 283 & \cite{Zhai2013a} & 0.11 & - & - & - & - & no \\
+ & & & & & & & & \\
 Tarim & 290 & \cite{Xu2014a} & 0.35 & - & - & - & - & no \\
+ & & & & & & & & \\
 Magdalen & 360 & \cite{Brendan_Murphy1999a} & 0.42 & - & - & - & - & no \\
+ & & & & & & & & \\
 Vilyui & 374 & \cite{Ricci2013a} & 1.14 & - & - & - & - & no \\
+ & & & & & & & & \\
 Kola-Dnieper & 380 & \cite{Arzamastsev2014a} & 5.90 & - & - & - & - & no \\
+ & & & & & & & & \\
 Suordakh & 450 & \cite{Khudoley2013a} & 0.02 & - & - & - & - & no \\
+ & & & & & & & & \\
 Kalkarindji & 511 & \cite{Jourdan2014a} & 3.54 & 0.17 & \cite{Thorne2014a} & 0.05 & 116.3 & no \\
+ & & & & & & & & \\
 Franklin & 720 & \cite{Denyszyn2009a} & 2.62 & 0.04 & \cite{Buchan2004a} & 0.02 & 121.8 & no \\
@@
 \begin{center}
-	\includegraphics[width=0.9\textwidth]{Manuscript/Figures/LIP_Areas.pdf}
-	\caption{\textbf{A)} Paleolatitude of emplacement (centroid) and the initial area estimate of each LIP in the compilation.
+	\includegraphics[width=0.8\textwidth]{Manuscript/Figures/LIP_Areas.pdf}
+	\caption{\textbf{A)} LIPs included in this analysis.
+The size of each circle reflects the initial area estimate of each LIP.
+The + indicates the timing and absolute paleolatitude of the centroid of each LIP at the time of emplacement.
 Car.-Col.
@@
 \textbf{D)} Heat map for the preferred post-emplacement model showing the latitudinal distribution of LIP area.
-\textbf{E)} Latitudinal extent of land ice away from the poles through time.}
+\textbf{E)} Latitudinal extent of land ice away from the poles through time \citep{Macdonald2018a}.}
 	\label{fig:LIP_area}

### 2018-11-15 Manuscript/Manuscript.tex
completion of implementation of round 1 edits

@@
 To assess the statistical significance of the correlation (or lack thereof) implied by the Pearson correlation coefficients, we simulated the four glacial episodes (Fig.
-\ref{fig:LIP_area}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and percent overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these 100,000 simulations (Fig.
+\ref{fig:LIP_area}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and \% overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these 100,000 simulations (Fig.
 \ref{fig:LIP_correlation}).
@@
 An increase in global weatherability associated with the emplacement of the Franklin LIP in the tropics could have been a contributing factor in addition to other such processes.
-
 The temporal overlap between Franklin LIP eruptions \citep{Denyszyn2009a} and the initiation of Sturtian glaciation \citep{Macdonald2010a, MacLennan2018a} remains compelling.
@@
 The latitude of such aerosol injections is important for their temporary effect on albedo which is maximized at low latitude, and their presence at high concentrations is pre-conditioned on eruption through sedimentary basins hosting evaporite deposits \citep{Macdonald2017a}.
-
 Overall, the conclusion that tropical LIP area associated with the Franklin is not uniquely high and therefore not a sole driver of Snowball Earth onset is consistent with the results of the Phanerozoic analysis.

### 2018-11-20 Manuscript/Manuscript.tex

@@
 One of the hypothesized effects of the emplacement of large igneous provinces (LIPs) is planetary cooling on million-year timescales associated with enhanced silicate weathering.
-A systematic analysis of LIP area and latitude through the Phanerozoic reveals that there is no significant correlation between total LIP area, nor LIP area in the tropics, and the extent of continental ice sheets.
+This study combines outlines of the original surface extent of LIPs through the Phanerozoic, a paleogeographic model, and LIP erosion models to estimate LIP area in all latitudinal bands through the Phanerozoic.
+This analysis reveals no significant correlation between total LIP area, nor LIP area in the tropics, and the extent of continental ice sheets.
 The largest peaks in tropical LIP area are at times of non-glacial climate.
@@
 Basaltic regions consume more CO$_2$ than regions where bedrock composition is closer to bulk continental crust because of the relatively high concentrations of Ca and Mg (that ultimately sequester carbon through precipitation as carbonate) and the high weathering rates of mafic lithologies \citep{Dessert2003a}.
-Data from basaltic watersheds show that chemical weathering rates are greatest in regions with high runoff, and as a result CO$_2$ consumption in basaltic regions is most pronounced in the tropical rain belt \citep{Dessert2003a, Hartmann2009a, Hartmann2014a}.
+Data from basaltic watersheds show that chemical weathering rates are greatest in regions with high runoff.
+As a result, CO$_2$ consumption in basaltic regions is most pronounced in the tropical rain belt \citep{Dessert2003a, Hartmann2009a, Hartmann2014a}.
 
@@
 In the Cenozoic, the movement of the Deccan LIP into the tropical rain belt, in concert with the low-latitude emplacement of the Ethiopian Traps, has been implicated in drawing down CO$_2$ levels in the lead-up to Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
+More recently, \citet{Johansson2018a} used a paleogeographic model to suggest that tropical LIP area correlates with Phanerozoic climate change.
 
@@
 
+This study combines outlines of the original surface extent of LIPs through the Phanerozoic, a paleogeographic model, and LIP erosion models to estimate LIP area in all latitudinal bands through the Phanerozoic.
+We then compare these LIP area time series to the latitudinal extent of continental ice sheets - a proxy for Earth's long term climate state.
+
 Outlines of the original surface extent of LIPs through the Phanerozoic (Fig.
@@
 The compilation of \citet{Ernst2017a} seeks to reconstruct the original surface extent of LIPs with the caveat that there can be significant uncertainty with doing so, particularly for older more deeply eroded LIPs.
-These polygons encapsulate all of the rocks associated with a given LIP, including dikes, sills, and layered intrusions in addition to subaerial volcanics (Fig.
+These polygons encapsulate all of the rocks associated with a given LIP, including dikes, sills, and layered intrusions, in addition to subaerial volcanics (Fig.
 \ref{fig:LIP_map}).
@@
 \ref{fig:LIP_preservation}).
-While there are significant uncertainties associated with the area estimates, this compilation suggests that an exponential decay function is an appropriate first-order model for the progressive reduction in LIP area.
+While there are significant uncertainties associated with the area estimates, this compilation suggests that an exponential decay function is an appropriate first-order model for the progressive reduction in LIP area (Fig.
+\ref{fig:LIP_preservation}).
 The best-fit exponential function results in a LIP area half-life of 29~Myr.
@@
 \ref{fig:LIP_preservation}).
-Although the root mean square error of the fit is low (0.14), there is variability in the half-lives for LIPs if they are fit individually with an exponential decay function, such that they range from shorter half-lives of $\sim$20~Myr up to $\sim$120~Myr for the Kalkarindji and Deccan Traps (Table \ref{tab:LIPs}).
+Although the exponential fit to the 13 unburied LIPs is good (i.e.
+it yields a low root mean square error of 0.14), if each LIP is fit individually with an exponential decay function, there is variability in the estimated half-lives such that they range from shorter half-lives of $\sim$20~Myr up to $\sim$120~Myr for the Kalkarindji and Deccan Traps (Table \ref{tab:LIPs}).
 In the analysis of LIP area through time, we implement decay scenarios informed by this analysis.
@@
 The LIP reconstructions used in this study also include pre-Phanerozoic LIPs.
-However, given the imposed exponential decay since emplacement, the inclusion of these LIPs does not significantly impact the calculated LIP areas throughout the Phanerozoic, which is the focus of this analysis.
+However, given the imposed exponential decay since emplacement, the inclusion of these LIPs does not significantly influence the calculated LIP areas throughout the Phanerozoic, which is the focus of this analysis.
 
@@
 Our LIP reconstruction differs from that of \citet{Johansson2018a} in several respects.
-In contrast to the decay and decay+burial scenarios implemented on estimates of original LIP extent in this study, \citet{Johansson2018a} uses a static extent for each LIP throughout the reconstruction with some of the polygons corresponding to the present-day surface extent and some representing the original extent that includes currently buried portions of the LIP (e.g.
-the Keweenawan Midcontinent Rift for which the implemented extent is from geophysical data that largely correponds to buried subsurface exposures).
+In contrast to the decay and decay+burial scenarios implemented on estimates of original LIP extent in this study, \citet{Johansson2018a} uses a static extent for each LIP throughout the reconstruction.
+In their analysis, some of the polygons correspond to the present-day surface extent and some represent the original extent that includes currently buried portions of the LIP (e.g.
+the Keweenawan Midcontinent Rift for which the implemented extent is from geophysical data that largely corresponds to buried subsurface exposures).
 
@@
 
-In the `t$_{1/2}$ = 36~Myr' scenario, we observe three primary peaks and one minor peak in the calculated LIP area within the tropics (Fig.
+In the `t$_{1/2}$ = 36~Myr' scenario, we observe three major peaks and one minor peak in the calculated LIP area within the tropics (Fig.
 \ref{fig:LIP_area}C).
-The first primary peak ca.
-510~Ma is associated with the emplacement of the Kalkarindji LIP, the second primary peak ca.
-380~Ma is associated with the emplacement of the Kola-Dnieper LIP, the third primary peak ca.
+The first major peak ca.
+510~Ma is associated with the emplacement of the Kalkarindji LIP, the second major peak ca.
+380~Ma is associated with the emplacement of the Kola-Dnieper LIP, the third major peak ca.
 200~Ma is associated with the emplacement of the CAMP, and the minor Cenozoic peak is associated with both the ca.
@@
 All other scenarios yield a near zero or weak negative correlation coefficient.
-Qualitatively, the weak positive correlation of the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario relative to the other scenarios can be primarily attributed to the complete removal of the CAMP, which was emplaced during an extended interval of ice-free conditions, as well as the effect of the longer decay half-life extending the duration of the earlier two peaks, such that they overlap more with the Late Ordovician and Permo-Carboniferous glacial intervals.
+The weak positive correlation of the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario relative to the other scenarios can be primarily attributed to the complete removal of the CAMP, which was emplaced during an extended interval of ice-free conditions, as well as the effect of the longer decay half-life extending the duration of the earlier two peaks, such that they overlap more with the Late Ordovician and Permo-Carboniferous glacial intervals.
 
@@
 
-The results from this analysis indicate that when the entire LIP database is considered, there is no significant relationship between total LIP area or LIP area in the tropics and the extent of continental ice sheets (Fig.
-\ref{fig:LIP_correlation}).
-While this result need not imply that there is no increase in global weatherability from the emplacement of LIPs, it does suggest that LIP area is not the fundamental control on Earth's long-term climate state.
-
-In the original models that proposed the `Fire and Ice' hypothesis as an explanation for the onset of the Sturtian `Snowball Earth' glaciation, chemical weathering was modeled as an Arrhenius relationship where it was solely a function of temperature and run-off \citep{Donnadieu2004a}.
+In the original models that proposed the `Fire and Ice' hypothesis as an explanation for the onset of the Sturtian `Snowball Earth' glaciation, chemical weathering was modeled as a function of temperature and run-off only \citep{Donnadieu2004a}.
 However, such an approach neglects the effects of soil shielding and regolith development in low-relief regions.
@@
 We have focused this analysis on the Phanerozoic record given that well-constrained paleogeographic models are available for the past $\sim$520~Myr.
-Furthermore, the approach of seeking to evaluate correlation between LIP area and glaciation is complicated for Neoproterozoic Snowball Earth events, since ice-albedo runaway leads to persistent global glaciation absent of continued forcing through normal carbon cycle processes until sufficient CO$_2$ builds up in the absence of silicate weathering to drive deglaciation.
+The approach of seeking to evaluate correlation between LIP area and glaciation is further complicated for Neoproterozoic Snowball Earth events because ice-albedo runaway leads to persistent global glaciation on timescales of tens of millions of years absent of continued forcing through normal carbon cycle processes until sufficient CO$_2$ to drive deglaciation builds up in the absence of silicate weathering \citep{Hoffman2017a}.
+Moreover, cooling past the critical threshold for rapid global glaciation may have occurred on a sub-million year timescale that is not recorded in the geological carbon cycle (e.g.
+\citealp{Macdonald2017a}).
 Nevertheless, the hypothesis of tropical LIP area associated with the ca.
 720~Ma Franklin LIP increasing global weatherability and contributing to the onset of the Sturtian Snowball Earth is a major motivating driver behind conducting this analysis.
+
 How does the tropical LIP area associated with the Franklin LIP compare to that observed in the Phanerozoic?
@@
 \ref{fig:LIP_area}C).
+
 Together, these results indicate that the tropical area of the Franklin LIP, when compared to that of Phanerozoic as well as other Precambrian LIPs, did not have a singularly large area in the tropics.
-Given that similar peaks in tropical LIP area are not associated with the onset of glacial periods, additional processes must have been at play in drawing down atmospheric CO$_2$ in the lead-up to the Sturtian Snowball Earth.
-One such process could have been an increase in global weatherability associated with arc accretion within terranes of the present-day Arabian-Nubian Shield \citep{Park2018a}.
+Given that similar peaks in tropical LIP area are not associated with the onset of glacial periods, additional processes beyond an increase in weatherability due to LIP area in the tropics must have been at play in the initiation of the Sturtian Snowball Earth.
+One such process could have been an increase in global weatherability associated with the collision and accretion of arc terranes within the present-day Arabian-Nubian Shield \citep{Park2018a}.
+A low-latitude continent configuration and abundant arc-continent collisions in the tropics may have led to a cool background climate and the emplacement of the Franklin LIP may have pushed the climate system over the edge.
+However, tropical LIP area associated with the Franklin was not uniquely high, and therefore an associated increase in global weatherability was not the sole driver of Snowball Earth onset, consistent with the results of the Phanerozoic analysis.
+
 An increase in global weatherability associated with the emplacement of the Franklin LIP in the tropics could have been a contributing factor in addition to other such processes.
-The temporal overlap between Franklin LIP eruptions \citep{Denyszyn2009a} and the initiation of Sturtian glaciation \citep{Macdonald2010a, MacLennan2018a} remains compelling.
-This overlap continues to suggest that other aspects of LIP emplacement, such as the injection of sulfur aerosols in the stratosphere \citep{Macdonald2017a}, could have played a role in the initiation of low-latitude glaciation.
-The latitude of such aerosol injections is important for their temporary effect on albedo which is maximized at low latitude, and their presence at high concentrations is pre-conditioned on eruption through sedimentary basins hosting evaporite deposits \citep{Macdonald2017a}.
-Overall, the conclusion that tropical LIP area associated with the Franklin is not uniquely high and therefore not a sole driver of Snowball Earth onset is consistent with the results of the Phanerozoic analysis.
-LIP area and its associated impact on global weatherability is not the fundamental control on Earth's long-term climate state.
+The temporal overlap between Franklin LIP eruptions and the initiation of Sturtian glaciation remains compelling \citep{Macdonald2017a}.
+This overlap continues to suggest that other aspects of LIP emplacement, such as relieving P, Fe, or S limitations on biological productivity \citep{Tziperman2011a, Cox2016a}, or the injection of sulfur aerosols in the stratosphere \citep{Macdonald2017a}, could have played a role in the initiation of low-latitude glaciation.
+The latitude and background climate of such aerosol injections is important for their temporary effect on albedo which is maximized at low latitude, and their presence at high concentrations is pre-conditioned on eruption through sedimentary basins hosting evaporite deposits \citep{Macdonald2017a}.
+Sulfur aerosols associated with low-latitude LIPs may have also caused short-term cooling due to changes in albedo during the Phanerozoic, as has been suggested for the CAMP \citep{Schoene2010a} and putative LIPs associated with the Hirnantian ice advance \citep{Jones2017a}.
+However, because these cooling events did not result in a Snowball Earth, the climate would have returned to its background climate state within thousands of years \citep{Macdonald2017a}, which would not be observed in our analysis.
+
+The results from this analysis indicate that when the entire LIP database is considered, there is no significant relationship between total LIP area or LIP area in the tropics and the extent of continental ice sheets (Fig.
+\ref{fig:LIP_correlation}).
+While this result need not imply that there is no increase in global weatherability from the emplacement of LIPs, it does suggest that LIP area is not the fundamental control on Earth's long-term climate state.
 
@@
 Richard Ernst provided GIS compilations of LIP extent and present-day exposure that were essential to the analysis.
-Discussions with Yves Godd\'eris were valuable in contributing to aspects of the interpretation.
+Discussions with Yves Godd\'eris made possible through the France Berkeley Fund were valuable in contributing to aspects of the interpretation.
 Park was supported by NSF Grant 1547434 awarded to Swanson-Hysell.

### 2018-11-22 Manuscript/Manuscript.tex
incorporate Lorraine's edits, before sending a version to Ernst

@@
 One of the hypothesized effects of the emplacement of large igneous provinces (LIPs) is planetary cooling on million-year timescales associated with enhanced silicate weathering.
-This study combines outlines of the original surface extent of LIPs through the Phanerozoic, a paleogeographic model, and LIP erosion models to estimate LIP area in all latitudinal bands through the Phanerozoic.
+This study combines reconstructions of the original surface extent of LIPs through the Phanerozoic, a paleogeographic model, and LIP erosion models to estimate LIP area in all latitudinal bands through the Phanerozoic.
 This analysis reveals no significant correlation between total LIP area, nor LIP area in the tropics, and the extent of continental ice sheets.
@@
 
-Global weatherability is the sum of factors aside from climate itself, such as the latitudinal distribution of continents and mountain belts, that contribute to overall global weathering and associated CO$_2$ consumption.
+Global weatherability is the sum of factors aside from climate itself that contribute to overall global weathering and associated CO$_2$ consumption, such as the latitudinal distribution of continents and mountain belts.
 On a planet with high weatherability, the CO$_2$ input from volcanism can be removed via silicate weathering at a lower atmospheric CO$_2$ concentration than on a less weatherable planet.
@@
 
-This study combines outlines of the original surface extent of LIPs through the Phanerozoic, a paleogeographic model, and LIP erosion models to estimate LIP area in all latitudinal bands through the Phanerozoic.
+This study combines reconstructions of the original surface extent of LIPs through the Phanerozoic, a paleogeographic model, and LIP erosion models to estimate LIP area in all latitudinal bands through the Phanerozoic.
 We then compare these LIP area time series to the latitudinal extent of continental ice sheets - a proxy for Earth's long term climate state.
@@
 \ref{fig:LIP_correlation}).
-All other scenarios yield a near zero or weak negative correlation coefficient.
+All other scenarios (including both total LIP area and tropical LIP area) yield a near zero or weak negative correlation coefficient (Fig.
+\ref{fig:LIP_correlation}).
 The weak positive correlation of the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario relative to the other scenarios can be primarily attributed to the complete removal of the CAMP, which was emplaced during an extended interval of ice-free conditions, as well as the effect of the longer decay half-life extending the duration of the earlier two peaks, such that they overlap more with the Late Ordovician and Permo-Carboniferous glacial intervals.
@@
 regolith development itself) in a number of the provinces; 2) the current extent of LIP exposure may be incomplete and an underestimate for some of the provinces; 3) the initial LIP extents are typically poorly constrained and are likely overestimates which could be resulting in inflated interpreted decay rates; and 4) the relationship between LIP area and volume is poorly constrained.
-Future efforts that improve the LIP database, such as developing better-constrained estimates of original LIP extent, constraining burial and uplift histories, and refining the timing of eruptions associated with LIPs, will improve analyses such as that in this contribution.
+Future efforts that improve the LIP database, such as developing better-constrained estimates of original LIP extent, constraining burial and uplift histories, and refining the timing of eruptions associated with LIPs, will improve analyses that consider the LIP record in its entirety, such as that in this contribution.
 
@@
 	\includegraphics[width=\textwidth]{Manuscript/Figures/LIP_Map.pdf}
-	\caption{Map of current extent of volcanic lithologies associated with LIPs that erupted between 520~Ma and the present, as well as the estimates of the initial LIP area used in the area analysis (modified slightly from \citealp{Ernst2017a}).}
+	\caption{Map of current extent of volcanic lithologies associated with LIPs that erupted between 520~Ma and the present, as well as the estimates of the initial LIP area used in the area analysis (modified slightly from \citealp{Ernst2017a} to ensure that all currently exposed volcanic lithologies are encapsulated by the initial LIP area polygons).}
 	\label{fig:LIP_map}

### 2019-03-19 Manuscript/Manuscript.tex

@@
 The total LIP area and the LIP area reconstructed within the tropical rain belt (15\textdegree~S to 15\textdegree~N) were calculated for the various decay and burial scenarios at a resolution of 5~Myr (Fig.
-\ref{fig:LIP_area}B and C).
+\ref{fig:LIP_Areas}B and C).
 We use 15\textdegree~S to 15\textdegree~N as the definition of the tropical rain belt, as those latitudes approximately correspond with precipitation greater than 1.0~m/yr in modern climatalogical data \citep{Kalnay1996a}.
@@
 To evaluate the relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic (\citealp{Macdonald2018a}; Fig.
-\ref{fig:LIP_area}E).
+\ref{fig:LIP_Areas}E).
 The goal in doing so is to evaluate the hypothesis that there is a correlation between LIP area in the tropics and Earth's long-term climate state.
@@
 In the `t$_{1/2}$ = 36~Myr' scenario, we observe three major peaks and one minor peak in the calculated LIP area within the tropics (Fig.
-\ref{fig:LIP_area}C).
+\ref{fig:LIP_Areas}C).
 The first major peak ca.
@@
 66~Ma Deccan LIP into the tropics (Fig.
-\ref{fig:LIP_area}A).
+\ref{fig:LIP_Areas}A).
 When we account for burial (`t$_{1/2}$ = 36~Myr + 50\% burial' and `t$_{1/2}$ = 36~Myr + 100\% burial' scenarios), only the latter two of these four peaks are affected -- the ca.
@@
 The only scenario which results in a positive Pearson correlation coefficient (0.10) between LIP area in the tropics (Fig.
-\ref{fig:LIP_area}C) and the ice extent record (Fig.
-\ref{fig:LIP_area}E) is the scenario with the slow decay rate and complete burial of LIPs associated with rifting (the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario; Fig.
+\ref{fig:LIP_Areas}C) and the ice extent record (Fig.
+\ref{fig:LIP_Areas}E) is the scenario with the slow decay rate and complete burial of LIPs associated with rifting (the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario; Fig.
 \ref{fig:LIP_correlation}).
@@
 To assess the statistical significance of the correlation (or lack thereof) implied by the Pearson correlation coefficients, we simulated the four glacial episodes (Fig.
-\ref{fig:LIP_area}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and \% overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these 100,000 simulations (Fig.
+\ref{fig:LIP_Areas}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and \% overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these 100,000 simulations (Fig.
 \ref{fig:LIP_correlation}).
@@
 This Franklin LIP tropical area at the time of emplacement is approximately equivalent to the Cenozoic peak, and is smaller than the other Phanerozoic peaks (Fig.
-\ref{fig:LIP_area}C).
+\ref{fig:LIP_Areas}C).
 The ca.
@@
 We reconstruct the paleolatitude of the Umkondo LIP at the time of emplacement using the paleomagnetic pole of \citet{Swanson-Hysell2015b}, and find that effectively all of the LIP (or $\sim$2.0~Mkm$^{2}$) erupted within the tropics, an area that is slightly smaller than that for the Franklin LIP (Fig.
-\ref{fig:LIP_area}C).
+\ref{fig:LIP_Areas}C).
 
@@
 \begin{center}
-	\includegraphics[width=0.8\textwidth]{Manuscript/Figures/LIP_Areas.pdf}
+	\includegraphics[width=0.9\textwidth]{Manuscript/Figures/LIP_Areas.pdf}
 	\caption{\textbf{A)} LIPs included in this analysis.
@@
 The arrows to the right indicate reconstructed tropical LIP area at the time of emplacement for the ca.
-720~Ma Franklin LIP and the ca.
-1109~Ma Umkondo LIP.
-\textbf{D)} Heat map for the preferred post-emplacement model showing the latitudinal distribution of LIP area.
-\textbf{E)} Latitudinal extent of land ice away from the poles through time \citep{Macdonald2018a}.}
-	\label{fig:LIP_area}
+720 and 1109~Ma Franklin and Umkondo LIPs.
+The black circles show the paleogeographic reconstruction times in Fig.
+\ref{fig:Reconstruction_Snapshots}.
+\textbf{D)} Heat map showing the latitudinal distribution of LIP area for the preferred post-emplacement model.
+\textbf{E)} Latitudinal extent of land ice away from the poles through time \citep{Macdonald2018a} and compilation of pCO$_{2}$ proxies \citep{Foster2017a}.
+Grey bands indicate 68 and 95\% confidence intervals for a LOESS fit through the pCO$_{2}$ proxy data.
+Error bars indicate standardized uncertainties as developed in \citep{Foster2017a}.}
+	\label{fig:LIP_Areas}
+\end{center}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
+	\includegraphics[width=0.9\textwidth]{Manuscript/Figures/Reconstruction_Snapshots.pdf}
+	\caption{}
+	\label{fig:Reconstruction_Snapshots}
 \end{center}

### 2019-03-20 Manuscript/Manuscript.tex

@@
 However, such proxy compilations are problematic as they are poorly calibrated in deep time, prone to alteration, and reliant on a myriad of assumptions.
-We thus prefer to use the latitudinal extent of land ice to reflect Earth's overall climate state throughout the Phanerozoic.
+Even when stringent quality criteria and the latest understanding of each of the \textit{p}CO$_2$ proxies have been applied to available \textit{p}CO$_2$ records \citep{Foster2017a}, both significant uncertainty in the estimated \textit{p}CO$_2$ for any given data point as well as disagreement between techniques remain (Fig.
+\ref{fig:LIP_Areas}E).
+For instance, in the Late Triassic ($\sim$240-200~Ma), estimates of \textit{p}CO$_2$ span $\sim$3000~ppm.
+Even a probabilistic approach to a large \textit{p}CO$_2$ proxy data set can not constrain \textit{p}CO$_2$ at the 95\% confidence level to within a few hundred ppm for any given time interval, especially when we look deeper in time than the Cenozoic (\citealp{Foster2017a}; Fig.
+\ref{fig:LIP_Areas}E).
+Nevertheless, despite these shortcomings, the \textit{p}CO$_2$ proxy record is broadly consistent with the ice extent record anyway (Fig.
+\ref{fig:LIP_Areas}E) - sparse \textit{p}CO$_2$ proxy data decreases $\sim$400-310~Ma as the Late Devonian glacial interval occurs and the Permo-Carboniferous glacial interval begins and waxes, \textit{p}CO$_2$ proxy data roughly increases $\sim$310-240~Ma as the Permo-Carboniferous glacial interval wanes and ends, \textit{p}CO$_2$ proxy data remains relatively high $\sim$240-40~Ma when no glacial intervals occur, and \textit{p}CO$_2$ proxy data roughly decreases $\sim$40-0~Ma as the Cenozoic glacial interval begins.
+Given these considerations, we thus prefer to use the latitudinal extent of land ice to reflect Earth's overall climate state throughout the Phanerozoic at this time3.
 
@@
 \textbf{D)} Heat map showing the latitudinal distribution of LIP area for the preferred post-emplacement model.
-\textbf{E)} Latitudinal extent of land ice away from the poles through time \citep{Macdonald2018a} and compilation of pCO$_{2}$ proxies \citep{Foster2017a}.
-Grey bands indicate 68 and 95\% confidence intervals for a LOESS fit through the pCO$_{2}$ proxy data.
-Error bars indicate standardized uncertainties as developed in \citep{Foster2017a}.}
+\textbf{E)} Latitudinal extent of land ice away from the poles through time \citep{Macdonald2018a} and compilation of \textit{p}CO$_{2}$ proxies \citep{Foster2017a} (y-axis reversed, and in log-scale).
+Error bars indicate standardized uncertainties, and grey bands indicate 68 and 95\% confidence intervals for Monte Carlo resampled LOESS fits through the \textit{p}CO$_{2}$ proxy data.}
 	\label{fig:LIP_Areas}
@@
 \begin{center}
-	\includegraphics[width=0.9\textwidth]{Manuscript/Figures/Reconstruction_Snapshots.pdf}
-	\caption{}
+	\includegraphics[width=\textwidth]{Manuscript/Figures/Reconstruction_Snapshots.pdf}
+	\caption{Paleogeographic reconstructions for times that correspond to peaks of LIP area in the tropics (Fig.
+\ref{fig:LIP_Areas}C).
+The opacity of LIP polygons indicates their effective weathering area at the time of the reconstruction as a percentage of initial LIP area, under the preferred post-emplacement model of `t$_{1/2}$ = 36~Myr + 50\% burial'.}
 	\label{fig:Reconstruction_Snapshots}

### 2019-03-27 Code/recon_tools.py
add preliminary evaporites analysis

@@ -559,6 +559,42 @@ def get_lengths_in_bands(reconstructed_feature_geometries, lat_mins, lat_maxs):
     return accumulated_lengths, length_polylines, polyline_attributes
 
 
+def get_points_in_bands(reconstructed_feature_geometries, lat_mins, lat_maxs):
+    """
+    Get the number of all features in each latitude band.
+
+    Parameters
+    ----------
+    reconstructed_feature_geometries : list
+        list of reconstructed features
+    lat_mins : array-like
+        array-like of latitude minimums
+    lat_maxs : array_like
+        array_like of latitude maximums
+
+    Returns
+    -------
+    points : array
+        list of total number of points in each latitude band
+    """
+    # storage vector
+    points = np.zeros(len(lat_mins))
+
+    # iterate over each point
+    for i in range(len(reconstructed_feature_geometries)):
+
+        # get the latitude of the point
+        current_point = reconstructed_feature_geometries[i].get_reconstructed_geometry()
+        current_lat = current_point.to_lat_lon_array()[:,0]
+
+        # put the point where it belongs
+        for j in range(len(points)):
+            if current_lat>lat_mins[j] and current_lat<lat_maxs[j]:
+                points[j] = points[j] + 1
+
+    return points
+
+
 ########## PLOTTING FUNCTIONS ##########
 
 
@@ -598,6 +634,36 @@ def plot_reconstructed_features(ax, reconstructed_feature, **kwargs):
         # note that ccrs.Geodetic() must be used to make sure that polygons that cross the dateline are drawn properly
         poly = Polygon(zip(lons,lats))
         ax.add_geometries([poly], ccrs.Geodetic(), **kwargs)
+
+
+def plot_reconstructed_points(ax, reconstructed_feature, **kwargs):
+    """
+    Plot reconstructed points output by pygplates.
+
+    Parameters
+    ----------
+    ax : axis handle
+        axis set up by cartopy on which to plot
+
+    reconstructed_feature : reconstructed features
+        reconstructed features output by pygplates
+
+    Other Parameters
+    ----------------
+    **kwargs passed to plt.add_geometries
+
+    Returns
+    -------
+    None.
+    """
+    for n in range(len(reconstructed_feature)):
+
+        # pull out lat/lon vertices
+        lat_lon_array = reconstructed_feature[n].get_reconstructed_geometry().to_lat_lon_array()
[…]

### 2019-03-27 Manuscript/Manuscript.tex

@@
 \fancyhf{}
-\lhead{Park et al., in preparation for AGU book entitled \textit{Environmental Change and Large Igneous Provinces}}
+\lhead{Park et al., in review for AGU book entitled \textit{Environmental Change and Large Igneous Provinces: The Deadly Kiss of LIPs}}
 \rhead{\thepage}
@@
 
-\noindent\textit{This article is in preparation for submission to the AGU book \textit{Environmental Change and Large Igneous Provinces}.}
+\noindent\textit{This article is in review for the AGU book \textit{Environmental Change and Large Igneous Provinces: The Deadly Kiss of LIPs}.}
 
@@
 \ref{fig:LIP_Areas}B and C).
-We use 15\textdegree~S to 15\textdegree~N as the definition of the tropical rain belt, as those latitudes approximately correspond with precipitation greater than 1.0~m/yr in modern climatalogical data \citep{Kalnay1996a}.
-The consistent paleolatitudes of evaporite deposits in the subtropics demonstrate the stability of the large-scale atmospheric circulation that gives rise to intense precipitation in the tropical rain belt \citep{Evans2006a}.
 
-To evaluate the relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic (\citealp{Macdonald2018a}; Fig.
+We use 15\textdegree~S to 15\textdegree~N as the definition of the tropical rain belt, as those latitudes approximately correspond with a sharp increase in zonal mean precipitation to values greater than 1.0~m/yr in modern climatalogical data \citep{Kalnay1996a}.
+In \citet{Evans2006a}, the reconstructed paleolatitudes of evaporite deposits throughout the Phanerozoic are shown to be consistently in the subtropics, suggesting the stability of the large-scale atmospheric circulation that gives rise to intense precipitation in the tropical rain belt.
+However, subsequent work by \citet{Boucot2013a} and \citet{Cao2018a} has shown that evaporite deposits did form at or near the equator at times in the Phanerozoic.
+While some of this variability could be attributed to waxing/waning of the width of the tropical rain belt as a whole, it is important to note that there can be large deviations in local precipitation from the zonal mean due to factors such as orographic effects, ocean current circulation, or continentality (i.e.
+how dispersed or amalgamated the continents are).
+For instance, much of the Early Cretaceous low-latitude evaporite deposits were formed in basins which were located deep within arid continental interiors at the time of deposition \citep{Boucot2013a, Cao2018a}.
+A limitation of the LIP analysis described in this study is that it does not account for such deviations in local precipitation from the zonal mean (due to the infeasibility of running a global climate model at each time-step in the analysis).
+However, given that evaporite deposits are distributed bimodally about the equator in the subtropics for the vast majority ($\sim$80\%) of the past $\sim$420~Myr \citep{Cao2018a}, assuming overall stability of the large-scale atmospheric circulation that gives rise to intense precipitation in the tropical rain belt appears to be warranted to first-order.
+Nevertheless, to test the sensitivity of our results to the width of the tropical rain belt, we also performed the tropical LIP area calculations assuming a tropical rain belt of $\pm$10\textdegree and $\pm$20\textdegree.
+We find that both of these additional analyses result in a conclusion that is qualitatively the same as that found for an assumed tropical rain belt of $\pm$15\textdegree REF.
+
+To evaluate the relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic (\citealp{Macdonald2019a}; Fig.
 \ref{fig:LIP_Areas}E).
@@
 Another approach would be to compare LIP area records to proxy compilations of \textit{p}CO$_2$ (as done by \citealp{Johansson2018a}).
-However, such proxy compilations are problematic as they are poorly calibrated in deep time, prone to alteration, and reliant on a myriad of assumptions.
+However, such \textit{p}CO$_2$ proxies are problematic as they are poorly calibrated in deep time, prone to alteration, and reliant on a myriad of assumptions.
 Even when stringent quality criteria and the latest understanding of each of the \textit{p}CO$_2$ proxies have been applied to available \textit{p}CO$_2$ records \citep{Foster2017a}, both significant uncertainty in the estimated \textit{p}CO$_2$ for any given data point as well as disagreement between techniques remain (Fig.
@@
 \ref{fig:LIP_Areas}E) - sparse \textit{p}CO$_2$ proxy data decreases $\sim$400-310~Ma as the Late Devonian glacial interval occurs and the Permo-Carboniferous glacial interval begins and waxes, \textit{p}CO$_2$ proxy data roughly increases $\sim$310-240~Ma as the Permo-Carboniferous glacial interval wanes and ends, \textit{p}CO$_2$ proxy data remains relatively high $\sim$240-40~Ma when no glacial intervals occur, and \textit{p}CO$_2$ proxy data roughly decreases $\sim$40-0~Ma as the Cenozoic glacial interval begins.
-Given these considerations, we thus prefer to use the latitudinal extent of land ice to reflect Earth's overall climate state throughout the Phanerozoic at this time3.
+Given these considerations, we thus prefer to use the latitudinal extent of land ice to reflect Earth's overall climate state throughout the Phanerozoic.
 
@@
 The ca.
-1109~Ma Umkondo is another Precambrian LIP that is constrained to have erupted in the tropics, although it is not known to be associated with any glaciation (no glacial deposits are found within the contemporaneous Midcontinent Rift basin; \citealp{Swanson-Hysell2018a}).
+1109~Ma Umkondo is another Precambrian LIP that is constrained to have erupted in the tropics, although it is not known to be associated with any glaciation (no glacial deposits are found within the contemporaneous Midcontinent Rift basin; \citealp{Swanson-Hysell2019a}).
 We reconstruct the paleolatitude of the Umkondo LIP at the time of emplacement using the paleomagnetic pole of \citet{Swanson-Hysell2015b}, and find that effectively all of the LIP (or $\sim$2.0~Mkm$^{2}$) erupted within the tropics, an area that is slightly smaller than that for the Franklin LIP (Fig.
@@
 \textbf{D)} Heat map showing the latitudinal distribution of LIP area for the preferred post-emplacement model.
-\textbf{E)} Latitudinal extent of land ice away from the poles through time \citep{Macdonald2018a} and compilation of \textit{p}CO$_{2}$ proxies \citep{Foster2017a} (y-axis reversed, and in log-scale).
+\textbf{E)} Latitudinal extent of land ice away from the poles through time \citep{Macdonald2019a} and compilation of \textit{p}CO$_{2}$ proxies \citep{Foster2017a} (y-axis reversed, and in log-scale).
 Error bars indicate standardized uncertainties, and grey bands indicate 68 and 95\% confidence intervals for Monte Carlo resampled LOESS fits through the \textit{p}CO$_{2}$ proxy data.}

### 2019-03-28 Manuscript/Manuscript.tex
checkpoint after initial implementation of changes in response to reviewer comments

@@
 
-Outlines of the original surface extent of LIPs through the Phanerozoic (Fig.
-\ref{fig:LIP_map}) were slightly modified from the compilation of \citet{Ernst2017a}, and emplacement ages were taken from the literature (Table \ref{tab:LIPs}).
-The compilation of \citet{Ernst2017a} seeks to reconstruct the original surface extent of LIPs with the caveat that there can be significant uncertainty with doing so, particularly for older more deeply eroded LIPs.
+Outlines of the original surface extent of continental LIPs through the Phanerozoic (Fig.
+\ref{fig:LIP_map}) were slightly modified (to ensure that all currently exposed volcanic lithologies are encapsulated by the initial LIP area polygons) from the compilation of \citet{Ernst2017a} and \citet{Ernst2019a}, and emplacement ages were taken from the literature (Table \ref{tab:LIPs}).
+The LIP original surface extent compilation seeks to reconstruct the original surface extent of LIPs with the caveat that there can be significant uncertainty with doing so, particularly for older more deeply eroded LIPs.
 These polygons encapsulate all of the rocks associated with a given LIP, including dikes, sills, and layered intrusions, in addition to subaerial volcanics (Fig.
@@
 However, this approach likely provides the best estimates available of original extent for ancient LIPs.
-The extents of presently-exposed volcanics associated with LIPs that were used for present-day area estimates, and the construction of the original extent polygons by \citet{Ernst2017a}, were taken from a number of resources including the PLATES compilation \citep{Coffin2006a} and more recent compilation efforts associated with the LIPs Reconstruction Project (\citealp{Ernst2013a}; Table \ref{tab:LIPs}).
+The extents of presently-exposed volcanics associated with LIPs that were used for present-day area estimates, and the construction of the original extent polygons by \citet{Ernst2017a} and \citet{Ernst2019a}, were taken from a number of resources including the PLATES compilation \citep{Coffin2006a} and more recent compilation efforts associated with the LIPs Reconstruction Project (\citealp{Ernst2013a}; Table \ref{tab:LIPs}).
 
@@
 The goal in doing so is to evaluate the hypothesis that there is a correlation between LIP area in the tropics and Earth's long-term climate state.
-The land ice record is an imperfect tracker of climate, as it is insensitive to changes in temperature during non-glacial intervals and is influenced by additional factors such as the physical geography of the continents during glacial intervals.
+The land ice record is an imperfect tracker of climate, as it is insensitive to changes in temperature during non-glacial intervals, is influenced by additional factors such as the physical geography of the continents during glacial intervals, and is potentially vulnerable to removal from the observable geologic record via erosion and/or burial.
 Nevertheless, it forms a physical record of Earth's climate through time and delineates glacial and non-glacial climate states.
@@
 This approach accounts for the reality that spurious correlation can arise between auto-correlated data sets such as these, where each value is not independent, but is instead dependent on the previous state of the system.
-For the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario, 28\% of the randomly timed glacial intervals correlate better with the LIP area in the tropics.
+For the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario, 28\% of the randomly timed glacial interval simulations correlate better with the LIP area in the tropics.
 With an associated p-value of 0.28, the null hypothesis that glacial intervals do not correlate to LIP area in the tropics cannot be rejected.
@@
 \ref{fig:LIP_correlation}).
-While this result need not imply that there is no increase in global weatherability from the emplacement of LIPs, it does suggest that LIP area is not the fundamental control on Earth's long-term climate state.
+While this result need not imply that there is no increase in global weatherability from the emplacement of LIPs, it does suggest that changes in planetary weatherability associated with LIPs are not the fundamental control on whether Earth is in a glacial or non-glacial climate state.
 
@@
 \toprule
-\textbf{name} & \textbf{age} & \textbf{age ref.} & \textbf{original} & \textbf{present} & \textbf{present area ref.} & \textbf{present/} & \textbf{half-life}$^{1}$ & \textbf{buried?} \\
+\textbf{name} & \textbf{age} & \textbf{age ref.} & \textbf{original$^{1}$} & \textbf{present} & \textbf{present area ref.} & \textbf{present/$^{2}$} & \textbf{half-life}$^{3}$ & \textbf{buried?} \\
  & \textbf{(Ma)} & & \textbf{area} & \textbf{area} & & \textbf{original} & \textbf{(Myr)} & \\
@@
  & & & & & & & & \\
-Karoo & 183 & \cite{Burgess2015a} & 3.21 & 0.15 & de Kock compilation$^{2}$ & 0.05 & 41.3 & no \\
+Karoo & 183 & \cite{Burgess2015a} & 3.21 & 0.15 & de Kock compilation$^{4}$ & 0.05 & 41.3 & no \\
  & & & & & & & & \\
@@
  & & & & & & & & \\
-CAMP & 201 & \cite{Blackburn2013a} & 11.46 & 0.23 & Marzoli and Parisio compilation$^{3}$ & 0.02 & 35.7 & partial \\
+CAMP & 201 & \cite{Blackburn2013a} & 11.46 & 0.23 & Marzoli and Parisio compilation$^{5}$ & 0.02 & 35.7 & partial \\
  & & & & & & & & \\
@@
 \begin{tablenotes}[para,flushleft]
-$^{1}$assuming exponential decay with form $N(t) = 2^{t/t_{1/2}}$.
+$^{1}$calculated from the continental portion of polygons within the LIP original surface extent compilation of \citet{Ernst2017a} and \citet{Ernst2019a}.
 
-$^{2}$from ArcGIS compilation produced by M. de Kock for the LIPs Reconstruction Project \citep{Ernst2013a}.
+$^{2}$present area / original area
 
-$^{3}$from ArcGIS compilation produced by A. Marzoli and L. Parisio for the LIPs Reconstruction Project \citep{Ernst2013a}.
+$^{3}$assuming exponential decay with form $N(t) = 2^{t/t_{1/2}}$.
+
+$^{4}$from ArcGIS compilation produced by M. de Kock for the LIPs Reconstruction Project \citep{Ernst2013a}.
+
+$^{5}$from ArcGIS compilation produced by A. Marzoli and L. Parisio for the LIPs Reconstruction Project \citep{Ernst2013a}.
 \end{tablenotes}
 \label{tab:LIPs}
+\end{table}
+
+\begin{table}[h!]
+\scriptsize
+\caption{Statistics of correlation between large igneous province area and ice extent.}
+\resizebox{\textwidth}{!}{
+\begin{tabular}{>{\raggedright}p{2.5cm}cccccccccccccccc}
+\toprule
+\textbf{} & \multicolumn{4}{c}{\textbf{total}} & \multicolumn{4}{c}{\textbf{within tropics $\pm$10$^{\circ}$}} & \multicolumn{4}{c}{\textbf{within tropics $\pm$15$^{\circ}$}} & \multicolumn{4}{c}{\textbf{within tropics $\pm$20$^{\circ}$}} \\
+\textbf{} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c}{\textbf{\% overlap}} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c}{\textbf{\% overlap}} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c}{\textbf{\% overlap}} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c}{\textbf{\% overlap}} \\
+\textbf{scenario} & \textbf{val.}$^{1}$ & \textbf{p-val.}$^{2}$ & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} \\
+\midrule
+t$_{1/2}$ = 36~Myr & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X \\
+&&&&&&&&&&&&&&&& \\
+t$_{1/2}$ = 36~Myr + 50\% burial & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X \\
+&&&&&&&&&&&&&&&& \\
+t$_{1/2}$ = 36~Myr + 100\% burial & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X \\
+&&&&&&&&&&&&&&&& \\
+t$_{1/2}$ = 120~Myr + 100\% burial & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X \\
+\bottomrule
+\end{tabular}
+}
+\begin{tablenotes}[para,flushleft]
+$^{1}$`val.' refers to the computed correlation coefficient/\% overlap between LIP area and the actual ice extent record.
+
+$^{2}$`p-val.' refers to the fraction of randomly timed glacial interval simulations that correlate/overlap better with LIP area than the actual ice extent record (i.e.
+the p-value with respect to the null hypothesis of no correlation/overlap).
+\end{tablenotes}
+\label{tab:stats}
 \end{table}
@@
 	\includegraphics[width=\textwidth]{Manuscript/Figures/LIP_Map.pdf}
-	\caption{Map of current extent of volcanic lithologies associated with LIPs that erupted between 520~Ma and the present, as well as the estimates of the initial LIP area used in the area analysis (modified slightly from \citealp{Ernst2017a} to ensure that all currently exposed volcanic lithologies are encapsulated by the initial LIP area polygons).}
+	\caption{Map of current extent of volcanic lithologies associated with LIPs that erupted between 520~Ma and the present, as well as the estimates of the initial LIP area used in the area analysis (modified slightly from \citealt{Ernst2017a} and \citealt{Ernst2019a} to ensure that all currently exposed volcanic lithologies are encapsulated by the initial LIP area polygons).}
 	\label{fig:LIP_map}

### 2019-04-09 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 One of the hypothesized effects of the emplacement of large igneous provinces (LIPs) is planetary cooling on million-year timescales associated with enhanced silicate weathering.
-This study combines reconstructions of the original surface extent of LIPs through the Phanerozoic, a paleogeographic model, and LIP erosion models to estimate LIP area in all latitudinal bands through the Phanerozoic.
+This study combines reconstructions of the original surface extent of LIPs through the Phanerozoic, a paleogeographic model, and a paramaterization of LIP erosion to estimate LIP area in all latitudinal bands through the Phanerozoic.
 This analysis reveals no significant correlation between total LIP area, nor LIP area in the tropics, and the extent of continental ice sheets.
@@
 In the Cenozoic, the movement of the Deccan LIP into the tropical rain belt, in concert with the low-latitude emplacement of the Ethiopian Traps, has been implicated in drawing down CO$_2$ levels in the lead-up to Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
-More recently, \citet{Johansson2018a} used a paleogeographic model to suggest that tropical LIP area correlates with Phanerozoic climate change.
+More recently, \citet{Johansson2018a} used a paleogeographic model to suggest that tropical LIP area correlates with Phanerozoic climate change through comparisons to \textit{p}CO$_{2}$ proxy records.
 
@@
 
-This study combines reconstructions of the original surface extent of LIPs through the Phanerozoic, a paleogeographic model, and LIP erosion models to estimate LIP area in all latitudinal bands through the Phanerozoic.
+This study combines reconstructions of the original surface extent of LIPs through the Phanerozoic, a paleogeographic model, and a parameterization of LIP erosion to estimate LIP area in all latitudinal bands through the Phanerozoic.
 We then compare these LIP area time series to the latitudinal extent of continental ice sheets - a proxy for Earth's long term climate state.
@@
 The polygons also assume complete surface coverage between wide-spread remnants, creating further potential for these outlines being over-estimates.
-However, this approach likely provides the best estimates available of original extent for ancient LIPs.
+However, despite these uncertainties, this approach likely provides the best estimates available of original extent for ancient LIPs.
 The extents of presently-exposed volcanics associated with LIPs that were used for present-day area estimates, and the construction of the original extent polygons by \citet{Ernst2017a} and \citet{Ernst2019a}, were taken from a number of resources including the PLATES compilation \citep{Coffin2006a} and more recent compilation efforts associated with the LIPs Reconstruction Project (\citealp{Ernst2013a}; Table \ref{tab:LIPs}).
@@
 After LIPs are emplaced, they will progressively erode.
-In order to account for this decrease in area with time, \citet{Godderis2017a} took the approach of fitting an exponential decay function to estimates of the original surface extent and the current surface extent for 5 LIPs.
-They used the resulting exponential decay constants to develop a first-order model to estimate changing LIP area through time.
-We extend this approach to 19 basaltic LIPs for which we have estimates of the original surface extent of the province and the current surface extent of rocks associated with it (Fig.
+In order to account for the associated decrease in area with time, \citet{Godderis2017a} took the approach of fitting an exponential decay function to estimates of the original surface extent and the current surface extent for 5 LIPs.
+They used the resulting exponential decay constants to develop a first-order parameterization to estimate changing LIP area through time.
+We extend this approach to 19 basaltic LIPs for which there are estimates of the original surface extent of the province and the current surface extent of rocks associated with it (Fig.
 \ref{fig:LIP_preservation}).
-While there are significant uncertainties associated with the area estimates, this compilation suggests that an exponential decay function is an appropriate first-order model for the progressive reduction in LIP area (Fig.
+While there are significant uncertainties associated with the area estimates, this compilation suggests that an exponential decay function is an appropriate first-order representation of the progressive reduction in LIP area (Fig.
 \ref{fig:LIP_preservation}).
@@
 The `50\% burial' scenario imposes instantaneous burial of 50$\%$ of the LIP area while the `100\% burial' imposes instantaneous burial of the entire LIP as an end-member scenario.
-The LIP area analysis uses all of the available combinations of these distinct decay and burial scenarios.
+A limitation of our treatment of LIPs that are co-located with rifting is that we instantaneously `bury' all of these LIPs to the same degree, when in fact the degree/timing of burial of these LIPs may vary substantially.
+The LIP area analysis uses all of the available combinations of the distinct decay and burial scenarios described above.
 
@@
 The LIP polygons and tectonic units were reconstructed from 520~Ma to the present utilizing the paleogeographic model of \citet{Torsvik2016a} in the spin axis reference frame (anchor plate ID of 1).
-This paleogeographic model was updated to include a revision to Ordovician Laurentia \citep{Swanson-Hysell2017a} and Paleozoic Asia \citep{Domeier2018a}.
-Reconstructions and area calculations within latitude bands utilized the pyGPlates function library and custom Python scripts documented within a Jupyter notebook that reproduces the analysis and development of the associated visualizations.
-The total LIP area and the LIP area reconstructed within the tropical rain belt (15\textdegree~S to 15\textdegree~N) were calculated for the various decay and burial scenarios at a resolution of 5~Myr (Fig.
+This paleogeographic model was updated to include revisions to Ordovician Laurentia \citep{Swanson-Hysell2017a} and Paleozoic Asia \citep{Domeier2018a}.
+Reconstructions and area calculations within latitude bands utilized the pyGPlates function library and custom Python scripts.
+The total LIP area and the LIP area reconstructed within the tropical rain belt were calculated for the various decay and burial scenarios at a resolution of 5~Myr (Fig.
 \ref{fig:LIP_Areas}B and C).
+All the data and code necessary to reproduce the analyses and figures presented in this study can be downloaded from GitHub: \url{https://github.com/Swanson-Hysell-Group/Large_Igneous_Provinces}.
 
 We use 15\textdegree~S to 15\textdegree~N as the definition of the tropical rain belt, as those latitudes approximately correspond with a sharp increase in zonal mean precipitation to values greater than 1.0~m/yr in modern climatalogical data \citep{Kalnay1996a}.
-In \citet{Evans2006a}, the reconstructed paleolatitudes of evaporite deposits throughout the Phanerozoic are shown to be consistently in the subtropics, suggesting the stability of the large-scale atmospheric circulation that gives rise to intense precipitation in the tropical rain belt.
+In \citet{Evans2006a}, the reconstructed paleolatitudes of basins with thick, basin-wide evaporite deposition are shown to be consistently in the subtropics throughout the Phanerozoic (and potentially beyond), suggesting the stability of the large-scale atmospheric circulation that gives rise to intense precipitation in the tropical rain belt.
 However, subsequent work by \citet{Boucot2013a} and \citet{Cao2018a} has shown that evaporite deposits did form at or near the equator at times in the Phanerozoic.
-While some of this variability could be attributed to waxing/waning of the width of the tropical rain belt as a whole, it is important to note that there can be large deviations in local precipitation from the zonal mean due to factors such as orographic effects, ocean current circulation, or continentality (i.e.
-how dispersed or amalgamated the continents are).
+While some of this variability could be attributed to waxing/waning of the width of the tropical rain belt as a whole, it is important to note that there can be large deviations in local precipitation from the zonal mean due to factors such as monsoon-related precipitation, or continentality (i.e.
+how dispersed or amalgamated the continents are) which can lead to aridity in continental interiors.
 For instance, much of the Early Cretaceous low-latitude evaporite deposits were formed in basins which were located deep within arid continental interiors at the time of deposition \citep{Boucot2013a, Cao2018a}.
-A limitation of the LIP analysis described in this study is that it does not account for such deviations in local precipitation from the zonal mean (due to the infeasibility of running a global climate model at each time-step in the analysis).
-However, given that evaporite deposits are distributed bimodally about the equator in the subtropics for the vast majority ($\sim$80\%) of the past $\sim$420~Myr \citep{Cao2018a}, assuming overall stability of the large-scale atmospheric circulation that gives rise to intense precipitation in the tropical rain belt appears to be warranted to first-order.
-Nevertheless, to test the sensitivity of our results to the width of the tropical rain belt, we also performed the tropical LIP area calculations assuming a tropical rain belt of $\pm$10\textdegree and $\pm$20\textdegree.
-We find that both of these additional analyses result in a conclusion that is qualitatively the same as that found for an assumed tropical rain belt of $\pm$15\textdegree REF.
+Furthermore, in contrast to \citet{Evans2006a}, the compilation of evaporite deposits of \citet{Boucot2013a} contains sedimentary sequences in which the occurrence of evaporitic minerals is limited (e.g.
+to nodular evaporitic pseudomorphs).
+Such limited evaporitic mineral precipitation could be attributed to local processes such as periodic restriction of the basin that lead to saturation states that otherwise would not be expected for that latitude.
+Nevertheless, a limitation of the LIP analysis described in this study is that it does not account for deviations in local precipitation from the zonal mean (due to the infeasibility of running a global climate model at each time-step in the analysis).
+However, evaporite deposits are distributed bimodally about the equator in the subtropics for the vast majority of the past $\sim$420~Myr \citep{Cao2018a} and overall stability of the large-scale atmospheric circulation is predicted by climate dynamics \citep{Donohoe2017a}.
+Therefore, the assumption of enhanced precipitation in the tropics throughout the Phanerozoic appears to be warranted to first-order.
+To test the sensitivity of our results to the width of the tropical rain belt, we performed the tropical LIP area calculations with tropical rain belt widths of $\pm$10\textdegree and $\pm$20\textdegree, which yields similar results to a tropical rain belt width of $\pm$15\textdegree (Table \ref{tab:stats}).
 
@@
 The goal in doing so is to evaluate the hypothesis that there is a correlation between LIP area in the tropics and Earth's long-term climate state.
-The land ice record is an imperfect tracker of climate, as it is insensitive to changes in temperature during non-glacial intervals, is influenced by additional factors such as the physical geography of the continents during glacial intervals, and is potentially vulnerable to removal from the observable geologic record via erosion and/or burial.
+The land ice record is an imperfect tracker of climate, as it is insensitive to changes in temperature during non-glacial intervals, is influenced by additional factors such as the physical geography of the continents during glacial intervals, and is potentially vulnerable to removal from the observable geologic record via erosion and burial.
 Nevertheless, it forms a physical record of Earth's climate through time and delineates glacial and non-glacial climate states.
@@
 
-Another approach would be to compare LIP area records to proxy compilations of \textit{p}CO$_2$ (as done by \citealp{Johansson2018a}).
-However, such \textit{p}CO$_2$ proxies are problematic as they are poorly calibrated in deep time, prone to alteration, and reliant on a myriad of assumptions.
+Another approach would be to compare LIP area records to proxy compilations of \textit{p}CO$_2$ (as done by \citealt{Johansson2018a}).
+However, such \textit{p}CO$_2$ proxies are potentially problematic as they can be difficult to calibrate in deep time and can be affected by secondary alteration.
 Even when stringent quality criteria and the latest understanding of each of the \textit{p}CO$_2$ proxies have been applied to available \textit{p}CO$_2$ records \citep{Foster2017a}, both significant uncertainty in the estimated \textit{p}CO$_2$ for any given data point as well as disagreement between techniques remain (Fig.
@@
 \ref{fig:LIP_Areas}E).
-Nevertheless, despite these shortcomings, the \textit{p}CO$_2$ proxy record is broadly consistent with the ice extent record anyway (Fig.
-\ref{fig:LIP_Areas}E) - sparse \textit{p}CO$_2$ proxy data decreases $\sim$400-310~Ma as the Late Devonian glacial interval occurs and the Permo-Carboniferous glacial interval begins and waxes, \textit{p}CO$_2$ proxy data roughly increases $\sim$310-240~Ma as the Permo-Carboniferous glacial interval wanes and ends, \textit{p}CO$_2$ proxy data remains relatively high $\sim$240-40~Ma when no glacial intervals occur, and \textit{p}CO$_2$ proxy data roughly decreases $\sim$40-0~Ma as the Cenozoic glacial interval begins.
-Given these considerations, we thus prefer to use the latitudinal extent of land ice to reflect Earth's overall climate state throughout the Phanerozoic.
+Nevertheless, despite these shortcomings, the \textit{p}CO$_2$ proxy record is broadly consistent with the ice extent record (Fig.
+\ref{fig:LIP_Areas}E) - sparse \textit{p}CO$_2$ proxy data decreases $\sim$400-310~Ma as the Late Devonian glacial interval occurs and the Permo-Carboniferous glacial interval begins and waxes, \textit{p}CO$_2$ proxy data roughly increases $\sim$310-240~Ma as the Permo-Carboniferous glacial interval wanes and ends, \textit{p}CO$_2$ proxy data broadly remains relatively high $\sim$240-40~Ma when no glacial intervals are robustly documented, and \textit{p}CO$_2$ proxy data roughly decreases $\sim$40-0~Ma as the Cenozoic glacial interval begins.
+Given these considerations, we thus prefer to use the latitudinal extent of land ice to reflect Earth's overall climate state throughout the Phanerozoic, despite its own limitations.
+
+We note, however, that Earth's climate state, regardless of whether we choose to represent it using the latitudinal extent of continental ice sheets or proxy compilations of \textit{p}CO$_2$, is dependent on a variety of factors in addition to planetary weatherability.
+For instance, the CO$_{2}$ degassing flux of the solid Earth, ocean circulation patterns, and orbital parameters all could have been significantly different in the past.
+Changes in these additional factors may have modulated the sensitivity of Earth's climate to changes in planetary weatherability.
 
@@
 
-In the `t$_{1/2}$ = 36~Myr' scenario, we observe three major peaks and one minor peak in the calculated LIP area within the tropics (Fig.
+In the `t$_{1/2}$ = 36~Myr' scenario, we observe four main peaks in the calculated LIP area within the tropics (Fig.
 \ref{fig:LIP_Areas}C).
-The first major peak ca.
-510~Ma is associated with the emplacement of the Kalkarindji LIP, the second major peak ca.
-380~Ma is associated with the emplacement of the Kola-Dnieper LIP, the third major peak ca.
-200~Ma is associated with the emplacement of the CAMP, and the minor Cenozoic peak is associated with both the ca.
+The first peak ca.
+510~Ma is associated with the emplacement of the Kalkarindji LIP, the second peak ca.
+380~Ma is associated with the emplacement of the Kola-Dnieper LIP, the third peak ca.
+200~Ma is associated with the emplacement of the CAMP, and the fourth peak is associated with both the ca.
 30~Ma emplacement of the Afar LIP as well as the earlier drift of the ca.
-66~Ma Deccan LIP into the tropics (Fig.
-\ref{fig:LIP_Areas}A).
+66~Ma Deccan LIP into the tropics (Figs.
+\ref{fig:LIP_Areas}A and \ref{fig:Reconstruction_Snapshots}).
 When we account for burial (`t$_{1/2}$ = 36~Myr + 50\% burial' and `t$_{1/2}$ = 36~Myr + 100\% burial' scenarios), only the latter two of these four peaks are affected -- the ca.
-200~Ma peak is attenuated/removed due to the partial/complete burial of the CAMP, and the Cenozoic peak is attenuated due to the partial/complete removal of the Afar LIP.
+200~Ma peak is attenuated/removed due to the partial/complete burial of the CAMP, and the Cenozoic peak is attenuated due to the partial/complete burial of the Afar LIP.
 However, after accounting for burial, a minor area of LIPs remain in the tropics from ca.
@@
 
-The only scenario which results in a positive Pearson correlation coefficient (0.10) between LIP area in the tropics (Fig.
+The only scenario which results in a non-negative Pearson correlation coefficient (0.10) between LIP area in the tropics (Fig.
 \ref{fig:LIP_Areas}C) and the ice extent record (Fig.
@@
 
-To assess the statistical significance of the correlation (or lack thereof) implied by the Pearson correlation coefficients, we simulated the four glacial episodes (Fig.
+To assess the statistical significance of the correlation (or lack thereof) implied by the Pearson correlation coefficients, we applied the approach of \citet{Macdonald2019a} and simulated the four glacial episodes (Fig.
 \ref{fig:LIP_Areas}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and \% overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these 100,000 simulations (Fig.
@@
 With an associated p-value of 0.28, the null hypothesis that glacial intervals do not correlate to LIP area in the tropics cannot be rejected.
-Taking this approach, none of the positive or negative correlations that emerge between the LIP area scenarios are statistically significant.
+Taking this approach, none of the positive or negative correlations that emerge between the LIP area scenarios and the ice extent record are statistically significant.
 
@@
 Recent progress on understanding the relationships between landscapes, topography, and chemical weathering reveals that these effects are quite important \citep{Gabet2009a, Maher2014a}.
-As a result, more recent modeling of chemical weathering that has incorporated such processes highlights the importance of high-relief regions relative to low-relief ones \citep{Godderis2017b}.
+As a result, more recent modeling of chemical weathering that has incorporated such processes highlights the importance of high-relief regions relative to low-relief ones for setting global weatherability \citep{West2012a, Godderis2017b}.
 Without active uplift, soil shielding from regolith development on low-relief LIPs could significantly decrease the local weatherability of a LIP and mute its impact on global weatherability.
-In contrast, processes that lead to continued exhumation of mafic lithologies and the creation of steep topography, particularly in tropical regions, may exert a strong control on global weatherability and long-term climate.
-This interpretation underlies the hypothesis that arc-continent collisions in the tropics during the Ordovician \citep{Swanson-Hysell2017a} and the Cenozoic \citep{Jagoutz2016a} played a significant role in transitions into glacial climate states at those times.
+Furthermore, significant soil shielding can lead to a supply-limited weathering regime in which the weathering rate of the underlying bedrock becomes insensitive to kinetic and equilibrium factors such as temperature and runoff (factors that would, in the absence of soil shielding, lead to relatively high weathering rates in the tropical rain belt).
+In contrast, processes that lead to continued exhumation of mafic lithologies and the creation of steep topography that minimizes soil shielding, particularly in tropical regions, may exert a strong control on global weatherability and long-term climate.
+This interpretation underlies the hypothesis that arc-continent collisions in the tropics during the Ordovician \citep{Swanson-Hysell2017a} and the Cenozoic \citep{Jagoutz2016a} played a significant role in transitions into glacial climate states at those times - a correlation that appears robust throughout the Phanerozoic \citep{Macdonald2019a}.
 
@@
 The approach of seeking to evaluate correlation between LIP area and glaciation is further complicated for Neoproterozoic Snowball Earth events because ice-albedo runaway leads to persistent global glaciation on timescales of tens of millions of years absent of continued forcing through normal carbon cycle processes until sufficient CO$_2$ to drive deglaciation builds up in the absence of silicate weathering \citep{Hoffman2017a}.
-Moreover, cooling past the critical threshold for rapid global glaciation may have occurred on a sub-million year timescale that is not recorded in the geological carbon cycle (e.g.
+Moreover, cooling past the critical threshold for rapid global glaciation may have occurred on a sub-million year timescale (e.g.
 \citealp{Macdonald2017a}).
@@
 1109~Ma Umkondo is another Precambrian LIP that is constrained to have erupted in the tropics, although it is not known to be associated with any glaciation (no glacial deposits are found within the contemporaneous Midcontinent Rift basin; \citealp{Swanson-Hysell2019a}).
-We reconstruct the paleolatitude of the Umkondo LIP at the time of emplacement using the paleomagnetic pole of \citet{Swanson-Hysell2015b}, and find that effectively all of the LIP (or $\sim$2.0~Mkm$^{2}$) erupted within the tropics, an area that is slightly smaller than that for the Franklin LIP (Fig.
+We reconstruct the paleolatitude of the Umkondo LIP at the time of emplacement using the paleomagnetic pole of \citet{Swanson-Hysell2015b}, and find that effectively all of the LIP (or $\sim$2.0~Mkm$^{2}$) erupted within the tropics, an area that is slightly smaller than that estimated for the Franklin LIP (Fig.
 \ref{fig:LIP_Areas}C).
@@
 One such process could have been an increase in global weatherability associated with the collision and accretion of arc terranes within the present-day Arabian-Nubian Shield \citep{Park2018a}.
-A low-latitude continent configuration and abundant arc-continent collisions in the tropics may have led to a cool background climate and the emplacement of the Franklin LIP may have pushed the climate system over the edge.
+A low-latitude continent configuration and abundant arc-continent collisions in the tropics may have led to a cool background climate, and the emplacement of the Franklin LIP may have further increased global weatherability to the point where the ice-albedo runaway could take effect.
 However, tropical LIP area associated with the Franklin was not uniquely high, and therefore an associated increase in global weatherability was not the sole driver of Snowball Earth onset, consistent with the results of the Phanerozoic analysis.
 
-An increase in global weatherability associated with the emplacement of the Franklin LIP in the tropics could have been a contributing factor in addition to other such processes.
-The temporal overlap between Franklin LIP eruptions and the initiation of Sturtian glaciation remains compelling \citep{Macdonald2017a}.
-This overlap continues to suggest that other aspects of LIP emplacement, such as relieving P, Fe, or S limitations on biological productivity \citep{Tziperman2011a, Cox2016a}, or the injection of sulfur aerosols in the stratosphere \citep{Macdonald2017a}, could have played a role in the initiation of low-latitude glaciation.
+The temporal overlap between Franklin LIP eruptions and the initiation of Sturtian glaciation remains compelling \citep{Macdonald2010a, MacLennan2018a}.
+This overlap continues to suggest that other aspects of LIP emplacement, such as the injection of sulfur aerosols in the stratosphere \citep{Macdonald2017a}, could have played a role in the initiation of low-latitude glaciation.
 The latitude and background climate of such aerosol injections is important for their temporary effect on albedo which is maximized at low latitude, and their presence at high concentrations is pre-conditioned on eruption through sedimentary basins hosting evaporite deposits \citep{Macdonald2017a}.
-Sulfur aerosols associated with low-latitude LIPs may have also caused short-term cooling due to changes in albedo during the Phanerozoic, as has been suggested for the CAMP \citep{Schoene2010a} and putative LIPs associated with the Hirnantian ice advance \citep{Jones2017a}.
+Sulfur aerosols associated with low-latitude LIPs may have also caused short-term cooling due to changes in albedo during the Phanerozoic, as has been suggested for the CAMP \citep{Schoene2010a}.
 However, because these cooling events did not result in a Snowball Earth, the climate would have returned to its background climate state within thousands of years \citep{Macdonald2017a}, which would not be observed in our analysis.
 
-The results from this analysis indicate that when the entire LIP database is considered, there is no significant relationship between total LIP area or LIP area in the tropics and the extent of continental ice sheets (Fig.
+The results from this analysis indicate that when the entire LIP database is considered in this way, there is no significant relationship between total LIP area or LIP area in the tropics and the extent of continental ice sheets (Fig.
 \ref{fig:LIP_correlation}).
@@
 \toprule
-\textbf{name} & \textbf{age} & \textbf{age ref.} & \textbf{original$^{1}$} & \textbf{present} & \textbf{present area ref.} & \textbf{present/$^{2}$} & \textbf{half-life}$^{3}$ & \textbf{buried?} \\
+\textbf{name} & \textbf{age} & \textbf{age ref.} & \textbf{original$^{1}$} & \textbf{present$^{2}$} & \textbf{present area ref.} & \textbf{present/$^{3}$} & \textbf{half-life}$^{4}$ & \textbf{buried?} \\
  & \textbf{(Ma)} & & \textbf{area} & \textbf{area} & & \textbf{original} & \textbf{(Myr)} & \\
@@
  & & & & & & & & \\
-HALIP & 95 & \cite{Kingsbury2018a} & 3.60 & 0.15 & \cite{Moosdorf2015a} & 0.04 & 20.7 & no \\
+HALIP & 95 & \cite{Kingsbury2018a} & 3.60 & 0.15 & \cite{Hartmann2012a} & 0.04 & 20.7 & no \\
  & & & & & & & & \\
@@
  & & & & & & & & \\
-Karoo & 183 & \cite{Burgess2015a} & 3.21 & 0.15 & de Kock compilation$^{4}$ & 0.05 & 41.3 & no \\
+Karoo & 183 & \cite{Burgess2015a} & 3.21 & 0.15 & de Kock compilation$^{5}$ & 0.05 & 41.3 & no \\
  & & & & & & & & \\
@@
  & & & & & & & & \\
-CAMP & 201 & \cite{Blackburn2013a} & 11.46 & 0.23 & Marzoli and Parisio compilation$^{5}$ & 0.02 & 35.7 & partial \\
+CAMP & 201 & \cite{Blackburn2013a} & 11.46 & 0.23 & Marzoli and Parisio compilation$^{6}$ & 0.02 & 35.7 & partial \\
  & & & & & & & & \\
@@
  & & & & & & & & \\
-Magdalen & 360 & \cite{Brendan_Murphy1999a} & 0.42 & - & - & - & - & no \\
+Magdalen & 360 & \cite{Murphy1999a} & 0.42 & - & - & - & - & no \\
  & & & & & & & & \\
@@
 \begin{tablenotes}[para,flushleft]
-$^{1}$calculated from the continental portion of polygons within the LIP original surface extent compilation of \citet{Ernst2017a} and \citet{Ernst2019a}.
+\vspace{0.15cm}
+$^{1}$obtained via calculating the area of the continental portions of polygons within the LIP original surface extent compilation of \citet{Ernst2017a} and \citet{Ernst2019a}, shown as blue polygons in Fig.
+\ref{fig:LIP_map}.
+\vspace{0.15cm}
 
-$^{2}$present area / original area
+$^{2}$obtained via calculating the area of polygons from the noted reference of presently-exposed volcanics associated with LIPs, shown as orange polygons in Fig. \ref{fig:LIP_map}.
+\vspace{0.15cm}
 
-$^{3}$assuming exponential decay with form $N(t) = 2^{t/t_{1/2}}$.
+$^{3}$present area / original area
+\vspace{0.15cm}
 
-$^{4}$from ArcGIS compilation produced by M. de Kock for the LIPs Reconstruction Project \citep{Ernst2013a}.
+$^{4}$assuming exponential decay with form $N(t) = 2^{t/t_{1/2}}$.
+\vspace{0.15cm}
 
-$^{5}$from ArcGIS compilation produced by A. Marzoli and L. Parisio for the LIPs Reconstruction Project \citep{Ernst2013a}.
+$^{5}$from ArcGIS compilation produced by M. de Kock for the LIPs Reconstruction Project \citep{Ernst2013a}.
+\vspace{0.15cm}
+
+$^{6}$from ArcGIS compilation produced by A. Marzoli and L. Parisio for the LIPs Reconstruction Project \citep{Ernst2013a}.
 \end{tablenotes}
@@
 \resizebox{\textwidth}{!}{
-\begin{tabular}{>{\raggedright}p{2.5cm}cccccccccccccccc}
+\begin{tabular}{>{\raggedright}p{2.5cm}|cccc|cccc|cccc|cccc}
 \toprule
-\textbf{} & \multicolumn{4}{c}{\textbf{total}} & \multicolumn{4}{c}{\textbf{within tropics $\pm$10$^{\circ}$}} & \multicolumn{4}{c}{\textbf{within tropics $\pm$15$^{\circ}$}} & \multicolumn{4}{c}{\textbf{within tropics $\pm$20$^{\circ}$}} \\
-\textbf{} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c}{\textbf{\% overlap}} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c}{\textbf{\% overlap}} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c}{\textbf{\% overlap}} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c}{\textbf{\% overlap}} \\
-\textbf{scenario} & \textbf{val.}$^{1}$ & \textbf{p-val.}$^{2}$ & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} \\
+\textbf{} & \multicolumn{4}{c|}{\textbf{within tropics $\pm$15$^{\circ}$}$^{1}$} & \multicolumn{4}{c|}{\textbf{total}$^{6}$} & \multicolumn{4}{c|}{\textbf{within tropics $\pm$10$^{\circ}$}} & \multicolumn{4}{c}{\textbf{within tropics $\pm$20$^{\circ}$}} \\
+&&&&&&&&&&&&&&&& \\
+\textbf{} & \multicolumn{2}{c}{\textbf{correlation}$^{2}$} & \multicolumn{2}{c|}{\textbf{\% overlap}$^{3}$} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c|}{\textbf{\% overlap}} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c|}{\textbf{\% overlap}} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c}{\textbf{\% overlap}} \\
+\textbf{scenario} & \textbf{val.}$^{4}$ & \textbf{p-val.}$^{5}$ & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} \\
 \midrule
-t$_{1/2}$ = 36~Myr & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X \\
+t$_{1/2}$ = 36~Myr & -0.19 & 0.81 & 13 & 0.94 & -0.26 & 0.85 & 30 & 0.95 & -0.22 & 0.86 & 13 & 0.95 & -0.17 & 0.77 & 9 & 0.96 \\
 &&&&&&&&&&&&&&&& \\
-t$_{1/2}$ = 36~Myr + 50\% burial & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X \\
+t$_{1/2}$ = 36~Myr + 50\% burial & -0.14 & 0.73 & 22 & 0.85 & -0.25 & 0.84 & 52 & 0.97 & -0.17 & 0.79 & 9 & 0.90 & -0.10 & 0.67 & 26 & 0.86 \\
 &&&&&&&&&&&&&&&& \\
-t$_{1/2}$ = 36~Myr + 100\% burial & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X \\
+t$_{1/2}$ = 36~Myr + 100\% burial & -0.02 & 0.45 & 22 & 0.65 & -0.14 & 0.74 & 65 & 0.94 & -0.08 & 0.54 & 9 & 0.76 & 0.04 & 0.37 & 22 & 0.65 \\
 &&&&&&&&&&&&&&&& \\
-t$_{1/2}$ = 120~Myr + 100\% burial & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X \\
+t$_{1/2}$ = 120~Myr + 100\% burial & 0.10 & 0.32 & 35 & 0.72 & 0.00 & 0.51 & 100 & 1.00 & 0.02 & 0.40 & 26 & 0.77 & 0.19 & 0.24 & 48 & 0.66 \\
 \bottomrule
@@
 \begin{tablenotes}[para,flushleft]
-$^{1}$`val.' refers to the computed correlation coefficient/\% overlap between LIP area and the actual ice extent record.
+\vspace{0.15cm}
+$^{1}$our preferred definition of the tropical rain belt.
+\vspace{0.15cm}
 
-$^{2}$`p-val.' refers to the fraction of randomly timed glacial interval simulations that correlate/overlap better with LIP area than the actual ice extent record (i.e.
+$^{2}$Pearson correlation coefficient between LIP area and the actual ice extent record.
+\vspace{0.15cm}
+
+$^{3}$\% of time when both LIP area is \textgreater30\% of the maximum and ice extent is \textgreater10\textdegree\xspace from the poles.
+\vspace{0.15cm}
+
+$^{4}$`val.' refers to the computed correlation coefficient/\% overlap between LIP area and the actual ice extent record.
+\vspace{0.15cm}
+
+$^{5}$`p-val.' refers to the fraction of randomly timed glacial interval simulations that correlate/overlap better with LIP area than the actual ice extent record (i.e.
 the p-value with respect to the null hypothesis of no correlation/overlap).
+\vspace{0.15cm}
+
+$^{6}$all latitudes.
 \end{tablenotes}
@@
 The ratio of estimates of the present-day area to that of the original area are shown for 19 basaltic LIPs.
-An exponential fit is made to the 13 basaltic LIPs that have not been buried after emplacement (Table \ref{tab:LIPs}), which yields a half-life of $\sim$36~Myr.
+An exponential fit is made to the 13 basaltic LIPs that are interpreted to not have been buried after emplacement (Table \ref{tab:LIPs}), which yields a half-life of $\sim$36~Myr.
 RMSE = root mean square error.}
@@
 = Caribbean-Colombian.
-\textbf{B)} Total LIP area through time for the different post-emplacement models.
+\textbf{B)} Total LIP area through time for the different post-emplacent models.
 Only the `no decay' model excludes pre-Phanerozoic LIPs.
@@
 720 and 1109~Ma Franklin and Umkondo LIPs.
-The black circles show the paleogeographic reconstruction times in Fig.
+The triangles show the paleogeographic reconstruction times in Fig.
 \ref{fig:Reconstruction_Snapshots}.
 \textbf{D)} Heat map showing the latitudinal distribution of LIP area for the preferred post-emplacement model.
-\textbf{E)} Latitudinal extent of land ice away from the poles through time \citep{Macdonald2019a} and compilation of \textit{p}CO$_{2}$ proxies \citep{Foster2017a} (y-axis reversed, and in log-scale).
+\textbf{E)} Latitudinal extent of land ice away from the poles through time \citep{Macdonald2019a} and compilation of \textit{p}CO$_{2}$ proxies \citep{Foster2017a} (\textit{p}CO$_{2}$ y-axis reversed, and in log-scale).
 Error bars indicate standardized uncertainties, and grey bands indicate 68 and 95\% confidence intervals for Monte Carlo resampled LOESS fits through the \textit{p}CO$_{2}$ proxy data.}
@@
 \ref{fig:LIP_Areas}C).
-The opacity of LIP polygons indicates their effective weathering area at the time of the reconstruction as a percentage of initial LIP area, under the preferred post-emplacement model of `t$_{1/2}$ = 36~Myr + 50\% burial'.}
+The opacity of LIP polygons indicates their interpreted remainingu area at the time of the reconstruction as a percentage of initial LIP area, under the preferred post-emplacement model of `t$_{1/2}$ = 36~Myr + 50\% burial'.}
 	\label{fig:Reconstruction_Snapshots}

### 2019-04-10 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 
-One of the hypothesized effects of the emplacement of large igneous provinces (LIPs) is planetary cooling on million-year timescales associated with enhanced silicate weathering.
-This study combines reconstructions of the original surface extent of LIPs through the Phanerozoic, a paleogeographic model, and a paramaterization of LIP erosion to estimate LIP area in all latitudinal bands through the Phanerozoic.
+One of the hypothesized effects of the emplacement of large igneous provinces (LIPs) is planetary cooling on million-year timescales associated with enhanced silicate weathering of the highly weatherable basalt.
+This study combines reconstructions of the original surface extent and emplacement ages of LIPs through the Phanerozoic, a paleogeographic model, and a paramaterization of LIP erosion to estimate LIP area in all latitudinal bands through the Phanerozoic.
 This analysis reveals no significant correlation between total LIP area, nor LIP area in the tropics, and the extent of continental ice sheets.
@@
 Basaltic regions consume more CO$_2$ than regions where bedrock composition is closer to bulk continental crust because of the relatively high concentrations of Ca and Mg (that ultimately sequester carbon through precipitation as carbonate) and the high weathering rates of mafic lithologies \citep{Dessert2003a}.
-Data from basaltic watersheds show that chemical weathering rates are greatest in regions with high runoff.
+Furthermore, data from basaltic watersheds show that chemical weathering rates are greatest in regions with high runoff.
 As a result, CO$_2$ consumption in basaltic regions is most pronounced in the tropical rain belt \citep{Dessert2003a, Hartmann2009a, Hartmann2014a}.
 
-One aspect of large igneous province (LIP) emplacement that has been hypothesized to relate to long-term global climate is the effect such emplacement could have on increasing global weatherability and driving cooling.
+One aspect of large igneous province (LIP) emplacement that has been hypothesized to relate to long-term global climate is the effect such emplacement of mafic lithologies could have on increasing global weatherability and driving cooling.
 In particular, the emplacement of LIPs in the tropics has been hypothesized to be associated with specific episodes of climatic cooling on Earth.
@@
 
-This study combines reconstructions of the original surface extent of LIPs through the Phanerozoic, a paleogeographic model, and a parameterization of LIP erosion to estimate LIP area in all latitudinal bands through the Phanerozoic.
+This study combines reconstructions of the original surface extent and emplacement ages of LIPs through the Phanerozoic, a paleogeographic model, and a parameterization of LIP erosion to estimate LIP area in all latitudinal bands through the Phanerozoic.
 We then compare these LIP area time series to the latitudinal extent of continental ice sheets - a proxy for Earth's long term climate state.
@@
 However, despite these uncertainties, this approach likely provides the best estimates available of original extent for ancient LIPs.
-The extents of presently-exposed volcanics associated with LIPs that were used for present-day area estimates, and the construction of the original extent polygons by \citet{Ernst2017a} and \citet{Ernst2019a}, were taken from a number of resources including the PLATES compilation \citep{Coffin2006a} and more recent compilation efforts associated with the LIPs Reconstruction Project (\citealp{Ernst2013a}; Table \ref{tab:LIPs}).
+The extents of presently-exposed volcanics associated with LIPs that were used for present-day area estimates, and the sources that went into the construction of the original extent polygons by \citet{Ernst2017a} and \citet{Ernst2019a}, were taken from a number of resources including the PLATES compilation \citep{Coffin2006a} and more recent compilation efforts associated with the LIPs Reconstruction Project (\citealp{Ernst2013a}; Table \ref{tab:LIPs}).
 
@@
 The best-fit exponential function results in a LIP area half-life of 29~Myr.
-However, since we explicitly account for LIP burial separately in our models (see below), we prefer to exclude from our estimation of a representative LIP area half-life the 6 of 19 LIPs which have been partially or completely buried.
+However, since we explicitly account for LIP burial separately in our area analysis (see below), we prefer to exclude from our estimation of a representative LIP area half-life the 6 of 19 LIPs which are inferred to have been partially or completely buried.
 This latter approach yields a slightly higher best-fit half-life of 36~Myr (Fig.
@@
 it yields a low root mean square error of 0.14), if each LIP is fit individually with an exponential decay function, there is variability in the estimated half-lives such that they range from shorter half-lives of $\sim$20~Myr up to $\sim$120~Myr for the Kalkarindji and Deccan Traps (Table \ref{tab:LIPs}).
-In the analysis of LIP area through time, we implement decay scenarios informed by this analysis.
-Scenario `t$_{1/2}$ = 36~Myr' uses the best-fit half-life of 36~Myr, while `t$_{1/2}$ = 120~Myr' uses the slower decay.
+In the analysis of LIP area through time, we implement decay scenarios informed by this analysis: the `t$_{1/2}$ = 36~Myr' scenario uses the best-fit half-life of 36~Myr, while the `t$_{1/2}$ = 120~Myr' scenario uses the slower decay.
 Given that the post-emplacement weathering and erosional history of each LIP should be dependent on the tectonic and climatic setting that each LIP experiences during and after emplacement, this approach is simplistic but provides a framework for analysis.
 The LIP reconstructions used in this study also include pre-Phanerozoic LIPs.
-However, given the imposed exponential decay since emplacement, the inclusion of these LIPs does not significantly influence the calculated LIP areas throughout the Phanerozoic, which is the focus of this analysis.
+However, given the imposed exponential decay since emplacement, the inclusion of these LIPs does not significantly influence the calculated LIP areas through the Phanerozoic, which is the focus of this analysis.
 
-Of the tectonic factors that could alter exposed LIP area, the most consequential is near-immediate burial by sediments of LIP volcanics that are co-located with a rift basin.
+Of the tectonic factors that could alter exposed LIP area, the most consequential is near-immediate burial of LIP volcanics that are co-located with a rift basin by sediments.
 There are numerous examples in the record where there is partial or complete burial of a LIP associated with rifting and thermal subsidence (Table \ref{tab:LIPs}).
@@
 The `50\% burial' scenario imposes instantaneous burial of 50$\%$ of the LIP area while the `100\% burial' imposes instantaneous burial of the entire LIP as an end-member scenario.
-A limitation of our treatment of LIPs that are co-located with rifting is that we instantaneously `bury' all of these LIPs to the same degree, when in fact the degree/timing of burial of these LIPs may vary substantially.
+A limitation of our treatment of LIPs that are co-located with rifting is that we `bury' all of these LIPs instantly at the time of emplacement and to the same degree, when in fact the degree/timing of burial of these LIPs may vary substantially.
 The LIP area analysis uses all of the available combinations of the distinct decay and burial scenarios described above.
@@
 The original extent LIP polygons were assigned a plate ID corresponding to a tectonic unit on Earth using the polygons of \citet{Torsvik2016a} for the Phanerozoic.
-The LIP polygons and tectonic units were reconstructed from 520~Ma to the present utilizing the paleogeographic model of \citet{Torsvik2016a} in the spin axis reference frame (anchor plate ID of 1).
+The LIP polygons and tectonic units were reconstructed from 520~Ma to the present (e.g.
+Fig.
+\ref{fig:Reconstruction_Snapshots}) utilizing the paleogeographic model of \citet{Torsvik2016a} in the spin axis reference frame (anchor plate ID of 1).
 This paleogeographic model was updated to include revisions to Ordovician Laurentia \citep{Swanson-Hysell2017a} and Paleozoic Asia \citep{Domeier2018a}.
@@
 We use 15\textdegree~S to 15\textdegree~N as the definition of the tropical rain belt, as those latitudes approximately correspond with a sharp increase in zonal mean precipitation to values greater than 1.0~m/yr in modern climatalogical data \citep{Kalnay1996a}.
-In \citet{Evans2006a}, the reconstructed paleolatitudes of basins with thick, basin-wide evaporite deposition are shown to be consistently in the subtropics throughout the Phanerozoic (and potentially beyond), suggesting the stability of the large-scale atmospheric circulation that gives rise to intense precipitation in the tropical rain belt.
+In \citet{Evans2006a}, the reconstructed paleolatitudes of basins with thick, basin-wide evaporite deposition are shown to be consistently in the subtropics throughout the Phanerozoic (and potentially beyond), suggesting that the large-scale atmospheric circulation that gives rise to intense precipitation in the tropical rain belt is stable through time.
 However, subsequent work by \citet{Boucot2013a} and \citet{Cao2018a} has shown that evaporite deposits did form at or near the equator at times in the Phanerozoic.
@@
 to nodular evaporitic pseudomorphs).
-Such limited evaporitic mineral precipitation could be attributed to local processes such as periodic restriction of the basin that lead to saturation states that otherwise would not be expected for that latitude.
+Such limited evaporitic mineral precipitation could be attributed to local processes such as periodic restriction of the basin that led to saturation states that otherwise would not be expected for that latitude.
 Nevertheless, a limitation of the LIP analysis described in this study is that it does not account for deviations in local precipitation from the zonal mean (due to the infeasibility of running a global climate model at each time-step in the analysis).
-However, evaporite deposits are distributed bimodally about the equator in the subtropics for the vast majority of the past $\sim$420~Myr \citep{Cao2018a} and overall stability of the large-scale atmospheric circulation is predicted by climate dynamics \citep{Donohoe2017a}.
+However, evaporite deposits, including those in which the occurrence of evaporitic minerals is limited, are distributed bimodally about the equator in the subtropics for the vast majority of the past $\sim$420~Myr \citep{Cao2018a} and overall stability of the large-scale atmospheric circulation is predicted by climate dynamics \citep{Donohoe2017a}.
 Therefore, the assumption of enhanced precipitation in the tropics throughout the Phanerozoic appears to be warranted to first-order.
-To test the sensitivity of our results to the width of the tropical rain belt, we performed the tropical LIP area calculations with tropical rain belt widths of $\pm$10\textdegree and $\pm$20\textdegree, which yields similar results to a tropical rain belt width of $\pm$15\textdegree (Table \ref{tab:stats}).
+To test the sensitivity of our results to the width of the tropical rain belt, we performed the tropical LIP area calculations with tropical rain belt widths of $\pm$10\textdegree\xspace and $\pm$20\textdegree, which yields similar results to a tropical rain belt width of $\pm$15\textdegree\xspace (Table \ref{tab:stats}).
 
-To evaluate the relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic (\citealp{Macdonald2019a}; Fig.
+To evaluate the relationship between Earth's climate state and total and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic (\citealp{Macdonald2019a}; Fig.
 \ref{fig:LIP_Areas}E).
@@
 We take two approaches for comparison between the LIP area reconstruction and the record of ice extent.
-The first is to analyze the correlation between LIP area and the extent of ice away from the pole (Fig.
-\ref{fig:LIP_correlation}).
+The first is to calculate the Pearson correlation coefficient between LIP area and the extent of ice away from the pole.
 The second is to consider the overlap between intervals of high LIP area (defined as LIP area \textgreater30\% of the maximum in a given post-emplacement model) and intervals of glacial climate (defined as ice extent \textgreater10\textdegree\xspace from the poles).
@@
 
-To assess the statistical significance of the correlation (or lack thereof) implied by the Pearson correlation coefficients, we applied the approach of \citet{Macdonald2019a} and simulated the four glacial episodes (Fig.
+To assess the statistical significance of the correlation implied by the Pearson correlation coefficients (or lack thereof), we applied the approach of \citet{Macdonald2019a} and simulated the four glacial episodes (Fig.
 \ref{fig:LIP_Areas}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and \% overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these 100,000 simulations (Fig.
@@
 This approach accounts for the reality that spurious correlation can arise between auto-correlated data sets such as these, where each value is not independent, but is instead dependent on the previous state of the system.
-For the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario, 28\% of the randomly timed glacial interval simulations correlate better with the LIP area in the tropics.
-With an associated p-value of 0.28, the null hypothesis that glacial intervals do not correlate to LIP area in the tropics cannot be rejected.
-Taking this approach, none of the positive or negative correlations that emerge between the LIP area scenarios and the ice extent record are statistically significant.
+For the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario, 72\% of the randomly timed glacial interval simulations correlate better with LIP area in the tropics.
+With an associated p-value of 0.72, the null hypothesis that glacial intervals do not correlate to LIP area in the tropics cannot be rejected.
+Taking this approach, none of the positive or negative correlations that emerge between the LIP area scenarios and the ice extent record are statistically significant (Table \ref{tab:stats}).
 
@@
 However, such an approach neglects the effects of soil shielding and regolith development in low-relief regions.
-Recent progress on understanding the relationships between landscapes, topography, and chemical weathering reveals that these effects are quite important \citep{Gabet2009a, Maher2014a}.
-As a result, more recent modeling of chemical weathering that has incorporated such processes highlights the importance of high-relief regions relative to low-relief ones for setting global weatherability \citep{West2012a, Godderis2017b}.
-Without active uplift, soil shielding from regolith development on low-relief LIPs could significantly decrease the local weatherability of a LIP and mute its impact on global weatherability.
-Furthermore, significant soil shielding can lead to a supply-limited weathering regime in which the weathering rate of the underlying bedrock becomes insensitive to kinetic and equilibrium factors such as temperature and runoff (factors that would, in the absence of soil shielding, lead to relatively high weathering rates in the tropical rain belt).
+Recent progress on understanding the relationships between landscapes, topography, and chemical weathering reveals that these effects are important \citep{Gabet2009a, Maher2014a}.
+Given enough of it, soil shielding can lead to a supply-limited weathering regime in which the weathering rate of the underlying bedrock becomes insensitive to kinetic and equilibrium factors such as temperature and runoff - factors that would, in the absence of soil shielding, lead to relatively high weathering rates in the tropical rain belt.
+As a result, more recent modeling of chemical weathering incorporates such processes and highlights the importance of high-relief regions relative to low-relief ones for setting global weatherability \citep{West2012a, Godderis2017b}.
+LIPs are often emplaced in low-relief areas, and as such, without active uplift, soil shielding from regolith development on these low-relief LIPs could significantly decrease the local weatherability of a LIP and mute its impact on global weatherability.
+In this way, soil shielding could explain the lack of correlation between tropical LIP area and ice extent (Figs.
+\ref{fig:LIP_Areas} and \ref{fig:LIP_correlation}).
 In contrast, processes that lead to continued exhumation of mafic lithologies and the creation of steep topography that minimizes soil shielding, particularly in tropical regions, may exert a strong control on global weatherability and long-term climate.
@@
 
-A complication with the interpretation of soil shielding and limited weathering of LIPs is the rapid area decay rate inferred from the comparison of current LIP extent to estimated original extent (Fig.
+A complication with the interpretation of soil shielding and limited weathering of LIPs is the rapid area decay rate (t$_{1/2}$ = 36~Myr) inferred from the comparison of current LIP extent to estimated original extent (Fig.
 \ref{fig:LIP_preservation}).
@@
 We have focused this analysis on the Phanerozoic record given that well-constrained paleogeographic models are available for the past $\sim$520~Myr.
-The approach of seeking to evaluate correlation between LIP area and glaciation is further complicated for Neoproterozoic Snowball Earth events because ice-albedo runaway leads to persistent global glaciation on timescales of tens of millions of years absent of continued forcing through normal carbon cycle processes until sufficient CO$_2$ to drive deglaciation builds up in the absence of silicate weathering \citep{Hoffman2017a}.
+The approach of seeking to evaluate correlation between LIP area and glaciation is further complicated for Neoproterozoic Snowball Earth events because ice-albedo runaway leads to persistent global glaciation on timescales of tens of millions of years without continued forcing through normal carbon cycle processes until sufficient CO$_2$ to drive deglaciation builds up in the absence of silicate weathering \citep{Hoffman2017a}.
 Moreover, cooling past the critical threshold for rapid global glaciation may have occurred on a sub-million year timescale (e.g.
 \citealp{Macdonald2017a}).
-Nevertheless, the hypothesis of tropical LIP area associated with the ca.
+Nevertheless, evaluating the hypothesis of tropical LIP area associated with the ca.
 720~Ma Franklin LIP increasing global weatherability and contributing to the onset of the Sturtian Snowball Earth is a major motivating driver behind conducting this analysis.
@@
 
-Together, these results indicate that the tropical area of the Franklin LIP, when compared to that of Phanerozoic as well as other Precambrian LIPs, did not have a singularly large area in the tropics.
-Given that similar peaks in tropical LIP area are not associated with the onset of glacial periods, additional processes beyond an increase in weatherability due to LIP area in the tropics must have been at play in the initiation of the Sturtian Snowball Earth.
+Together, these results indicate that the Franklin LIP, when compared to Phanerozoic as well as other Precambrian LIPs, did not have a uniquely large area in the tropics.
+Given that similar (and larger) peaks in tropical LIP area are not associated with the onset of glacial periods, additional processes beyond an increase in weatherability due to LIP area in the tropics must have been at play in the initiation of the Sturtian Snowball Earth.
 One such process could have been an increase in global weatherability associated with the collision and accretion of arc terranes within the present-day Arabian-Nubian Shield \citep{Park2018a}.
@@
 
-The temporal overlap between Franklin LIP eruptions and the initiation of Sturtian glaciation remains compelling \citep{Macdonald2010a, MacLennan2018a}.
+However, the temporal overlap between Franklin LIP eruptions and the initiation of Sturtian glaciation remains compelling \citep{Macdonald2010a, MacLennan2018a}.
 This overlap continues to suggest that other aspects of LIP emplacement, such as the injection of sulfur aerosols in the stratosphere \citep{Macdonald2017a}, could have played a role in the initiation of low-latitude glaciation.
-The latitude and background climate of such aerosol injections is important for their temporary effect on albedo which is maximized at low latitude, and their presence at high concentrations is pre-conditioned on eruption through sedimentary basins hosting evaporite deposits \citep{Macdonald2017a}.
-Sulfur aerosols associated with low-latitude LIPs may have also caused short-term cooling due to changes in albedo during the Phanerozoic, as has been suggested for the CAMP \citep{Schoene2010a}.
-However, because these cooling events did not result in a Snowball Earth, the climate would have returned to its background climate state within thousands of years \citep{Macdonald2017a}, which would not be observed in our analysis.
+The temporary effect on albedo of such aerosols is maximized when they are injected into the atmosphere at low-latitudes into a cool background climate, and their presence at high concentrations is pre-conditioned on eruption through sedimentary basins hosting evaporite deposits \citep{Macdonald2017a}.
+Changes in albedo due to sulfur aerosols associated with low-latitude LIPs may have also caused short-term cooling due to changes in albedo during the Phanerozoic, as has been suggested for the CAMP \citep{Schoene2010a}.
+However, because these cooling events did not result in ice-albedo runaway and a Snowball Earth, the climate would have returned to its background climate state within thousands of years \citep{Macdonald2017a}, which would not be observed in our analysis.
 
-The results from this analysis indicate that when the entire LIP database is considered in this way, there is no significant relationship between total LIP area or LIP area in the tropics and the extent of continental ice sheets (Fig.
-\ref{fig:LIP_correlation}).
+The results from this analysis indicate that when the entire LIP database is considered in this way, there is no significant relationship between total LIP area or LIP area in the tropics and the extent of continental ice sheets.
 While this result need not imply that there is no increase in global weatherability from the emplacement of LIPs, it does suggest that changes in planetary weatherability associated with LIPs are not the fundamental control on whether Earth is in a glacial or non-glacial climate state.
@@
 Richard Ernst provided GIS compilations of LIP extent and present-day exposure that were essential to the analysis.
-Discussions with Yves Godd\'eris made possible through the France Berkeley Fund were valuable in contributing to aspects of the interpretation.
+Discussions with Yves Godd\'eris made possible through the France-Berkeley Fund contributed valuably to aspects of the interpretation.
 Park was supported by NSF Grant 1547434 awarded to Swanson-Hysell.
@@
 the p-value with respect to the null hypothesis of no correlation/overlap).
+P-values \textless0.05 indicate that we can reject the null hypothesis at the 95\% confidence level.
 \vspace{0.15cm}
@@
 	\includegraphics[width=\textwidth]{Manuscript/Figures/LIP_Map.pdf}
-	\caption{Map of current extent of volcanic lithologies associated with LIPs that erupted between 520~Ma and the present, as well as the estimates of the initial LIP area used in the area analysis (modified slightly from \citealt{Ernst2017a} and \citealt{Ernst2019a} to ensure that all currently exposed volcanic lithologies are encapsulated by the initial LIP area polygons).}
+	\caption{Map of current extent of volcanic lithologies associated with LIPs that erupted between 520~Ma and the present, as well as the estimates of the initial LIP extent (modified slightly from \citealt{Ernst2017a} and \citealt{Ernst2019a} to ensure that all currently exposed volcanic lithologies are encapsulated by the initial LIP area polygons) used in the area analysis.}
 	\label{fig:LIP_map}
@@
 	\includegraphics[width=0.6\textwidth]{Manuscript/Figures/LIP_Preservation.pdf}
-	\caption{LIP preservation through time.
+	\caption{LIP erosion through time.
 The ratio of estimates of the present-day area to that of the original area are shown for 19 basaltic LIPs.
@@
 = Caribbean-Colombian.
-\textbf{B)} Total LIP area through time for the different post-emplacent models.
-Only the `no decay' model excludes pre-Phanerozoic LIPs.
-\textbf{C)} Tropical LIP area through time for the different post-emplacement models.
+\textbf{B)} Total LIP area through time for the different post-emplacent scenarios.
+Only the `no decay' scenario excludes pre-Phanerozoic LIPs.
+\textbf{C)} Tropical LIP area through time for the different post-emplacement scenarios.
 The arrows to the right indicate reconstructed tropical LIP area at the time of emplacement for the ca.
@@
 \ref{fig:Reconstruction_Snapshots}.
-\textbf{D)} Heat map showing the latitudinal distribution of LIP area for the preferred post-emplacement model.
-\textbf{E)} Latitudinal extent of land ice away from the poles through time \citep{Macdonald2019a} and compilation of \textit{p}CO$_{2}$ proxies \citep{Foster2017a} (\textit{p}CO$_{2}$ y-axis reversed, and in log-scale).
-Error bars indicate standardized uncertainties, and grey bands indicate 68 and 95\% confidence intervals for Monte Carlo resampled LOESS fits through the \textit{p}CO$_{2}$ proxy data.}
+\textbf{D)} Contour plot showing the latitudinal distribution of LIP area for the preferred post-emplacement model.
+\textbf{E)} Latitudinal extent of land ice away from the poles \citep{Macdonald2019a} and compilation of \textit{p}CO$_{2}$ proxies \citep{Foster2017a} (\textit{p}CO$_{2}$ y-axis reversed, and in log-scale).
+Error bars indicate standardized uncertainties, and grey bands indicate 68 and 95\% confidence intervals for Monte Carlo resampled LOESS fits through the \textit{p}CO$_{2}$ proxy data \citep{Foster2017a}.}
 	\label{fig:LIP_Areas}
@@
 \ref{fig:LIP_Areas}C).
-The opacity of LIP polygons indicates their interpreted remainingu area at the time of the reconstruction as a percentage of initial LIP area, under the preferred post-emplacement model of `t$_{1/2}$ = 36~Myr + 50\% burial'.}
+The opacity of LIP polygons indicates their parameterized remaining area at the time of the reconstruction as a percentage of initial LIP area, under the preferred post-emplacement scenario of `t$_{1/2}$ = 36~Myr + 50\% burial'.}
 	\label{fig:Reconstruction_Snapshots}
@@
 	\includegraphics[width=0.85\textwidth]{Manuscript/Figures/overlap_correlation_cropped.pdf}
-	\caption{The \% overlap and Pearson correlation coefficients for various LIP post-emplacement scenarios are shown with circles.
-These values are compared to histograms that show the range of values that arise when comparing the LIP area record to glacial intervals that have been shifted randomly in time 100,000 times.}
+	\caption{The \% overlap and Pearson correlation coefficients between LIP area and the actual ice-extent record are shown with circles.
+These values are compared to histograms that show the range of values that arise when comparing the LIP area record to glacial intervals that have been shifted randomly in time 100,000 times.
+The fraction of randomly timed glacial interval simulations that correlate/overlap better with LIP area than the actual ice-extent record is the p-value shown in Table \ref{tab:stats}.}
 	\label{fig:LIP_correlation}

### 2019-04-22 Manuscript/Manuscript.tex

@@
 
-One of the hypothesized effects of the emplacement of large igneous provinces (LIPs) is planetary cooling on million-year timescales associated with enhanced silicate weathering of the highly weatherable basalt.
+One of the hypothesized effects of the emplacement of large igneous provinces (LIPs) is planetary cooling on million-year timescales associated with enhanced silicate weathering of the freshly-emplaced basalt.
 This study combines reconstructions of the original surface extent and emplacement ages of LIPs through the Phanerozoic, a paleogeographic model, and a paramaterization of LIP erosion to estimate LIP area in all latitudinal bands through the Phanerozoic.
@@
 
-Global weatherability is the sum of factors aside from climate itself that contribute to overall global weathering and associated CO$_2$ consumption, such as the latitudinal distribution of continents and mountain belts.
+Global weatherability is the sum of factors aside from climate itself that contribute to overall global weathering and associated CO$_2$ consumption, such as the latitudinal distribution of continents and mountain belts \citep{Kump1997a}.
 On a planet with high weatherability, the CO$_2$ input from volcanism can be removed via silicate weathering at a lower atmospheric CO$_2$ concentration than on a less weatherable planet.
@@
 720~Ma Franklin LIP in the tropics, in concert with elevated runoff rates associated with continental break-up, has been implicated as a major contributor to the cooling that initiated the Sturtian `Snowball Earth' \citep{Donnadieu2004b, Macdonald2010a, Cox2016a}.
-In the Cenozoic, the movement of the Deccan LIP into the tropical rain belt, in concert with the low-latitude emplacement of the Ethiopian Traps, has been implicated in drawing down CO$_2$ levels in the lead-up to Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
-More recently, \citet{Johansson2018a} used a paleogeographic model to suggest that tropical LIP area correlates with Phanerozoic climate change through comparisons to \textit{p}CO$_{2}$ proxy records.
+In the Cenozoic, the movement of the Deccan LIP into the tropical rain belt, together with the low-latitude emplacement of the Ethiopian Traps, has been implicated in drawing down CO$_2$ levels in the lead-up to Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
+More recently, \citet{Johansson2018a} used a paleogeographic model to suggest that tropical LIP area correlates with Phanerozoic climate change through comparison with a \textit{p}CO$_{2}$ proxy compilation.
 
@@
 For some LIPs, this approach may lead to an overestimate of original surface extent, given that subsurface intrusions could extend over a broader area than the surface volcanics.
-The polygons also assume complete surface coverage between wide-spread remnants, creating further potential for these outlines being over-estimates.
+The polygons also assume complete surface coverage between wide-spread remnants, creating further potential for these original extent outlines to be over-estimates.
 However, despite these uncertainties, this approach likely provides the best estimates available of original extent for ancient LIPs.
@@
 They used the resulting exponential decay constants to develop a first-order parameterization to estimate changing LIP area through time.
-We extend this approach to 19 basaltic LIPs for which there are estimates of the original surface extent of the province and the current surface extent of rocks associated with it (Fig.
+We extend this approach to 19 basaltic LIPs for which there are estimates of the original surface extent of the province and the current surface extent of rocks associated with the province (Fig.
 \ref{fig:LIP_preservation}).
@@
 In their analysis, some of the polygons correspond to the present-day surface extent and some represent the original extent that includes currently buried portions of the LIP (e.g.
-the Keweenawan Midcontinent Rift for which the implemented extent is from geophysical data that largely corresponds to buried subsurface exposures).
+the Keweenawan Midcontinent Rift for which the implemented extent in \citet{Johansson2018a} is from geophysical data that largely corresponds to buried subsurface exposures).
 
@@
 \ref{fig:LIP_Areas}B and C).
-All the data and code necessary to reproduce the analyses and figures presented in this study can be downloaded from GitHub: \url{https://github.com/Swanson-Hysell-Group/Large_Igneous_Provinces}.
+All the data and code necessary to reproduce the analyses and figures presented in this study can be downloaded from GitHub (\url{https://github.com/Swanson-Hysell-Group/Large_Igneous_Provinces}) and are also archived in Zenodo (URL TO BE ADDED HERE).
 
 We use 15\textdegree~S to 15\textdegree~N as the definition of the tropical rain belt, as those latitudes approximately correspond with a sharp increase in zonal mean precipitation to values greater than 1.0~m/yr in modern climatalogical data \citep{Kalnay1996a}.
+Other parameters that could be used to define the tropical rain belt are runoff and evaporation minus precipitation (E--P).
+In modern climatological data, runoff sharply increases to values above 0.6~kg/m$^{2}$ between approximately 10 and 20\textdegree\xspace, and E--P sharply decreases to values below 1.5~kg/m$^{2}$/day between approximately 10 and 15\textdegree.
+Given that it is unclear as to which of these parameters correlates most strongly with long term weathering rates in natural environments, we also tested the sensitivity of our results to the assumed width of the tropical rain belt by performing the tropical LIP area calculations with tropical rain belt widths of $\pm$10\textdegree\xspace and $\pm$20\textdegree.
+Testing different tropical rain belt widths also has the indirect effect of accounting for uncertainties in the paleolatitude of LIPs in the paleogeographic model.
+We find that the $\pm$10\textdegree\xspace and $\pm$20\textdegree\xspace tropical rain belts yield similar results to that defined as $\pm$15\textdegree\xspace (Table \ref{tab:stats}).
+
 In \citet{Evans2006a}, the reconstructed paleolatitudes of basins with thick, basin-wide evaporite deposition are shown to be consistently in the subtropics throughout the Phanerozoic (and potentially beyond), suggesting that the large-scale atmospheric circulation that gives rise to intense precipitation in the tropical rain belt is stable through time.
@@
 However, evaporite deposits, including those in which the occurrence of evaporitic minerals is limited, are distributed bimodally about the equator in the subtropics for the vast majority of the past $\sim$420~Myr \citep{Cao2018a} and overall stability of the large-scale atmospheric circulation is predicted by climate dynamics \citep{Donohoe2017a}.
-Therefore, the assumption of enhanced precipitation in the tropics throughout the Phanerozoic appears to be warranted to first-order.
-To test the sensitivity of our results to the width of the tropical rain belt, we performed the tropical LIP area calculations with tropical rain belt widths of $\pm$10\textdegree\xspace and $\pm$20\textdegree, which yields similar results to a tropical rain belt width of $\pm$15\textdegree\xspace (Table \ref{tab:stats}).
+Therefore, the assumption of enhanced precipitation in the tropics throughout the Phanerozoic is warranted.
 
@@
 The land ice record is an imperfect tracker of climate, as it is insensitive to changes in temperature during non-glacial intervals, is influenced by additional factors such as the physical geography of the continents during glacial intervals, and is potentially vulnerable to removal from the observable geologic record via erosion and burial.
+Furthermore, the threshold \textit{p}CO$_2$ for establishing a glacial climate is dependent on ocean circulation and changing solar luminosity.
 Nevertheless, it forms a physical record of Earth's climate through time and delineates glacial and non-glacial climate states.
@@
 \ref{fig:LIP_Areas}E).
+For the pedogenic carbonate \dC proxy, which forms the majority of the pre-Cenozoic data in the compilation of \citet{Foster2017a}, such scatter could result from diagenesis \citep{Michel2016a} and the sensitivity of the \textit{p}CO$_2$ estimates on assumptions regarding soil respired CO$_{2}$ \citep{Montanez2013a}.
 Nevertheless, despite these shortcomings, the \textit{p}CO$_2$ proxy record is broadly consistent with the ice extent record (Fig.
-\ref{fig:LIP_Areas}E) - sparse \textit{p}CO$_2$ proxy data decreases $\sim$400-310~Ma as the Late Devonian glacial interval occurs and the Permo-Carboniferous glacial interval begins and waxes, \textit{p}CO$_2$ proxy data roughly increases $\sim$310-240~Ma as the Permo-Carboniferous glacial interval wanes and ends, \textit{p}CO$_2$ proxy data broadly remains relatively high $\sim$240-40~Ma when no glacial intervals are robustly documented, and \textit{p}CO$_2$ proxy data roughly decreases $\sim$40-0~Ma as the Cenozoic glacial interval begins.
+\ref{fig:LIP_Areas}E) - \textit{p}CO$_2$ proxy data decreases $\sim$400-310~Ma as the Late Devonian glacial interval occurs and the Permo-Carboniferous glacial interval begins and waxes, \textit{p}CO$_2$ proxy data roughly increases $\sim$310-240~Ma as the Permo-Carboniferous glacial interval wanes and ends, \textit{p}CO$_2$ proxy data broadly remains relatively high $\sim$240-40~Ma when no glacial intervals are robustly documented, and \textit{p}CO$_2$ proxy data roughly decreases $\sim$40-0~Ma as the Cenozoic glacial interval begins.
 Given these considerations, we thus prefer to use the latitudinal extent of land ice to reflect Earth's overall climate state throughout the Phanerozoic, despite its own limitations.
-
-We note, however, that Earth's climate state, regardless of whether we choose to represent it using the latitudinal extent of continental ice sheets or proxy compilations of \textit{p}CO$_2$, is dependent on a variety of factors in addition to planetary weatherability.
-For instance, the CO$_{2}$ degassing flux of the solid Earth, ocean circulation patterns, and orbital parameters all could have been significantly different in the past.
-Changes in these additional factors may have modulated the sensitivity of Earth's climate to changes in planetary weatherability.
 
@@
 
-The only scenario which results in a non-negative Pearson correlation coefficient (0.10) between LIP area in the tropics (Fig.
+The only scenario which results in a non-negative Pearson correlation coefficient (0.10) between LIP area in the tropics ($\pm$15\textdegree; Fig.
 \ref{fig:LIP_Areas}C) and the ice extent record (Fig.
@@
 This approach accounts for the reality that spurious correlation can arise between auto-correlated data sets such as these, where each value is not independent, but is instead dependent on the previous state of the system.
-For the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario, 72\% of the randomly timed glacial interval simulations correlate better with LIP area in the tropics.
+For the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario, 72\% of the randomly timed glacial interval simulations correlate better with LIP area in the tropics than the actual ice extent record.
 With an associated p-value of 0.72, the null hypothesis that glacial intervals do not correlate to LIP area in the tropics cannot be rejected.
@@
 However, such an approach neglects the effects of soil shielding and regolith development in low-relief regions.
-Recent progress on understanding the relationships between landscapes, topography, and chemical weathering reveals that these effects are important \citep{Gabet2009a, Maher2014a}.
-Given enough of it, soil shielding can lead to a supply-limited weathering regime in which the weathering rate of the underlying bedrock becomes insensitive to kinetic and equilibrium factors such as temperature and runoff - factors that would, in the absence of soil shielding, lead to relatively high weathering rates in the tropical rain belt.
+Recent progress on understanding the relationships between landscapes, topography, and chemical weathering reveals that these effects are important \citep{Gabet2009a, Maher2014a, Godderis2017b}.
+Soil shielding can lead to a transport-limited weathering regime in which the weathering rate of the underlying bedrock becomes insensitive to kinetic and equilibrium factors such as temperature and runoff - factors that would, in the absence of soil shielding, lead to relatively high weathering rates in the tropical rain belt.
 As a result, more recent modeling of chemical weathering incorporates such processes and highlights the importance of high-relief regions relative to low-relief ones for setting global weatherability \citep{West2012a, Godderis2017b}.
@@
 A low-latitude continent configuration and abundant arc-continent collisions in the tropics may have led to a cool background climate, and the emplacement of the Franklin LIP may have further increased global weatherability to the point where the ice-albedo runaway could take effect.
-However, tropical LIP area associated with the Franklin was not uniquely high, and therefore an associated increase in global weatherability was not the sole driver of Snowball Earth onset, consistent with the results of the Phanerozoic analysis.
+However, tropical LIP area associated with the Franklin was not uniquely high, and therefore an associated increase in global weatherability was likely not the sole driver of Snowball Earth onset, consistent with the results of the Phanerozoic analysis.
 
 However, the temporal overlap between Franklin LIP eruptions and the initiation of Sturtian glaciation remains compelling \citep{Macdonald2010a, MacLennan2018a}.
-This overlap continues to suggest that other aspects of LIP emplacement, such as the injection of sulfur aerosols in the stratosphere \citep{Macdonald2017a}, could have played a role in the initiation of low-latitude glaciation.
-The temporary effect on albedo of such aerosols is maximized when they are injected into the atmosphere at low-latitudes into a cool background climate, and their presence at high concentrations is pre-conditioned on eruption through sedimentary basins hosting evaporite deposits \citep{Macdonald2017a}.
-Changes in albedo due to sulfur aerosols associated with low-latitude LIPs may have also caused short-term cooling due to changes in albedo during the Phanerozoic, as has been suggested for the CAMP \citep{Schoene2010a}.
-However, because these cooling events did not result in ice-albedo runaway and a Snowball Earth, the climate would have returned to its background climate state within thousands of years \citep{Macdonald2017a}, which would not be observed in our analysis.
+This overlap could support arguments that other aspects of LIP emplacement, such as the injection of sulfur aerosols in the stratosphere \citep{Macdonald2017a}, played a role in the initiation of low-latitude glaciation.
+The temporary effect on albedo of such aerosols is maximized when they are injected into the atmosphere at low-latitudes into a cool background climate and their presence at high concentrations is pre-conditioned on eruption through sedimentary basins hosting evaporite deposits, as could have been the case for the Franklin LIP \citep{Macdonald2017a}.
+However, in the cases in which such aerosol-driven cooling does not result in ice-albedo runaway and a Snowball Earth, the climate would return to its background climate state within years \citep{Macdonald2017a}.
+Furthermore, LIP emplacement could instead cause transient warming associated with elevated CO$_{2}$ outgassing leading to transiently high \textit{p}CO$_{2}$, as has been argued for the CAMP \citep{Schaller2011a, Schaller2012a}.
 
-The results from this analysis indicate that when the entire LIP database is considered in this way, there is no significant relationship between total LIP area or LIP area in the tropics and the extent of continental ice sheets.
+The results from this analysis indicate that when the entire LIP database is considered through paleogeographic reconstruction and this parameterization of erosion, there is no significant relationship between total LIP area or LIP area in the tropics and the extent of continental ice sheets.
 While this result need not imply that there is no increase in global weatherability from the emplacement of LIPs, it does suggest that changes in planetary weatherability associated with LIPs are not the fundamental control on whether Earth is in a glacial or non-glacial climate state.
+This result points to variable CO$_{2}$ outgassing or other changes in planetary weatherability such as mountain building in the tropics as more significant drivers.
 

### 2019-04-22 Manuscript/Manuscript.tex

@@
 % Captions will be left justified
-\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off,font=small]{caption}
 
@@
 \textbf{E)} Latitudinal extent of land ice away from the poles \citep{Macdonald2019a} and compilation of \textit{p}CO$_{2}$ proxies \citep{Foster2017a} (\textit{p}CO$_{2}$ y-axis reversed, and in log-scale).
-Error bars indicate standardized uncertainties, and grey bands indicate 68 and 95\% confidence intervals for Monte Carlo resampled LOESS fits through the \textit{p}CO$_{2}$ proxy data \citep{Foster2017a}.}
+Error bars indicate standardized uncertainties, and grey bands indicate 68 and 95\% confidence intervals for Monte Carlo resampled LOESS fits to the \textit{p}CO$_{2}$ proxy data \citep{Foster2017a}.
+Note that there are \textit{p}CO$_{2}$ estimates \textless100~ppm that are cut off in this plot.}
 	\label{fig:LIP_Areas}

### 2019-04-22 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 Other parameters that could be used to define the tropical rain belt are runoff and evaporation minus precipitation (E--P).
-In modern climatological data, runoff sharply increases to values above 0.6~kg/m$^{2}$ between approximately 10 and 20\textdegree\xspace, and E--P sharply decreases to values below 1.5~kg/m$^{2}$/day between approximately 10 and 15\textdegree.
+In modern climatological data, runoff sharply increases to values above 0.6~kg/m$^{2}$ between approximately 10 and 20\textdegree\xspace \citep{Kalnay1996a}, and E--P sharply decreases to values below 1.5~kg/m$^{2}$/day between approximately 10 and 15\textdegree \citep{Trenberth2011a}.
 Given that it is unclear as to which of these parameters correlates most strongly with long term weathering rates in natural environments, we also tested the sensitivity of our results to the assumed width of the tropical rain belt by performing the tropical LIP area calculations with tropical rain belt widths of $\pm$10\textdegree\xspace and $\pm$20\textdegree.
@@
 \begin{center}
+	\includegraphics[width=0.6\textwidth]{Manuscript/Figures/Climatology.pdf}
+	\caption{Zonally averaged modern climatological data from \citet{Kalnay1996a} and \citet{Trenberth2011a} used to define the tropical rain belt.
+Precipitation and evaporation minus precipitation data are from land and ocean, whereas runoff data are from land only.
+Note that all of these metrics increase/decrease sharply between 10 and 20\textdegree\xspace.}
+	\label{fig:Climatology}
+\end{center}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
 	\includegraphics[width=0.9\textwidth]{Manuscript/Figures/LIP_Areas.pdf}

### 2019-04-22 Manuscript/Manuscript.tex
post-revisions checkpoint

--- 2019-04-22 Manuscript/Manuscript.tex
@@
 Other parameters that could be used to define the tropical rain belt are runoff and evaporation minus precipitation (E--P).
-In modern climatological data, runoff sharply increases to values above 0.6~kg/m$^{2}$ between approximately 10 and 20\textdegree\xspace, and E--P sharply decreases to values below 1.5~kg/m$^{2}$/day between approximately 10 and 15\textdegree.
-Given that it is unclear as to which of these parameters correlates most strongly with long term weathering rates in natural environments, we also tested the sensitivity of our results to the assumed width of the tropical rain belt by performing the tropical LIP area calculations with tropical rain belt widths of $\pm$10\textdegree\xspace and $\pm$20\textdegree.
+In modern climatological data, runoff sharply increases to values above 0.6~kg/m$^{2}$ between approximately $\pm$10\textdegree\xspace and $\pm$20\textdegree\xspace \citep{Kalnay1996a}, and E--P sharply decreases to values below 1.5~kg/m$^{2}$/day between approximately $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace \citep{Trenberth2011a}.
+Given that it is unclear as to which of these parameters correlates most strongly with long term weathering rates in natural environments, we also tested the sensitivity of our results to the assumed width of the tropical rain belt by performing the tropical LIP area calculations with tropical rain belt widths of $\pm$10\textdegree\xspace and $\pm$20\textdegree\xspace.
 Testing different tropical rain belt widths also has the indirect effect of accounting for uncertainties in the paleolatitude of LIPs in the paleogeographic model.
@@
 \begin{center}
+	\includegraphics[width=0.8\textwidth]{Manuscript/Figures/Climatology.pdf}
+	\caption{Zonally averaged modern climatological data used to define the tropical rain belt \citep{Kalnay1996a, Trenberth2011a}.
+Precipitation and evaporation minus precipitation data are from land and ocean, whereas runoff data are from land only.
+Note that all of these metrics increase/decrease sharply between $\pm$10\textdegree\xspace and $\pm$20\textdegree\xspace.}
+	\label{fig:Climatology}
+\end{center}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
 	\includegraphics[width=0.9\textwidth]{Manuscript/Figures/LIP_Areas.pdf}

### 2019-04-22 Manuscript/Manuscript.tex
new checkpoint post-revisions

--- 2019-04-22 Manuscript/Manuscript.tex
@@
 
-We use 15\textdegree~S to 15\textdegree~N as the definition of the tropical rain belt, as those latitudes approximately correspond with a sharp increase in zonal mean precipitation to values greater than 1.0~m/yr in modern climatalogical data \citep{Kalnay1996a}.
+We use 15\textdegree~S to 15\textdegree~N as the definition of the tropical rain belt, as those latitudes approximately correspond with a sharp increase in zonal mean precipitation to values greater than 1.0~m/yr in modern climatalogical data (\citealp{Kalnay1996a}; Fig.
+\ref{fig:Climatology}).
 Other parameters that could be used to define the tropical rain belt are runoff and evaporation minus precipitation (E--P).
-In modern climatological data, runoff sharply increases to values above 0.6~kg/m$^{2}$ between approximately 10 and 20\textdegree\xspace, and E--P sharply decreases to values below 1.5~kg/m$^{2}$/day between approximately 10 and 15\textdegree.
-Given that it is unclear as to which of these parameters correlates most strongly with long term weathering rates in natural environments, we also tested the sensitivity of our results to the assumed width of the tropical rain belt by performing the tropical LIP area calculations with tropical rain belt widths of $\pm$10\textdegree\xspace and $\pm$20\textdegree.
-Testing different tropical rain belt widths also has the indirect effect of accounting for uncertainties in the paleolatitude of LIPs in the paleogeographic model.
-We find that the $\pm$10\textdegree\xspace and $\pm$20\textdegree\xspace tropical rain belts yield similar results to that defined as $\pm$15\textdegree\xspace (Table \ref{tab:stats}).
+In modern climatological data, runoff sharply increases to values above 0.6~kg/m$^{2}$ between approximately $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace (\citealp{Kalnay1996a}; Fig.
+\ref{fig:Climatology}), and E--P sharply decreases to values below 1.5~kg/m$^{2}$/day also between approximately $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace (\citealp{Trenberth2011a}; Fig.
+\ref{fig:Climatology}).
+Given that it is unclear as to which of these parameters correlates most strongly with long term weathering rates in natural environments, we also tested the sensitivity of our results to the assumed width of the tropical rain belt by performing the tropical LIP area calculations with a tropical rain belt width of $\pm$10\textdegree\xspace.
+We also calculated the area of LIPs within $\pm$20\textdegree\xspace of the equator (a width that includes part of the arid subtropics) in order to account for uncertainties in the paleolatitude of LIPs in the paleogeographic model.
+We find that both of these additional analyses (LIP area calculated within $\pm$10\textdegree\xspace and $\pm$20\textdegree\xspace of the equator) yield similar results to that when LIP area is calculated within $\pm$15\textdegree\xspace of the equator (Table \ref{tab:stats}).
 
@@
 \begin{center}
+	\includegraphics[width=\textwidth]{Manuscript/Figures/Climatology.pdf}
+	\caption{Zonally averaged modern climatological data used to define the tropical rain belt.
+The global precipitation \citep{Kalnay1996a} and evaporation minus precipitation \citep{Trenberth2011a} data include land and ocean pixels, whereas global runoff and temperature \citep{Kalnay1996a} data are from land only.
+The peak in runoff $\sim$-50\textdegree\xspace is due to anomalously high runoff in the southern Andes, which represents almost all of the land in that latitude belt.
+Temperature data for Antarctica are off scale.
+Precipitation, runoff, and evaporation minus precipitation all increase/decrease sharply between $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace.}
+	\label{fig:Climatology}
+\end{center}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
 	\includegraphics[width=0.9\textwidth]{Manuscript/Figures/LIP_Areas.pdf}

### 2019-04-29 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 
-One of the hypothesized effects of the emplacement of large igneous provinces (LIPs) is planetary cooling on million-year timescales associated with enhanced silicate weathering of the freshly-emplaced basalt.
+One of the hypothesized effects of large igneous provinces (LIPs) is planetary cooling on million-year timescales associated with enhanced silicate weathering of the freshly-emplaced basalt.
 This study combines reconstructions of the original surface extent and emplacement ages of LIPs through the Phanerozoic, a paleogeographic model, and a paramaterization of LIP erosion to estimate LIP area in all latitudinal bands through the Phanerozoic.
@@
 On a planet with high weatherability, the CO$_2$ input from volcanism can be removed via silicate weathering at a lower atmospheric CO$_2$ concentration than on a less weatherable planet.
-Basaltic regions consume more CO$_2$ than regions where bedrock composition is closer to bulk continental crust because of the relatively high concentrations of Ca and Mg (that ultimately sequester carbon through precipitation as carbonate) and the high weathering rates of mafic lithologies \citep{Dessert2003a}.
-Furthermore, data from basaltic watersheds show that chemical weathering rates are greatest in regions with high runoff.
+Basaltic regions consume more CO$_2$ than regions where bedrock composition is closer to bulk continental crust because mafic lithologies have relatively high concentrations of Ca and Mg (that ultimately sequester carbon through precipitation as carbonate), constitute minerals with relatively high reactivity \citep{Gislason2003a}, and have relatively high weathering rates \citep{Dessert2003a, Ibarra2016a}.
+Furthermore, data from basaltic watersheds show that chemical weathering rates are greatest in regions with high runoff and temperature.
 As a result, CO$_2$ consumption in basaltic regions is most pronounced in the tropical rain belt \citep{Dessert2003a, Hartmann2009a, Hartmann2014a}.
 
-One aspect of large igneous province (LIP) emplacement that has been hypothesized to relate to long-term global climate is the effect such emplacement of mafic lithologies could have on increasing global weatherability and driving cooling.
+One aspect of large igneous province (LIP) emplacement that has been hypothesized to relate to long-term global climate is the effect that associated mafic lithologies could have on increasing global weatherability and driving cooling.
 In particular, the emplacement of LIPs in the tropics has been hypothesized to be associated with specific episodes of climatic cooling on Earth.
@@
 The LIP original surface extent compilation seeks to reconstruct the original surface extent of LIPs with the caveat that there can be significant uncertainty with doing so, particularly for older more deeply eroded LIPs.
-These polygons encapsulate all of the rocks associated with a given LIP, including dikes, sills, and layered intrusions, in addition to subaerial volcanics (Fig.
+These polygons encapsulate all of the preserved rocks associated with a given LIP, including dikes, sills, and layered intrusions, in addition to subaerial volcanics (Fig.
 \ref{fig:LIP_map}).
-For some LIPs, this approach may lead to an overestimate of original surface extent, given that subsurface intrusions could extend over a broader area than the surface volcanics.
+For some LIPs, this approach may lead to an over-estimate of original surface extent, given that subsurface intrusions could extend over a broader area than the surface volcanics.
 The polygons also assume complete surface coverage between wide-spread remnants, creating further potential for these original extent outlines to be over-estimates.
+These original extent outlines could also under-estimate the surface area for some LIPs where flows have been eroded and feeder dikes are not exposed or poorly documented.
 However, despite these uncertainties, this approach likely provides the best estimates available of original extent for ancient LIPs.
@@
 
-After LIPs are emplaced, they will progressively erode.
+After LIPs are emplaced, they progressively erode.
 In order to account for the associated decrease in area with time, \citet{Godderis2017a} took the approach of fitting an exponential decay function to estimates of the original surface extent and the current surface extent for 5 LIPs.
@@
 The best-fit exponential function results in a LIP area half-life of 29~Myr.
-However, since we explicitly account for LIP burial separately in our area analysis (see below), we prefer to exclude from our estimation of a representative LIP area half-life the 6 of 19 LIPs which are inferred to have been partially or completely buried.
+However, since we explicitly account for LIP burial separately in our area analysis (see below), we exclude from our estimation of a representative LIP area half-life the 6 of 19 LIPs which are inferred to have been partially or completely buried.
 This latter approach yields a slightly higher best-fit half-life of 36~Myr (Fig.
@@
 Although the exponential fit to the 13 unburied LIPs is good (i.e.
-it yields a low root mean square error of 0.14), if each LIP is fit individually with an exponential decay function, there is variability in the estimated half-lives such that they range from shorter half-lives of $\sim$20~Myr up to $\sim$120~Myr for the Kalkarindji and Deccan Traps (Table \ref{tab:LIPs}).
+it yields a low root mean square error of 0.14), if each LIP is fit individually with an exponential decay function, there is variability in the estimated half-lives from $\sim$20~Myr up to $\sim$120~Myr for the Deccan Traps (Table \ref{tab:LIPs}).
 In the analysis of LIP area through time, we implement decay scenarios informed by this analysis: the `t$_{1/2}$ = 36~Myr' scenario uses the best-fit half-life of 36~Myr, while the `t$_{1/2}$ = 120~Myr' scenario uses the slower decay.
@@
 
-Of the tectonic factors that could alter exposed LIP area, the most consequential is near-immediate burial of LIP volcanics that are co-located with a rift basin by sediments.
+Of the tectonic factors that could alter exposed LIP area, the most consequential is near-immediate sediment burial of LIP volcanics that are co-located with a rift basin.
 There are numerous examples in the record where there is partial or complete burial of a LIP associated with rifting and thermal subsidence (Table \ref{tab:LIPs}).
@@
 The `50\% burial' scenario imposes instantaneous burial of 50$\%$ of the LIP area while the `100\% burial' imposes instantaneous burial of the entire LIP as an end-member scenario.
-A limitation of our treatment of LIPs that are co-located with rifting is that we `bury' all of these LIPs instantly at the time of emplacement and to the same degree, when in fact the degree/timing of burial of these LIPs may vary substantially.
+A limitation of our treatment of LIPs that are co-located with rifting is that we `bury' all of these LIPs instantly at the time of emplacement and to the same degree when in fact the degree/timing of burial of these LIPs may vary substantially.
 The LIP area analysis uses all of the available combinations of the distinct decay and burial scenarios described above.
@@
 
-We use 15\textdegree~S to 15\textdegree~N as the definition of the tropical rain belt, as those latitudes approximately correspond with a sharp increase in zonal mean precipitation to values greater than 1.0~m/yr in modern climatalogical data (\citealp{Kalnay1996a}; Fig.
+We use 15\textdegree~S to 15\textdegree~N as the definition of the tropical rain belt, as those latitudes approximately correspond with a sharp increase in zonal mean precipitation when approaching the equator to values greater than 1.0~m/yr in modern climatalogical data (\citealp{Kalnay1996a}; Fig.
 \ref{fig:Climatology}).
-Other parameters that could be used to define the tropical rain belt are runoff and evaporation minus precipitation (E--P).
-In modern climatological data, runoff sharply increases to values above 0.6~kg/m$^{2}$ between approximately $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace (\citealp{Kalnay1996a}; Fig.
-\ref{fig:Climatology}), and E--P sharply decreases to values below 1.5~kg/m$^{2}$/day also between approximately $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace (\citealp{Trenberth2011a}; Fig.
+Other parameters that could be used to define the tropical rain belt are runoff and precipitation minus evaporation (P--E).
+In modern climatological data when approaching the equator, runoff sharply increases to values above 0.6~kg/m$^{2}$ between approximately $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace (\citealp{Kalnay1996a}), and P--E sharply increases to values above 1.5~kg/m$^{2}$/day also between approximately $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace (\citealp{Trenberth2011a}; Fig.
 \ref{fig:Climatology}).
-Given that it is unclear as to which of these parameters correlates most strongly with long term weathering rates in natural environments, we also tested the sensitivity of our results to the assumed width of the tropical rain belt by performing the tropical LIP area calculations with a tropical rain belt width of $\pm$10\textdegree\xspace.
+While seasonally high precipitation within $\pm$15\textdegree\xspace of the equator associated with migration of the intertropical convergence zone could be a driver of high chemical weathering, annual mean runoff is often the value that is used within parameterizations of chemical weathering (e.g.
+\citealp{West2012a}).
+Using runoff or P--E favors a definition of the tropical rain belt that is $\pm$10\textdegree\xspace rather than $\pm$15\textdegree\xspace.
+Therefore, we tested the sensitivity of our results to the assumed width of the tropical rain belt by performing the tropical LIP area calculations with a tropical rain belt width of $\pm$10\textdegree\xspace.
 We also calculated the area of LIPs within $\pm$20\textdegree\xspace of the equator (a width that includes part of the arid subtropics) in order to account for uncertainties in the paleolatitude of LIPs in the paleogeographic model.
-We find that both of these additional analyses (LIP area calculated within $\pm$10\textdegree\xspace and $\pm$20\textdegree\xspace of the equator) yield similar results to that when LIP area is calculated within $\pm$15\textdegree\xspace of the equator (Table \ref{tab:stats}).
+We find that both of these additional analyses (LIP area calculated within $\pm$10\textdegree\xspace and $\pm$20\textdegree\xspace of the equator) yield similar results to those obtained when LIP area is calculated within $\pm$15\textdegree\xspace of the equator (Table \ref{tab:stats}).
 
-In \citet{Evans2006a}, the reconstructed paleolatitudes of basins with thick, basin-wide evaporite deposition are shown to be consistently in the subtropics throughout the Phanerozoic (and potentially beyond), suggesting that the large-scale atmospheric circulation that gives rise to intense precipitation in the tropical rain belt is stable through time.
+In \citet{Evans2006a}, the reconstructed paleolatitudes of basins with thick, basin-wide evaporite deposition are shown to be consistently in the subtropics throughout the Phanerozoic and the Proterozoic, suggesting that the large-scale atmospheric circulation that gives rise to intense precipitation in the tropical rain belt associated with upwelling and arid subtropical climate associated with downwelling is stable through time.
 However, subsequent work by \citet{Boucot2013a} and \citet{Cao2018a} has shown that evaporite deposits did form at or near the equator at times in the Phanerozoic.
-While some of this variability could be attributed to waxing/waning of the width of the tropical rain belt as a whole, it is important to note that there can be large deviations in local precipitation from the zonal mean due to factors such as monsoon-related precipitation, or continentality (i.e.
+While some of this variability could be attributed to waxing/waning of the width of the tropical rain belt as a whole, it is important to note that there can be large deviations in local precipitation from the zonal mean due to factors such as monsoon-related precipitation \citep{Trenberth2000a}, or continentality (i.e.
 how dispersed or amalgamated the continents are) which can lead to aridity in continental interiors.
@@
 to nodular evaporitic pseudomorphs).
-Such limited evaporitic mineral precipitation could be attributed to local processes such as periodic restriction of the basin that led to saturation states that otherwise would not be expected for that latitude.
-Nevertheless, a limitation of the LIP analysis described in this study is that it does not account for deviations in local precipitation from the zonal mean (due to the infeasibility of running a global climate model at each time-step in the analysis).
+Such limited evaporitic mineral precipitation could be attributed to seasonal evaporation that transiently led to saturation states that otherwise would not be expected for that latitude.
+Nevertheless, a limitation of the LIP analysis described in this study is that it does not account for deviations in local precipitation from the zonal mean (due to the infeasibility of running a highly-resolved global climate model at each time-step in the analysis).
 However, evaporite deposits, including those in which the occurrence of evaporitic minerals is limited, are distributed bimodally about the equator in the subtropics for the vast majority of the past $\sim$420~Myr \citep{Cao2018a} and overall stability of the large-scale atmospheric circulation is predicted by climate dynamics \citep{Donohoe2017a}.
-Therefore, the assumption of enhanced precipitation in the tropics throughout the Phanerozoic is warranted.
+Therefore, the assumption of enhanced precipitation and runoff in the tropics throughout the Phanerozoic is warranted.
 
@@
 The land ice record is an imperfect tracker of climate, as it is insensitive to changes in temperature during non-glacial intervals, is influenced by additional factors such as the physical geography of the continents during glacial intervals, and is potentially vulnerable to removal from the observable geologic record via erosion and burial.
-Furthermore, the threshold \textit{p}CO$_2$ for establishing a glacial climate is dependent on ocean circulation and changing solar luminosity.
+Furthermore, the threshold \textit{p}CO$_2$ for establishing a glacial climate is dependent on ocean circulation and changing solar luminosity (e.g.
+\citealp{Shevenell2004a, DeConto2008a}).
 Nevertheless, it forms a physical record of Earth's climate through time and delineates glacial and non-glacial climate states.
@@
 The second is to consider the overlap between intervals of high LIP area (defined as LIP area \textgreater30\% of the maximum in a given post-emplacement model) and intervals of glacial climate (defined as ice extent \textgreater10\textdegree\xspace from the poles).
-This overlap approach places less emphasis on the specific magnitudes of the compiled ice extent and LIP area records.
+This overlap approach places less emphasis on the specific magnitudes of the peaks in the compiled ice extent and LIP area records.
 
@@
 \ref{fig:LIP_Areas}E).
-For the pedogenic carbonate \dC proxy, which forms the majority of the pre-Cenozoic data in the compilation of \citet{Foster2017a}, such scatter could result from diagenesis \citep{Michel2016a} and the sensitivity of the \textit{p}CO$_2$ estimates on assumptions regarding soil respired CO$_{2}$ \citep{Montanez2013a}.
+For the pedogenic carbonate \dC proxy, which forms the majority of the pre-Cenozoic data in the compilation of \citet{Foster2017a}, such scatter could result from diagenesis \citep{Michel2016a} and the sensitivity of the \textit{p}CO$_2$ estimates on assumptions regarding soil-respired CO$_{2}$ \citep{Montanez2013a}.
 Nevertheless, despite these shortcomings, the \textit{p}CO$_2$ proxy record is broadly consistent with the ice extent record (Fig.
@@
 380~Ma is associated with the emplacement of the Kola-Dnieper LIP, the third peak ca.
-200~Ma is associated with the emplacement of the CAMP, and the fourth peak is associated with both the ca.
+200~Ma is associated with the emplacement of the Central Atlantic Magmatic Province (CAMP), and the fourth peak is associated with both the ca.
 30~Ma emplacement of the Afar LIP as well as the earlier drift of the ca.
@@
 However, after accounting for burial, a minor area of LIPs remain in the tropics from ca.
-130~Ma onwards, due to the EQUAMP, Caribbean-Colombian, and Deccan LIPs.
+130~Ma onwards, due to the Equatorial Atlantic Magmatic Province (EQUAMP), Caribbean-Colombian, and Deccan LIPs.
 Using the longer decay half-life of 120~Myr (the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario) increases the area of LIPs in the tropics at any given time step, and has the effect of extending the duration of each peak.
@@
 However, such an approach neglects the effects of soil shielding and regolith development in low-relief regions.
-Recent progress on understanding the relationships between landscapes, topography, and chemical weathering reveals that these effects are important \citep{Gabet2009a, Maher2014a, Godderis2017b}.
+Recent progress on understanding the relationships between landscapes, topography, and chemical weathering reveals that these effects are important \citep{Gabet2009a, Hartmann2014a, Maher2014a, Godderis2017b}.
 Soil shielding can lead to a transport-limited weathering regime in which the weathering rate of the underlying bedrock becomes insensitive to kinetic and equilibrium factors such as temperature and runoff - factors that would, in the absence of soil shielding, lead to relatively high weathering rates in the tropical rain belt.
 As a result, more recent modeling of chemical weathering incorporates such processes and highlights the importance of high-relief regions relative to low-relief ones for setting global weatherability \citep{West2012a, Godderis2017b}.
-LIPs are often emplaced in low-relief areas, and as such, without active uplift, soil shielding from regolith development on these low-relief LIPs could significantly decrease the local weatherability of a LIP and mute its impact on global weatherability.
+LIPs are often emplaced in relatively low-relief areas, and as such, without active uplift, soil shielding from regolith development on these low-relief LIPs could significantly decrease the local weatherability of a LIP and mute its impact on global weatherability.
 In this way, soil shielding could explain the lack of correlation between tropical LIP area and ice extent (Figs.
@@
 A couple considerations are relevant with respect to this analysis: 1) the current extent of LIP exposure is reduced in part by volcanics being covered by unconsolidated sediments (i.e.
-regolith development itself) in a number of the provinces; 2) the current extent of LIP exposure may be incomplete and an underestimate for some of the provinces; 3) the initial LIP extents are typically poorly constrained and are likely overestimates which could be resulting in inflated interpreted decay rates; and 4) the relationship between LIP area and volume is poorly constrained.
+regolith development itself) in a number of the provinces; 2) the current extent of LIP exposure may be incomplete and an underestimate for some of the provinces; 3) the initial LIP extents are typically poorly constrained and are likely over-estimates which could be resulting in inflated interpreted decay rates; and 4) the relationship between LIP area and volume is poorly constrained.
 Future efforts that improve the LIP database, such as developing better-constrained estimates of original LIP extent, constraining burial and uplift histories, and refining the timing of eruptions associated with LIPs, will improve analyses that consider the LIP record in its entirety, such as that in this contribution.
@@
 
-However, the temporal overlap between Franklin LIP eruptions and the initiation of Sturtian glaciation remains compelling \citep{Macdonald2010a, MacLennan2018a}.
+The temporal overlap between Franklin LIP eruptions and the initiation of Sturtian glaciation remains compelling \citep{Macdonald2010a, MacLennan2018a}.
 This overlap could support arguments that other aspects of LIP emplacement, such as the injection of sulfur aerosols in the stratosphere \citep{Macdonald2017a}, played a role in the initiation of low-latitude glaciation.
@@
 However, in the cases in which such aerosol-driven cooling does not result in ice-albedo runaway and a Snowball Earth, the climate would return to its background climate state within years \citep{Macdonald2017a}.
-Furthermore, LIP emplacement could instead cause transient warming associated with elevated CO$_{2}$ outgassing leading to transiently high \textit{p}CO$_{2}$, as has been argued for the CAMP \citep{Schaller2011a, Schaller2012a}.
+A contrasting effect is that, on 1~kyr to 1~Myr timescales, LIP emplacement could instead cause transient warming associated with elevated CO$_{2}$ outgassing leading to transiently high \textit{p}CO$_{2}$, as has been argued for the CAMP \citep{Schaller2011a, Schaller2012a}.
 
@@
 	\includegraphics[width=\textwidth]{Manuscript/Figures/LIP_Map.pdf}
-	\caption{Map of current extent of volcanic lithologies associated with LIPs that erupted between 520~Ma and the present, as well as the estimates of the initial LIP extent (modified slightly from \citealt{Ernst2017a} and \citealt{Ernst2019a} to ensure that all currently exposed volcanic lithologies are encapsulated by the initial LIP area polygons) used in the area analysis.}
+	\caption{Map of current extent of volcanic lithologies associated with LIPs that erupted between 520~Ma and the present, as well as the estimates of the initial LIP extent used in the area analysis (modified slightly from \citealt{Ernst2017a} and \citealt{Ernst2019a} to ensure that all currently exposed volcanic lithologies are encapsulated by the initial LIP area polygons).}
 	\label{fig:LIP_map}
@@
 	\caption{Zonally averaged modern climatological data used to define the tropical rain belt.
-The global precipitation \citep{Kalnay1996a} and evaporation minus precipitation \citep{Trenberth2011a} data include land and ocean pixels, whereas global runoff and temperature \citep{Kalnay1996a} data are from land only.
-The peak in runoff $\sim$-50\textdegree\xspace is due to anomalously high runoff in the southern Andes, which represents almost all of the land in that latitude belt.
+The global precipitation \citep{Kalnay1996a} and evaporation minus precipitation \citep{Trenberth2011a} data include land and ocean pixels, whereas global temperature \citep{Kalnay1996a} data are from land only.
 Temperature data for Antarctica are off scale.
-Precipitation, runoff, and evaporation minus precipitation all increase/decrease sharply between $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace.}
+Precipitation and precipitation minus evaporation both increase sharply between $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace.}
 	\label{fig:Climatology}
@@
 Error bars indicate standardized uncertainties, and grey bands indicate 68 and 95\% confidence intervals for Monte Carlo resampled LOESS fits to the \textit{p}CO$_{2}$ proxy data \citep{Foster2017a}.
-Note that there are \textit{p}CO$_{2}$ estimates \textless100~ppm that are cut off in this plot.}
+Note that there are \textit{p}CO$_{2}$ proxy estimates \textless100~ppm that are cut off in this plot.}
 	\label{fig:LIP_Areas}

### 2019-04-29 Manuscript/Manuscript.tex
incorporate v4 edits

@@
 
-Another approach would be to compare LIP area records to proxy compilations of \textit{p}CO$_2$ (as done by \citealt{Johansson2018a}).
+Another approach would be to compare LIP area records to proxy compilations of \textit{p}CO$_2$ (as done by \citealp{Johansson2018a}).
 However, such \textit{p}CO$_2$ proxies are potentially problematic as they can be difficult to calibrate in deep time and can be affected by secondary alteration.
@@
 	\includegraphics[width=\textwidth]{Manuscript/Figures/LIP_Map.pdf}
-	\caption{Map of current extent of volcanic lithologies associated with LIPs that erupted between 520~Ma and the present, as well as the estimates of the initial LIP extent used in the area analysis (modified slightly from \citealt{Ernst2017a} and \citealt{Ernst2019a} to ensure that all currently exposed volcanic lithologies are encapsulated by the initial LIP area polygons).}
+	\caption{Map of current extent of volcanic lithologies associated with LIPs that erupted between 520~Ma and the present, as well as the estimates of the initial LIP extent used in the area analysis (modified slightly from \citealp{Ernst2017a} and \citealp{Ernst2019a} to ensure that all currently exposed volcanic lithologies are encapsulated by the initial LIP area polygons).}
 	\label{fig:LIP_map}

### 2019-04-30 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 One of the hypothesized effects of large igneous provinces (LIPs) is planetary cooling on million-year timescales associated with enhanced silicate weathering of the freshly-emplaced basalt.
-This study combines reconstructions of the original surface extent and emplacement ages of LIPs through the Phanerozoic, a paleogeographic model, and a paramaterization of LIP erosion to estimate LIP area in all latitudinal bands through the Phanerozoic.
+This study combines reconstructions of the original surface extent and emplacement ages of LIPs, a paleogeographic model, and a paramaterization of LIP erosion to estimate LIP area in all latitudinal bands through the Phanerozoic.
 This analysis reveals no significant correlation between total LIP area, nor LIP area in the tropics, and the extent of continental ice sheets.
@@
 On a planet with high weatherability, the CO$_2$ input from volcanism can be removed via silicate weathering at a lower atmospheric CO$_2$ concentration than on a less weatherable planet.
-Basaltic regions consume more CO$_2$ than regions where bedrock composition is closer to bulk continental crust because mafic lithologies have relatively high concentrations of Ca and Mg (that ultimately sequester carbon through precipitation as carbonate), constitute minerals with relatively high reactivity \citep{Gislason2003a}, and have relatively high weathering rates \citep{Dessert2003a, Ibarra2016a}.
-Furthermore, data from basaltic watersheds show that chemical weathering rates are greatest in regions with high runoff and temperature.
+Basaltic regions consume more CO$_2$ than regions where the bedrock composition is closer to bulk continental crust because mafic lithologies have relatively high concentrations of Ca and Mg (that ultimately sequester carbon through precipitation as carbonate), constitute minerals with relatively high reactivity \citep{Gislason2003a}, and have relatively high weathering rates \citep{Dessert2003a, Ibarra2016a}.
+Furthermore, data from basaltic watersheds show that chemical weathering rates are highest in regions with high runoff and temperature.
 As a result, CO$_2$ consumption in basaltic regions is most pronounced in the tropical rain belt \citep{Dessert2003a, Hartmann2009a, Hartmann2014a}.
@@
 In particular, the emplacement of LIPs in the tropics has been hypothesized to be associated with specific episodes of climatic cooling on Earth.
-In the Neoproterozoic Era, the emplacement of the ca.
-720~Ma Franklin LIP in the tropics, in concert with elevated runoff rates associated with continental break-up, has been implicated as a major contributor to the cooling that initiated the Sturtian `Snowball Earth' \citep{Donnadieu2004b, Macdonald2010a, Cox2016a}.
+In the Neoproterozoic, the emplacement of the ca.
+720~Ma Franklin LIP in the tropics, in concert with elevated runoff rates associated with supercontinent break-up, has been implicated as a major contributor to the cooling that initiated the Sturtian `Snowball Earth' \citep{Donnadieu2004b, Macdonald2010a, Cox2016a}.
 In the Cenozoic, the movement of the Deccan LIP into the tropical rain belt, together with the low-latitude emplacement of the Ethiopian Traps, has been implicated in drawing down CO$_2$ levels in the lead-up to Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
@@
 
-This study combines reconstructions of the original surface extent and emplacement ages of LIPs through the Phanerozoic, a paleogeographic model, and a parameterization of LIP erosion to estimate LIP area in all latitudinal bands through the Phanerozoic.
+This study combines reconstructions of the original surface extent and emplacement ages of LIPs, a paleogeographic model, and a parameterization of LIP erosion to estimate LIP area in all latitudinal bands through the Phanerozoic.
 We then compare these LIP area time series to the latitudinal extent of continental ice sheets - a proxy for Earth's long term climate state.
@@
 However, despite these uncertainties, this approach likely provides the best estimates available of original extent for ancient LIPs.
-The extents of presently-exposed volcanics associated with LIPs that were used for present-day area estimates, and the sources that went into the construction of the original extent polygons by \citet{Ernst2017a} and \citet{Ernst2019a}, were taken from a number of resources including the PLATES compilation \citep{Coffin2006a} and more recent compilation efforts associated with the LIPs Reconstruction Project (\citealp{Ernst2013a}; Table \ref{tab:LIPs}).
+The extents of presently-exposed volcanics associated with LIPs that were used for present-day area estimates (Fig.
+\ref{fig:LIP_map}), and the sources that went into the construction of the original extent polygons by \citet{Ernst2017a} and \citet{Ernst2019a}, were taken from a number of resources including the PLATES compilation \citep{Coffin2006a} and more recent compilation efforts associated with the LIPs Reconstruction Project (\citealp{Ernst2013a}; Table \ref{tab:LIPs}).
 
@@
 However, since we explicitly account for LIP burial separately in our area analysis (see below), we exclude from our estimation of a representative LIP area half-life the 6 of 19 LIPs which are inferred to have been partially or completely buried.
-This latter approach yields a slightly higher best-fit half-life of 36~Myr (Fig.
+This latter approach yields a slightly longer best-fit half-life of 36~Myr (Fig.
 \ref{fig:LIP_preservation}).
@@
 In their analysis, some of the polygons correspond to the present-day surface extent and some represent the original extent that includes currently buried portions of the LIP (e.g.
-the Keweenawan Midcontinent Rift for which the implemented extent in \citet{Johansson2018a} is from geophysical data that largely corresponds to buried subsurface exposures).
+the Keweenawan Midcontinent Rift for which the implemented extent in \citealp{Johansson2018a} is from geophysical data that largely corresponds to buried subsurface exposures).
 
@@
 Other parameters that could be used to define the tropical rain belt are runoff and precipitation minus evaporation (P--E).
-In modern climatological data when approaching the equator, runoff sharply increases to values above 0.6~kg/m$^{2}$ between approximately $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace (\citealp{Kalnay1996a}), and P--E sharply increases to values above 1.5~kg/m$^{2}$/day also between approximately $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace (\citealp{Trenberth2011a}; Fig.
+When approaching the equator in modern climatological data, zonal mean runoff sharply increases to values above 0.6~kg/m$^{2}$ between approximately $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace (\citealp{Kalnay1996a}), and zonal mean P--E sharply increases to values above 1.5~kg/m$^{2}$/day also between approximately $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace (\citealp{Trenberth2011a}; Fig.
 \ref{fig:Climatology}).
@@
 
-In \citet{Evans2006a}, the reconstructed paleolatitudes of basins with thick, basin-wide evaporite deposition are shown to be consistently in the subtropics throughout the Phanerozoic and the Proterozoic, suggesting that the large-scale atmospheric circulation that gives rise to intense precipitation in the tropical rain belt associated with upwelling and arid subtropical climate associated with downwelling is stable through time.
+In \citet{Evans2006a}, the reconstructed paleolatitudes of basins with thick, basin-wide evaporite deposition are shown to be consistently in the subtropics throughout the Phanerozoic and the Proterozoic, suggesting that the large-scale atmospheric circulation that gives rise to intense precipitation in the tropical rain belt and an arid subtropical climate is stable through time.
 However, subsequent work by \citet{Boucot2013a} and \citet{Cao2018a} has shown that evaporite deposits did form at or near the equator at times in the Phanerozoic.
@@
 how dispersed or amalgamated the continents are) which can lead to aridity in continental interiors.
-For instance, much of the Early Cretaceous low-latitude evaporite deposits were formed in basins which were located deep within arid continental interiors at the time of deposition \citep{Boucot2013a, Cao2018a}.
+For instance, much of the Early Cretaceous low-latitude evaporite deposits were formed in basins that were located deep within arid continental interiors at the time of deposition \citep{Boucot2013a, Cao2018a}.
 Furthermore, in contrast to \citet{Evans2006a}, the compilation of evaporite deposits of \citet{Boucot2013a} contains sedimentary sequences in which the occurrence of evaporitic minerals is limited (e.g.
@@
 
-To evaluate the relationship between Earth's climate state and total and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic (\citealp{Macdonald2019a}; Fig.
+To evaluate the relationship between Earth's climate state and total and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets over the Phanerozoic (\citealp{Macdonald2019a}; Fig.
 \ref{fig:LIP_Areas}E).
 The goal in doing so is to evaluate the hypothesis that there is a correlation between LIP area in the tropics and Earth's long-term climate state.
-The land ice record is an imperfect tracker of climate, as it is insensitive to changes in temperature during non-glacial intervals, is influenced by additional factors such as the physical geography of the continents during glacial intervals, and is potentially vulnerable to removal from the observable geologic record via erosion and burial.
+The land ice record is an imperfect tracker of climate as it is insensitive to changes in temperature during non-glacial intervals, is influenced by additional factors such as the physical geography of the continents during glacial intervals, and is potentially vulnerable to removal from the observable geologic record via erosion and burial.
 Furthermore, the threshold \textit{p}CO$_2$ for establishing a glacial climate is dependent on ocean circulation and changing solar luminosity (e.g.
@@
 \ref{fig:LIP_correlation}).
-This approach accounts for the reality that spurious correlation can arise between auto-correlated data sets such as these, where each value is not independent, but is instead dependent on the previous state of the system.
+This approach accounts for the fact that spurious correlation can arise between auto-correlated data sets such as these, where each value is not independent, but is instead dependent on the previous state of the system.
 For the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario, 72\% of the randomly timed glacial interval simulations correlate better with LIP area in the tropics than the actual ice extent record.
@@
 
-In the original models that proposed the `Fire and Ice' hypothesis as an explanation for the onset of the Sturtian `Snowball Earth' glaciation, chemical weathering was modeled as a function of temperature and run-off only \citep{Donnadieu2004a}.
+In the original models that proposed the `Fire and Ice' hypothesis as an explanation for the onset of the Sturtian `Snowball Earth' glaciation, chemical weathering was modeled as a function of temperature and runoff only \citep{Donnadieu2004a}.
 However, such an approach neglects the effects of soil shielding and regolith development in low-relief regions.
@@
 
-The results from this analysis indicate that when the entire LIP database is considered through paleogeographic reconstruction and this parameterization of erosion, there is no significant relationship between total LIP area or LIP area in the tropics and the extent of continental ice sheets.
+The results from this analysis indicate that when the entire LIP database is considered in conjunction with a paleogeographic reconstruction and this parameterization of erosion, there is no significant relationship between total LIP area nor LIP area in the tropics and the extent of continental ice sheets.
 While this result need not imply that there is no increase in global weatherability from the emplacement of LIPs, it does suggest that changes in planetary weatherability associated with LIPs are not the fundamental control on whether Earth is in a glacial or non-glacial climate state.

### 2019-05-06 Manuscript/Manuscript.tex

@@
 We then compare these LIP area time series to the latitudinal extent of continental ice sheets - a proxy for Earth's long term climate state.
+This study builds upon the preliminary zonal LIP area analysis presented in the supplementary materials for \citet{Macdonald2019a} by more rigorously developing parameterizations of LIP erosion and exploring several geologically reasonable LIP post-emplacement scenarios (see below).
 
@@
 Other parameters that could be used to define the tropical rain belt are runoff and precipitation minus evaporation (P--E).
-When approaching the equator in modern climatological data, zonal mean runoff sharply increases to values above 0.6~kg/m$^{2}$ between approximately $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace (\citealp{Kalnay1996a}), and zonal mean P--E sharply increases to values above 1.5~kg/m$^{2}$/day also between approximately $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace (\citealp{Trenberth2011a}; Fig.
+When approaching the equator in modern climatological data, zonal mean runoff sharply increases to values above 0.25~m/yr between approximately $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace (\citealp{Fekete1999a}; Fig.
+\ref{fig:Climatology}), and zonal mean P--E sharply increases to values above 0.5~m/yr also between approximately $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace (\citealp{Trenberth2011a}; Fig.
 \ref{fig:Climatology}).
@@
 As a result, more recent modeling of chemical weathering incorporates such processes and highlights the importance of high-relief regions relative to low-relief ones for setting global weatherability \citep{West2012a, Godderis2017b}.
-LIPs are often emplaced in relatively low-relief areas, and as such, without active uplift, soil shielding from regolith development on these low-relief LIPs could significantly decrease the local weatherability of a LIP and mute its impact on global weatherability.
+LIPs are often emplaced in relatively low-relief areas, and as such, without active uplift, soil shielding from regolith development on these low-relief LIPs could significantly decrease the local weatherability of a LIP and mute its impact on global weatherability (as suggested in \citealp{Kent2013a}).
 In this way, soil shielding could explain the lack of correlation between tropical LIP area and ice extent (Figs.
@@
 How does the tropical LIP area associated with the Franklin LIP compare to that observed in the Phanerozoic?
-Using the paleomagnetic pole of \citet{Denyszyn2009a}, we reconstruct the paleolatitude of the Franklin LIP at the time of emplacement, and find that $\sim$99.7\% (or $\sim$2.6~Mkm$^{2}$) of the LIP erupted within 15\textdegree\xspace of the equator.
+Using the paleomagnetic pole of \citet{Denyszyn2009a}, we reconstruct the paleolatitude of the Franklin LIP at the time of emplacement, and find that $\sim$99.7\% (or $\sim$2.6~Mm$^{2}$) of the LIP erupted within 15\textdegree\xspace of the equator.
 This Franklin LIP tropical area at the time of emplacement is approximately equivalent to the Cenozoic peak, and is smaller than the other Phanerozoic peaks (Fig.
@@
 1109~Ma Umkondo is another Precambrian LIP that is constrained to have erupted in the tropics, although it is not known to be associated with any glaciation (no glacial deposits are found within the contemporaneous Midcontinent Rift basin; \citealp{Swanson-Hysell2019a}).
-We reconstruct the paleolatitude of the Umkondo LIP at the time of emplacement using the paleomagnetic pole of \citet{Swanson-Hysell2015b}, and find that effectively all of the LIP (or $\sim$2.0~Mkm$^{2}$) erupted within the tropics, an area that is slightly smaller than that estimated for the Franklin LIP (Fig.
+We reconstruct the paleolatitude of the Umkondo LIP at the time of emplacement using the paleomagnetic pole of \citet{Swanson-Hysell2015b}, and find that effectively all of the LIP (or $\sim$2.0~Mm$^{2}$) erupted within the tropics, an area that is slightly smaller than that estimated for the Franklin LIP (Fig.
 \ref{fig:LIP_Areas}C).
@@
 Given that similar (and larger) peaks in tropical LIP area are not associated with the onset of glacial periods, additional processes beyond an increase in weatherability due to LIP area in the tropics must have been at play in the initiation of the Sturtian Snowball Earth.
-One such process could have been an increase in global weatherability associated with the collision and accretion of arc terranes within the present-day Arabian-Nubian Shield \citep{Park2018a}.
-A low-latitude continent configuration and abundant arc-continent collisions in the tropics may have led to a cool background climate, and the emplacement of the Franklin LIP may have further increased global weatherability to the point where the ice-albedo runaway could take effect.
+One such process could have been unusually high planetary albedo associated with the low-latitude continental configuration of the supercontinent Rodinia \citep{Kirschvink1992a, Li2008a}.
+However, our analysis of zonal continental area reveals an almost invariant tropical continental area from $\sim$400~Ma to the present (Fig.
+\ref{fig:Continent_Areas}) and consequently no significant correlation between tropical continental area and the ice extent record, suggesting that a low-latitude continental configuration also could not have been the sole driver for the Sturtian Snowball Earth.
+Another potential contributing process is an increase in global weatherability associated with the collision and accretion of arc terranes within the present-day Arabian-Nubian Shield \citep{Park2018a}.
+Together, a low-latitude continental configuration and abundant arc-continent collisions in the tropics may have led to a cool background climate, and the emplacement of the Franklin LIP may have further increased global weatherability to the point where the ice-albedo runaway could take effect.
 However, tropical LIP area associated with the Franklin was not uniquely high, and therefore an associated increase in global weatherability was likely not the sole driver of Snowball Earth onset, consistent with the results of the Phanerozoic analysis.
@@
  & \textbf{(Ma)} & & \textbf{area} & \textbf{area} & & \textbf{original} & \textbf{(Myr)} & \\
- & & & \textbf{(Mkm$^{2}$)} & \textbf{(Mkm$^{2}$)} & & & & \\
+ & & & \textbf{(Mm$^{2}$)} & \textbf{(Mm$^{2}$)} & & & & \\
 \midrule
@@
 	\caption{Zonally averaged modern climatological data used to define the tropical rain belt.
-The global precipitation \citep{Kalnay1996a} and evaporation minus precipitation \citep{Trenberth2011a} data include land and ocean pixels, whereas global temperature \citep{Kalnay1996a} data are from land only.
+The global precipitation \citep{Kalnay1996a} and precipitation minus evaporation \citep{Trenberth2011a} data include land and ocean pixels, whereas global temperature data \citep{Kalnay1996a} are from land only, and runoff data \citep{Fekete1999a} are from land only excluding Antarctica.
+The peak in runoff $\sim$-50\textdegree\xspace is due to anomalously high runoff in the southern Andes, which represents almost all of the land in that latitude belt.
 Temperature data for Antarctica are off scale.
-Precipitation and precipitation minus evaporation both increase sharply between $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace.}
+Precipitation, precipitation minus evaporation, and runoff all increase sharply between $\pm$10\textdegree\xspace and $\pm$15\textdegree\xspace.}
 	\label{fig:Climatology}
@@
 
+\begin{figure}[h!]
+\begin{center}
+	\includegraphics[width=0.9\textwidth]{Manuscript/Figures/Continent_Areas.pdf}
+	\caption{\textbf{A)} Total continental area through time.
+In the paleogeographic model used in this study, tectonic units \citep{Torsvik2016a} are continually added to the model, leading to a net increase in total continental area in the model of $\sim$33\% over the Phanerozoic.
+However, estimates of continental crust growth (e.g.
+\citealp{Pujol2013a}) suggest that continental area was roughly constant through the Phanerozoic.
+We therefore normalize the total continental area curve in our model by assuming a fixed continental area through the Phanerozoic.
+\textbf{B)} Tropical continental area through time.
+We normalize the tropical continental area curve using the normalization ratio implied in (A).
+\textbf{C)} Contour plot showing the latitudinal distribution of continental area.
+\textbf{D)} Latitudinal extent of land ice away from the poles \citep{Macdonald2019a}.}
+	\label{fig:Continent_Areas}
+\end{center}
+\end{figure}
+
 \clearpage

### 2019-05-06 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 The global precipitation \citep{Kalnay1996a} and precipitation minus evaporation \citep{Trenberth2011a} data include land and ocean pixels, whereas global temperature data \citep{Kalnay1996a} are from land only, and runoff data \citep{Fekete1999a} are from land only excluding Antarctica.
-The peak in runoff $\sim$-50\textdegree\xspace is due to anomalously high runoff in the southern Andes, which represents almost all of the land in that latitude belt.
+The peak in runoff $\sim$-50\textdegree\xspace is due to anomalously high orographically-induced runoff in the southern Andes, which represents almost all of the land in that latitude belt.
 Temperature data for Antarctica are off scale.

### 2019-05-14 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 One of the hypothesized effects of large igneous provinces (LIPs) is planetary cooling on million-year timescales associated with enhanced silicate weathering of the freshly-emplaced basalt.
-This study combines reconstructions of the original surface extent and emplacement ages of LIPs, a paleogeographic model, and a paramaterization of LIP erosion to estimate LIP area in all latitudinal bands through the Phanerozoic.
+This study combines reconstructions of the original surface extent and emplacement ages of LIPs, a paleogeographic model, and a parameterization of LIP erosion to estimate LIP area in all latitudinal bands through the Phanerozoic.
 This analysis reveals no significant correlation between total LIP area, nor LIP area in the tropics, and the extent of continental ice sheets.
@@
 In the Cenozoic, the movement of the Deccan LIP into the tropical rain belt, together with the low-latitude emplacement of the Ethiopian Traps, has been implicated in drawing down CO$_2$ levels in the lead-up to Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
-More recently, \citet{Johansson2018a} used a paleogeographic model to suggest that tropical LIP area correlates with Phanerozoic climate change through comparison with a \textit{p}CO$_{2}$ proxy compilation.
+More recently, \citet{Johansson2018a} used paleogeographic reconstructions to suggest that tropical LIP area correlates with Phanerozoic climate change through comparison with a \textit{p}CO$_{2}$ proxy compilation.
 
@@
 We then compare these LIP area time series to the latitudinal extent of continental ice sheets - a proxy for Earth's long term climate state.
-This study builds upon the preliminary zonal LIP area analysis presented in the supplementary materials for \citet{Macdonald2019a} by more rigorously developing parameterizations of LIP erosion and exploring several geologically reasonable LIP post-emplacement scenarios (see below).
+This study builds upon a zonal LIP area analysis presented in the supplementary materials for \citet{Macdonald2019a} by more rigorously developing parameterizations of LIP erosion and exploring several geologically reasonable LIP post-emplacement scenarios (see below).
 
@@
 In the analysis of LIP area through time, we implement decay scenarios informed by this analysis: the `t$_{1/2}$ = 36~Myr' scenario uses the best-fit half-life of 36~Myr, while the `t$_{1/2}$ = 120~Myr' scenario uses the slower decay.
-Given that the post-emplacement weathering and erosional history of each LIP should be dependent on the tectonic and climatic setting that each LIP experiences during and after emplacement, this approach is simplistic but provides a framework for analysis.
+Given that the post-emplacement weathering and erosional history of each LIP should be dependent on the tectonic and climatic setting that each LIP experiences during and after emplacement, this approach is simplistic, but it provides a framework for analysis.
 The LIP reconstructions used in this study also include pre-Phanerozoic LIPs.
@@
 
-Of the tectonic factors that could alter exposed LIP area, the most consequential is near-immediate sediment burial of LIP volcanics that are co-located with a rift basin.
+Of the tectonic factors that could alter exposed LIP area, the most consequential is near-immediate burial by sediment of LIP volcanics that are co-located with a rift basin.
 There are numerous examples in the record where there is partial or complete burial of a LIP associated with rifting and thermal subsidence (Table \ref{tab:LIPs}).
@@
 \ref{fig:LIP_Areas}B and C).
-All the data and code necessary to reproduce the analyses and figures presented in this study can be downloaded from GitHub (\url{https://github.com/Swanson-Hysell-Group/Large_Igneous_Provinces}) and are also archived in Zenodo (URL TO BE ADDED HERE).
+All the data and code necessary to reproduce the analyses and figures presented in this study can be downloaded from GitHub (\url{https://github.com/Swanson-Hysell-Group/Large_Igneous_Provinces}) and are also archived in the Zenodo repository (URL TO BE ADDED HERE).
 
-We use 15\textdegree~S to 15\textdegree~N as the definition of the tropical rain belt, as those latitudes approximately correspond with a sharp increase in zonal mean precipitation when approaching the equator to values greater than 1.0~m/yr in modern climatalogical data (\citealp{Kalnay1996a}; Fig.
+The ascent of air near the equator associated with Earth's large-scale Hadley circulation promotes precipitation and leads to a low-latitude band of high rainfall known as the tropical rain belt.
+In contrast, the descending branches of the Hadley circulation in the subtropica are associated with aridity \citep{Manabe1969a}.
+We use 15\textdegree~S to 15\textdegree~N as a working definition of the tropical rain belt, as these latitudes approximately correspond with a sharp increase in zonal mean precipitation when approaching the equator to values greater than 1.0~m/yr in modern climatalogical data (\citealp{Kalnay1996a}; Fig.
 \ref{fig:Climatology}).
@@
 In \citet{Evans2006a}, the reconstructed paleolatitudes of basins with thick, basin-wide evaporite deposition are shown to be consistently in the subtropics throughout the Phanerozoic and the Proterozoic, suggesting that the large-scale atmospheric circulation that gives rise to intense precipitation in the tropical rain belt and an arid subtropical climate is stable through time.
-However, subsequent work by \citet{Boucot2013a} and \citet{Cao2018a} has shown that evaporite deposits did form at or near the equator at times in the Phanerozoic.
+However, subsequent work by \citet{Boucot2013a} and \citet{Cao2018a} has interpreted evaporite deposits to have formed at or near the equator at times in the Phanerozoic.
 While some of this variability could be attributed to waxing/waning of the width of the tropical rain belt as a whole, it is important to note that there can be large deviations in local precipitation from the zonal mean due to factors such as monsoon-related precipitation \citep{Trenberth2000a}, or continentality (i.e.
@@
 Furthermore, in contrast to \citet{Evans2006a}, the compilation of evaporite deposits of \citet{Boucot2013a} contains sedimentary sequences in which the occurrence of evaporitic minerals is limited (e.g.
-to nodular evaporitic pseudomorphs).
+to disseminated gypsum pseudomorphs).
 Such limited evaporitic mineral precipitation could be attributed to seasonal evaporation that transiently led to saturation states that otherwise would not be expected for that latitude.
@@
 However, our analysis of zonal continental area reveals an almost invariant tropical continental area from $\sim$400~Ma to the present (Fig.
-\ref{fig:Continent_Areas}) and consequently no significant correlation between tropical continental area and the ice extent record, suggesting that a low-latitude continental configuration also could not have been the sole driver for the Sturtian Snowball Earth.
-Another potential contributing process is an increase in global weatherability associated with the collision and accretion of arc terranes within the present-day Arabian-Nubian Shield \citep{Park2018a}.
+\ref{fig:Continent_Areas}) and consequently no significant correlation between tropical continental area and the ice extent record in the Phanerozoic, although it is intriguing that there are high and rising low-latitude continental area in the Ordovician.
+Similar to the LIP area analysis, this analysis suggests that a low-latitude continental configuration can not be invoked as the sole driver of planetary cooling, although it could be a contributing factor.
+Another potential contributing process for Neoproterozoic cooling leading up to the Sturtian glaciation is an increase in global weatherability associated with the collision and accretion of arc terranes within the present-day Arabian-Nubian Shield \citep{Park2018a}.
 Together, a low-latitude continental configuration and abundant arc-continent collisions in the tropics may have led to a cool background climate, and the emplacement of the Franklin LIP may have further increased global weatherability to the point where the ice-albedo runaway could take effect.
@@
 While this result need not imply that there is no increase in global weatherability from the emplacement of LIPs, it does suggest that changes in planetary weatherability associated with LIPs are not the fundamental control on whether Earth is in a glacial or non-glacial climate state.
-This result points to variable CO$_{2}$ outgassing or other changes in planetary weatherability such as mountain building in the tropics as more significant drivers.
 
@@
 	\caption{\textbf{A)} Total continental area through time.
-In the paleogeographic model used in this study, tectonic units \citep{Torsvik2016a} are continually added to the model, leading to a net increase in total continental area in the model of $\sim$33\% over the Phanerozoic.
+In the paleogeographic model used in this study, tectonic units \citep{Torsvik2016a} are progressively added to the model, leading to a net increase in total continental area in the model of $\sim$33\% over the Phanerozoic.
 However, estimates of continental crust growth (e.g.

### 2019-05-14 Manuscript/Manuscript.tex
v7 commit

@@
 \begin{center}
-	\includegraphics[width=\textwidth]{Manuscript/Figures/LIP_Map.pdf}
+	\includegraphics[width=\textwidth]{./Figures/LIP_Map.pdf}
 	\caption{Map of current extent of volcanic lithologies associated with LIPs that erupted between 520~Ma and the present, as well as the estimates of the initial LIP extent used in the area analysis (modified slightly from \citealp{Ernst2017a} and \citealp{Ernst2019a} to ensure that all currently exposed volcanic lithologies are encapsulated by the initial LIP area polygons).}
@@
 \begin{center}
-	\includegraphics[width=0.6\textwidth]{Manuscript/Figures/LIP_Preservation.pdf}
+	\includegraphics[width=0.6\textwidth]{./Figures/LIP_Preservation.pdf}
 	\caption{LIP erosion through time.
@@
 \begin{center}
-	\includegraphics[width=\textwidth]{Manuscript/Figures/Climatology.pdf}
+	\includegraphics[width=\textwidth]{./Figures/Climatology.pdf}
 	\caption{Zonally averaged modern climatological data used to define the tropical rain belt.
@@
 \begin{center}
-	\includegraphics[width=0.9\textwidth]{Manuscript/Figures/LIP_Areas.pdf}
+	\includegraphics[width=0.9\textwidth]{./Figures/LIP_Areas.pdf}
 	\caption{\textbf{A)} LIPs included in this analysis.
@@
 \begin{center}
-	\includegraphics[width=\textwidth]{Manuscript/Figures/Reconstruction_Snapshots.pdf}
+	\includegraphics[width=\textwidth]{./Figures/Reconstruction_Snapshots.pdf}
 	\caption{Paleogeographic reconstructions for times that correspond to peaks of LIP area in the tropics (Fig.
@@
 \begin{center}
-	\includegraphics[width=0.85\textwidth]{Manuscript/Figures/overlap_correlation_cropped.pdf}
+	\includegraphics[width=0.85\textwidth]{./Figures/overlap_correlation_cropped.pdf}
 	\caption{The \% overlap and Pearson correlation coefficients between LIP area and the actual ice-extent record are shown with circles.
@@
 \begin{center}
-	\includegraphics[width=0.9\textwidth]{Manuscript/Figures/Continent_Areas.pdf}
+	\includegraphics[width=0.9\textwidth]{./Figures/Continent_Areas.pdf}
 	\caption{\textbf{A)} Total continental area through time.

### 2019-05-15 Manuscript/Manuscript.tex
updates to text

@@
 
-\noindent\textit{This article is in review for the AGU book \textit{Environmental Change and Large Igneous Provinces: The Deadly Kiss of LIPs}.}
+\noindent\textit{This article is in review for the AGU book entitled \textit{Environmental Change and Large Igneous Provinces: The Deadly Kiss of LIPs}.}
 
@@
 This study combines reconstructions of the original surface extent and emplacement ages of LIPs, a paleogeographic model, and a parameterization of LIP erosion to estimate LIP area in all latitudinal bands through the Phanerozoic.
-We then compare these LIP area time series to the latitudinal extent of continental ice sheets - a proxy for Earth's long term climate state.
-This study builds upon a zonal LIP area analysis presented in the supplementary materials for \citet{Macdonald2019a} by more rigorously developing parameterizations of LIP erosion and exploring several geologically reasonable LIP post-emplacement scenarios (see below).
+We then compare these time series of zonal LIP area to the latitudinal extent of continental ice sheets - a proxy for Earth's long term climate state.
+This study builds upon a zonal LIP area analysis presented in the supplementary materials for \citet{Macdonald2019a} by more rigorously developing parameterizations of LIP erosion and exploring several geologically reasonable LIP post-emplacement scenarios.
 
@@
 These original extent outlines could also under-estimate the surface area for some LIPs where flows have been eroded and feeder dikes are not exposed or poorly documented.
-However, despite these uncertainties, this approach likely provides the best estimates available of original extent for ancient LIPs.
+However, despite these uncertainties, this approach likely provides the best estimates available of original surface extent for ancient LIPs.
 The extents of presently-exposed volcanics associated with LIPs that were used for present-day area estimates (Fig.
-\ref{fig:LIP_map}), and the sources that went into the construction of the original extent polygons by \citet{Ernst2017a} and \citet{Ernst2019a}, were taken from a number of resources including the PLATES compilation \citep{Coffin2006a} and more recent compilation efforts associated with the LIPs Reconstruction Project (\citealp{Ernst2013a}; Table \ref{tab:LIPs}).
+\ref{fig:LIP_map}) and the sources that went into the construction of the original extent polygons by \citet{Ernst2017a} and \citet{Ernst2019a} were taken from a number of resources including the PLATES compilation \citep{Coffin2006a} and more recent compilation efforts associated with the LIPs Reconstruction Project (\citealp{Ernst2013a}; Table \ref{tab:LIPs}).
 
 After LIPs are emplaced, they progressively erode.
-In order to account for the associated decrease in area with time, \citet{Godderis2017a} took the approach of fitting an exponential decay function to estimates of the original surface extent and the current surface extent for 5 LIPs.
-They used the resulting exponential decay constants to develop a first-order parameterization to estimate changing LIP area through time.
+In order to account for the associated decrease in area with time, \citet{Godderis2017a} took the approach of fitting an exponential decay function to estimates of the original surface extent and the current surface extent of 5 LIPs.
+They used the resulting exponential decay constants to develop a first-order parameterization of changing LIP area through time.
 We extend this approach to 19 basaltic LIPs for which there are estimates of the original surface extent of the province and the current surface extent of rocks associated with the province (Fig.
@@
 it yields a low root mean square error of 0.14), if each LIP is fit individually with an exponential decay function, there is variability in the estimated half-lives from $\sim$20~Myr up to $\sim$120~Myr for the Deccan Traps (Table \ref{tab:LIPs}).
-In the analysis of LIP area through time, we implement decay scenarios informed by this analysis: the `t$_{1/2}$ = 36~Myr' scenario uses the best-fit half-life of 36~Myr, while the `t$_{1/2}$ = 120~Myr' scenario uses the slower decay.
+In the analysis of LIP area through time, we implement decay scenarios informed by these results: the `t$_{1/2}$ = 36~Myr' scenario uses the best-fit half-life of 36~Myr, while the `t$_{1/2}$ = 120~Myr' scenario uses the slower decay.
 Given that the post-emplacement weathering and erosional history of each LIP should be dependent on the tectonic and climatic setting that each LIP experiences during and after emplacement, this approach is simplistic, but it provides a framework for analysis.
@@
 The `50\% burial' scenario imposes instantaneous burial of 50$\%$ of the LIP area while the `100\% burial' imposes instantaneous burial of the entire LIP as an end-member scenario.
-A limitation of our treatment of LIPs that are co-located with rifting is that we `bury' all of these LIPs instantly at the time of emplacement and to the same degree when in fact the degree/timing of burial of these LIPs may vary substantially.
+A limitation of our treatment of LIPs that are co-located with rifting is that we `bury' all of these LIPs instantly at the time of emplacement and to the same degree when in fact the degree and timing of burial of these LIPs may vary substantially.
 The LIP area analysis uses all of the available combinations of the distinct decay and burial scenarios described above.
 
-Our LIP reconstruction differs from that of \citet{Johansson2018a} in several respects.
+Our LIP reconstruction differs from that of \citet{Johansson2018a}.
 In contrast to the decay and decay+burial scenarios implemented on estimates of original LIP extent in this study, \citet{Johansson2018a} uses a static extent for each LIP throughout the reconstruction.
@@
 
-The original extent LIP polygons were assigned a plate ID corresponding to a tectonic unit on Earth using the polygons of \citet{Torsvik2016a} for the Phanerozoic.
+The original surface extent LIP polygons were assigned a plate ID corresponding to a tectonic unit on Earth using the polygons of \citet{Torsvik2016a} for the Phanerozoic.
 The LIP polygons and tectonic units were reconstructed from 520~Ma to the present (e.g.
@@
 The ascent of air near the equator associated with Earth's large-scale Hadley circulation promotes precipitation and leads to a low-latitude band of high rainfall known as the tropical rain belt.
-In contrast, the descending branches of the Hadley circulation in the subtropica are associated with aridity \citep{Manabe1969a}.
+In contrast, the descending branches of the Hadley circulation in the subtropics are associated with aridity \citep{Manabe1969a}.
 We use 15\textdegree~S to 15\textdegree~N as a working definition of the tropical rain belt, as these latitudes approximately correspond with a sharp increase in zonal mean precipitation when approaching the equator to values greater than 1.0~m/yr in modern climatalogical data (\citealp{Kalnay1996a}; Fig.
@@
 \citealp{West2012a}).
-Using runoff or P--E favors a definition of the tropical rain belt that is $\pm$10\textdegree\xspace rather than $\pm$15\textdegree\xspace.
+Using runoff or P--E favors a definition of the tropical rain belt that is closer to $\pm$10\textdegree\xspace rather than $\pm$15\textdegree\xspace.
 Therefore, we tested the sensitivity of our results to the assumed width of the tropical rain belt by performing the tropical LIP area calculations with a tropical rain belt width of $\pm$10\textdegree\xspace.
@@
 Nevertheless, it forms a physical record of Earth's climate through time and delineates glacial and non-glacial climate states.
-We take two approaches for comparison between the LIP area reconstruction and the record of ice extent.
+We take two approaches for comparison between the LIP area reconstructions and the record of ice extent.
 The first is to calculate the Pearson correlation coefficient between LIP area and the extent of ice away from the pole.
-The second is to consider the overlap between intervals of high LIP area (defined as LIP area \textgreater30\% of the maximum in a given post-emplacement model) and intervals of glacial climate (defined as ice extent \textgreater10\textdegree\xspace from the poles).
+The second is to consider the degree of overlap between intervals of high LIP area (defined as LIP area \textgreater30\% of the maximum in a given post-emplacement model) and intervals of glacial climate (defined as ice extent \textgreater10\textdegree\xspace from the poles).
 This overlap approach places less emphasis on the specific magnitudes of the peaks in the compiled ice extent and LIP area records.
 
-Another approach would be to compare LIP area records to proxy compilations of \textit{p}CO$_2$ (as done by \citealp{Johansson2018a}).
+Another approach would be to compare the LIP area reconstructions to proxy compilations of \textit{p}CO$_2$ (as done by \citealp{Johansson2018a}) instead of the latitudinal extent of continental ice sheets.
 However, such \textit{p}CO$_2$ proxies are potentially problematic as they can be difficult to calibrate in deep time and can be affected by secondary alteration.
@@
 
-A complication with the interpretation of soil shielding and limited weathering of LIPs is the rapid area decay rate (t$_{1/2}$ = 36~Myr) inferred from the comparison of current LIP extent to estimated original extent (Fig.
+A complication with the interpretation of soil shielding and limited weathering of LIPs is the rapid area decay rate (t$_{1/2}$ = 36~Myr) inferred from the comparison of current LIP surface extent to estimated original surface extent (Fig.
 \ref{fig:LIP_preservation}).
-A couple considerations are relevant with respect to this analysis: 1) the current extent of LIP exposure is reduced in part by volcanics being covered by unconsolidated sediments (i.e.
-regolith development itself) in a number of the provinces; 2) the current extent of LIP exposure may be incomplete and an underestimate for some of the provinces; 3) the initial LIP extents are typically poorly constrained and are likely over-estimates which could be resulting in inflated interpreted decay rates; and 4) the relationship between LIP area and volume is poorly constrained.
-Future efforts that improve the LIP database, such as developing better-constrained estimates of original LIP extent, constraining burial and uplift histories, and refining the timing of eruptions associated with LIPs, will improve analyses that consider the LIP record in its entirety, such as that in this contribution.
+A couple considerations are relevant with respect to this analysis: 1) the current surface extent of LIP exposure is reduced in part by volcanics being covered by unconsolidated sediments (i.e.
+regolith development itself) in a number of the provinces; 2) the current surface extent of LIP exposure may be incomplete and an underestimate for some of the provinces; 3) the initial LIP surface extents are typically poorly constrained and are likely over-estimates which could be resulting in inflated interpreted decay rates; and 4) the relationship between LIP area and volume is poorly constrained.
+Future efforts that improve the LIP database, such as developing better-constrained estimates of original LIP surface extent, constraining burial and uplift histories, and refining the timing of eruptions associated with LIPs, will improve analyses that consider the LIP record in its entirety, such as that in this contribution.
 
@@
 However, our analysis of zonal continental area reveals an almost invariant tropical continental area from $\sim$400~Ma to the present (Fig.
-\ref{fig:Continent_Areas}) and consequently no significant correlation between tropical continental area and the ice extent record in the Phanerozoic, although it is intriguing that there are high and rising low-latitude continental area in the Ordovician.
-Similar to the LIP area analysis, this analysis suggests that a low-latitude continental configuration can not be invoked as the sole driver of planetary cooling, although it could be a contributing factor.
+\ref{fig:Continent_Areas}) and consequently no significant correlation between tropical continental area and the ice extent record in the Phanerozoic, although it is intriguing that there is a high and rising low-latitude continental area in the Ordovician.
+Similar to the LIP area analysis, this continental area analysis suggests that a low-latitude continental configuration can not be invoked as the sole driver of planetary cooling, although it could be a contributing factor.
 Another potential contributing process for Neoproterozoic cooling leading up to the Sturtian glaciation is an increase in global weatherability associated with the collision and accretion of arc terranes within the present-day Arabian-Nubian Shield \citep{Park2018a}.
@@
 \toprule
-\textbf{} & \multicolumn{4}{c|}{\textbf{within tropics $\pm$15$^{\circ}$}$^{1}$} & \multicolumn{4}{c|}{\textbf{total}$^{6}$} & \multicolumn{4}{c|}{\textbf{within tropics $\pm$10$^{\circ}$}} & \multicolumn{4}{c}{\textbf{within tropics $\pm$20$^{\circ}$}} \\
+\textbf{} & \multicolumn{4}{c|}{\textbf{within tropics $\pm$15$^{\circ}$}} & \multicolumn{4}{c|}{\textbf{total}$^{5}$} & \multicolumn{4}{c|}{\textbf{within tropics $\pm$10$^{\circ}$}} & \multicolumn{4}{c}{\textbf{within tropics $\pm$20$^{\circ}$}} \\
 &&&&&&&&&&&&&&&& \\
-\textbf{} & \multicolumn{2}{c}{\textbf{correlation}$^{2}$} & \multicolumn{2}{c|}{\textbf{\% overlap}$^{3}$} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c|}{\textbf{\% overlap}} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c|}{\textbf{\% overlap}} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c}{\textbf{\% overlap}} \\
-\textbf{scenario} & \textbf{val.}$^{4}$ & \textbf{p-val.}$^{5}$ & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} \\
+\textbf{} & \multicolumn{2}{c}{\textbf{correlation}$^{1}$} & \multicolumn{2}{c|}{\textbf{\% overlap}$^{2}$} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c|}{\textbf{\% overlap}} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c|}{\textbf{\% overlap}} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c}{\textbf{\% overlap}} \\
+\textbf{scenario} & \textbf{val.}$^{3}$ & \textbf{p-val.}$^{4}$ & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} \\
 \midrule
@@
 \vspace{0.15cm}
-$^{1}$our preferred definition of the tropical rain belt.
+$^{1}$Pearson correlation coefficient between LIP area and the actual ice extent record.
 \vspace{0.15cm}
 
-$^{2}$Pearson correlation coefficient between LIP area and the actual ice extent record.
+$^{2}$\% of time when both LIP area is \textgreater30\% of the maximum and ice extent is \textgreater10\textdegree\xspace from the poles.
 \vspace{0.15cm}
 
-$^{3}$\% of time when both LIP area is \textgreater30\% of the maximum and ice extent is \textgreater10\textdegree\xspace from the poles.
+$^{3}$`val.' refers to the computed correlation coefficient/\% overlap between LIP area and the actual ice extent record.
 \vspace{0.15cm}
 
-$^{4}$`val.' refers to the computed correlation coefficient/\% overlap between LIP area and the actual ice extent record.
-\vspace{0.15cm}
-
-$^{5}$`p-val.' refers to the fraction of randomly timed glacial interval simulations that correlate/overlap better with LIP area than the actual ice extent record (i.e.
+$^{4}$`p-val.' refers to the fraction of randomly timed glacial interval simulations that correlate/overlap better with LIP area than the actual ice extent record (i.e.
 the p-value with respect to the null hypothesis of no correlation/overlap).
@@
 
-$^{6}$all latitudes.
+$^{5}$all latitudes.
 \end{tablenotes}
@@
 	\includegraphics[width=\textwidth]{./Figures/LIP_Map.pdf}
-	\caption{Map of current extent of volcanic lithologies associated with LIPs that erupted between 520~Ma and the present, as well as the estimates of the initial LIP extent used in the area analysis (modified slightly from \citealp{Ernst2017a} and \citealp{Ernst2019a} to ensure that all currently exposed volcanic lithologies are encapsulated by the initial LIP area polygons).}
+	\caption{Map of current surface extent of volcanic lithologies associated with LIPs that erupted between 520~Ma and the present, as well as the estimates of the initial LIP surface extent used in the area analysis (modified slightly from \citealp{Ernst2017a} and \citealp{Ernst2019a} to ensure that all currently exposed volcanic lithologies are encapsulated by the initial LIP surface area polygons).}
 	\label{fig:LIP_map}
@@
 	\caption{LIP erosion through time.
-The ratio of estimates of the present-day area to that of the original area are shown for 19 basaltic LIPs.
+The ratio of estimates of the present-day surface area to that of the original surface area are shown for 19 basaltic LIPs.
 An exponential fit is made to the 13 basaltic LIPs that are interpreted to not have been buried after emplacement (Table \ref{tab:LIPs}), which yields a half-life of $\sim$36~Myr.
@@
 	\caption{Zonally averaged modern climatological data used to define the tropical rain belt.
-The global precipitation \citep{Kalnay1996a} and precipitation minus evaporation \citep{Trenberth2011a} data include land and ocean pixels, whereas global temperature data \citep{Kalnay1996a} are from land only, and runoff data \citep{Fekete1999a} are from land only excluding Antarctica.
+The global precipitation \citep{Kalnay1996a} and precipitation minus evaporation \citep{Trenberth2011a} data include land and ocean pixels, global temperature data \citep{Kalnay1996a} are from land only, and runoff data \citep{Fekete1999a} are from land only excluding Antarctica.
 The peak in runoff $\sim$-50\textdegree\xspace is due to anomalously high orographically-induced runoff in the southern Andes, which represents almost all of the land in that latitude belt.
@@
 	\caption{\textbf{A)} LIPs included in this analysis.
-The size of each circle reflects the initial area estimate of each LIP.
+The size of each circle reflects the initial surface area estimate of each LIP.
 The + indicates the timing and absolute paleolatitude of the centroid of each LIP at the time of emplacement.
@@
 \ref{fig:Reconstruction_Snapshots}.
-\textbf{D)} Contour plot showing the latitudinal distribution of LIP area for the preferred post-emplacement model.
+\textbf{D)} Contour plot showing the latitudinal distribution of LIP area for one of the post-emplacement models.
 \textbf{E)} Latitudinal extent of land ice away from the poles \citep{Macdonald2019a} and compilation of \textit{p}CO$_{2}$ proxies \citep{Foster2017a} (\textit{p}CO$_{2}$ y-axis reversed, and in log-scale).

### 2019-05-16 Manuscript/Manuscript.tex

@@
 \ref{fig:LIP_Areas}B and C).
-All the data and code necessary to reproduce the analyses and figures presented in this study can be downloaded from GitHub (\url{https://github.com/Swanson-Hysell-Group/Large_Igneous_Provinces}) and are also archived in the Zenodo repository (URL TO BE ADDED HERE).
+All the data and code necessary to reproduce the analyses and figures presented in this study can be downloaded from GitHub (\url{https://github.com/Swanson-Hysell-Group/Large_Igneous_Provinces}).
 

### 2019-06-06 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 \ref{fig:LIP_Areas}B and C).
-All the data and code necessary to reproduce the analyses and figures presented in this study can be downloaded from GitHub (\url{https://github.com/Swanson-Hysell-Group/Large_Igneous_Provinces}).
+All the data and code necessary to reproduce the analyses and figures presented in this study can be downloaded from GitHub (\url{https://github.com/Swanson-Hysell-Group/2019_large_igneous_provinces}).
 

### 2019-11-28 Manuscript.tex
Updates from Overleaf

--- 2019-06-06 Manuscript/Manuscript.tex
@@
 \doublespacing
+
+% make index - for this to work, the .tex file needs to be in the root directory
+\usepackage{imakeidx}
+\makeindex[columns=2, title=Index]
 
@@
 \fancyhf{}
-\lhead{Park et al., in review for AGU book entitled \textit{Environmental Change and Large Igneous Provinces: The Deadly Kiss of LIPs}}
+\lhead{Park et al., accepted for publication in the AGU book entitled \textit{Environmental Change and large igneous provinces: The Deadly Kiss of LIPs}}
 \rhead{\thepage}
@@
 
-\noindent\textit{This article is in review for the AGU book entitled \textit{Environmental Change and Large Igneous Provinces: The Deadly Kiss of LIPs}.}
+\noindent\textit{This article is accepted for publication in the AGU book entitled \textit{Environmental Change and large igneous provinces: The Deadly Kiss of LIPs}.}
 
@@
 
-One of the hypothesized effects of large igneous provinces (LIPs) is planetary cooling on million-year timescales associated with enhanced silicate weathering of the freshly-emplaced basalt.
+One of the hypothesized effects of large igneous provinces (LIPs) is planetary cooling\index{cooling} on million-year timescales associated with enhanced silicate weathering\index{silicate weathering} of the freshly-emplaced basalt\index{basalt}.
 This study combines reconstructions of the original surface extent and emplacement ages of LIPs, a paleogeographic model, and a parameterization of LIP erosion to estimate LIP area in all latitudinal bands through the Phanerozoic.
@@
 Global weatherability is the sum of factors aside from climate itself that contribute to overall global weathering and associated CO$_2$ consumption, such as the latitudinal distribution of continents and mountain belts \citep{Kump1997a}.
-On a planet with high weatherability, the CO$_2$ input from volcanism can be removed via silicate weathering at a lower atmospheric CO$_2$ concentration than on a less weatherable planet.
-Basaltic regions consume more CO$_2$ than regions where the bedrock composition is closer to bulk continental crust because mafic lithologies have relatively high concentrations of Ca and Mg (that ultimately sequester carbon through precipitation as carbonate), constitute minerals with relatively high reactivity \citep{Gislason2003a}, and have relatively high weathering rates \citep{Dessert2003a, Ibarra2016a}.
-Furthermore, data from basaltic watersheds show that chemical weathering rates are highest in regions with high runoff and temperature.
-As a result, CO$_2$ consumption in basaltic regions is most pronounced in the tropical rain belt \citep{Dessert2003a, Hartmann2009a, Hartmann2014a}.
+On a planet with high weatherability, the CO$_2$ input from volcanism can be removed via silicate weathering\index{silicate weathering} at a lower atmospheric CO$_2$ concentration than on a less weatherable planet.
+Basaltic\index{basalt} regions consume more CO$_2$ than regions where the bedrock composition is closer to bulk continental crust because mafic lithologies have relatively high concentrations of Ca and Mg (that ultimately sequester carbon through precipitation as carbonate), constitute minerals with relatively high reactivity \citep{Gislason2003a}, and have relatively high weathering rates \citep{Dessert2003a, Ibarra2016a}.
+Furthermore, data from basaltic\index{basalt} watersheds show that chemical weathering rates are highest in regions with high runoff and temperature.
+As a result, CO$_2$ consumption in basaltic\index{basalt} regions is most pronounced in the tropical rain belt \citep{Dessert2003a, Hartmann2009a, Hartmann2014a}.
 
-One aspect of large igneous province (LIP) emplacement that has been hypothesized to relate to long-term global climate is the effect that associated mafic lithologies could have on increasing global weatherability and driving cooling.
-In particular, the emplacement of LIPs in the tropics has been hypothesized to be associated with specific episodes of climatic cooling on Earth.
+One aspect of large igneous province (LIP) emplacement that has been hypothesized to relate to long-term global climate is the effect that associated mafic lithologies could have on increasing global weatherability and driving cooling\index{cooling}.
+In particular, the emplacement of LIPs in the tropics has been hypothesized to be associated with specific episodes of climatic cooling\index{cooling} on Earth.
 In the Neoproterozoic, the emplacement of the ca.
-720~Ma Franklin LIP in the tropics, in concert with elevated runoff rates associated with supercontinent break-up, has been implicated as a major contributor to the cooling that initiated the Sturtian `Snowball Earth' \citep{Donnadieu2004b, Macdonald2010a, Cox2016a}.
+720~Ma Franklin LIP in the tropics, in concert with elevated runoff rates associated with supercontinent break-up, has been implicated as a major contributor to the cooling\index{cooling} that initiated the Sturtian `Snowball Earth' \citep{Donnadieu2004b, Macdonald2010a, Cox2016a}.
 In the Cenozoic, the movement of the Deccan LIP into the tropical rain belt, together with the low-latitude emplacement of the Ethiopian Traps, has been implicated in drawing down CO$_2$ levels in the lead-up to Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
@@
 
-This chapter seeks to address two interconnected questions: 1) how unique are the peaks in low-latitude LIP area that have been proposed to be associated with climatic cooling?; and 2) how strong is the overall relationship between tropical LIP area and glaciation?
+This chapter seeks to address two interconnected questions: 1) how unique are the peaks in low-latitude LIP area that have been proposed to be associated with climatic cooling\index{cooling}?; and 2) how strong is the overall relationship between tropical LIP area and glaciation?
 
@@
 They used the resulting exponential decay constants to develop a first-order parameterization of changing LIP area through time.
-We extend this approach to 19 basaltic LIPs for which there are estimates of the original surface extent of the province and the current surface extent of rocks associated with the province (Fig.
+We extend this approach to 19 basaltic\index{basalt} LIPs for which there are estimates of the original surface extent of the province and the current surface extent of rocks associated with the province (Fig.
 \ref{fig:LIP_preservation}).
@@
 There are numerous examples in the record where there is partial or complete burial of a LIP associated with rifting and thermal subsidence (Table \ref{tab:LIPs}).
-For example, the Afar LIP is both associated with the Ethiopian Traps which form plateau flood basalts as well as successful rifting in the region of the Red Sea that has resulted in burial (Fig.
+For example, the Afar LIP is both associated with the Ethiopian Traps which form plateau flood basalts\index{basalt} as well as successful rifting in the region of the Red Sea that has resulted in burial (Fig.
 \ref{fig:LIP_map}).
@@
 We have focused this analysis on the Phanerozoic record given that well-constrained paleogeographic models are available for the past $\sim$520~Myr.
-The approach of seeking to evaluate correlation between LIP area and glaciation is further complicated for Neoproterozoic Snowball Earth events because ice-albedo runaway leads to persistent global glaciation on timescales of tens of millions of years without continued forcing through normal carbon cycle processes until sufficient CO$_2$ to drive deglaciation builds up in the absence of silicate weathering \citep{Hoffman2017a}.
-Moreover, cooling past the critical threshold for rapid global glaciation may have occurred on a sub-million year timescale (e.g.
+The approach of seeking to evaluate correlation between LIP area and glaciation is further complicated for Neoproterozoic Snowball Earth events because ice-albedo runaway leads to persistent global glaciation on timescales of tens of millions of years without continued forcing through normal carbon cycle processes until sufficient CO$_2$ to drive deglaciation builds up in the absence of silicate weathering\index{silicate weathering} \citep{Hoffman2017a}.
+Moreover, cooling\index{cooling} past the critical threshold for rapid global glaciation may have occurred on a sub-million year timescale (e.g.
 \citealp{Macdonald2017a}).
@@
 \ref{fig:Continent_Areas}) and consequently no significant correlation between tropical continental area and the ice extent record in the Phanerozoic, although it is intriguing that there is a high and rising low-latitude continental area in the Ordovician.
-Similar to the LIP area analysis, this continental area analysis suggests that a low-latitude continental configuration can not be invoked as the sole driver of planetary cooling, although it could be a contributing factor.
-Another potential contributing process for Neoproterozoic cooling leading up to the Sturtian glaciation is an increase in global weatherability associated with the collision and accretion of arc terranes within the present-day Arabian-Nubian Shield \citep{Park2018a}.
+Similar to the LIP area analysis, this continental area analysis suggests that a low-latitude continental configuration can not be invoked as the sole driver of planetary cooling\index{cooling}, although it could be a contributing factor.
+Another potential contributing process for Neoproterozoic cooling\index{cooling} leading up to the Sturtian glaciation is an increase in global weatherability associated with the collision and accretion of arc terranes within the present-day Arabian-Nubian Shield \citep{Park2018a}.
 Together, a low-latitude continental configuration and abundant arc-continent collisions in the tropics may have led to a cool background climate, and the emplacement of the Franklin LIP may have further increased global weatherability to the point where the ice-albedo runaway could take effect.
@@
 The temporary effect on albedo of such aerosols is maximized when they are injected into the atmosphere at low-latitudes into a cool background climate and their presence at high concentrations is pre-conditioned on eruption through sedimentary basins hosting evaporite deposits, as could have been the case for the Franklin LIP \citep{Macdonald2017a}.
-However, in the cases in which such aerosol-driven cooling does not result in ice-albedo runaway and a Snowball Earth, the climate would return to its background climate state within years \citep{Macdonald2017a}.
+However, in the cases in which such aerosol-driven cooling\index{cooling} does not result in ice-albedo runaway and a Snowball Earth, the climate would return to its background climate state within years \citep{Macdonald2017a}.
 A contrasting effect is that, on 1~kyr to 1~Myr timescales, LIP emplacement could instead cause transient warming associated with elevated CO$_{2}$ outgassing leading to transiently high \textit{p}CO$_{2}$, as has been argued for the CAMP \citep{Schaller2011a, Schaller2012a}.
@@
 \begin{center}
-	\includegraphics[width=\textwidth]{./Figures/LIP_Map.pdf}
+	\includegraphics[width=\textwidth]{./Manuscript/Figures/LIP_Map.pdf}
 	\caption{Map of current surface extent of volcanic lithologies associated with LIPs that erupted between 520~Ma and the present, as well as the estimates of the initial LIP surface extent used in the area analysis (modified slightly from \citealp{Ernst2017a} and \citealp{Ernst2019a} to ensure that all currently exposed volcanic lithologies are encapsulated by the initial LIP surface area polygons).}
@@
 \begin{center}
-	\includegraphics[width=0.6\textwidth]{./Figures/LIP_Preservation.pdf}
+	\includegraphics[width=0.6\textwidth]{./Manuscript/Figures/LIP_Preservation.pdf}
 	\caption{LIP erosion through time.
@@
 \begin{center}
-	\includegraphics[width=\textwidth]{./Figures/Climatology.pdf}
+	\includegraphics[width=\textwidth]{./Manuscript/Figures/Climatology.pdf}
 	\caption{Zonally averaged modern climatological data used to define the tropical rain belt.
@@
 \begin{center}
-	\includegraphics[width=0.9\textwidth]{./Figures/LIP_Areas.pdf}
+	\includegraphics[width=0.9\textwidth]{./Manuscript/Figures/LIP_Areas.pdf}
 	\caption{\textbf{A)} LIPs included in this analysis.
@@
 \begin{center}
-	\includegraphics[width=\textwidth]{./Figures/Reconstruction_Snapshots.pdf}
+	\includegraphics[width=\textwidth]{./Manuscript/Figures/Reconstruction_Snapshots.pdf}
 	\caption{Paleogeographic reconstructions for times that correspond to peaks of LIP area in the tropics (Fig.
@@
 \begin{center}
-	\includegraphics[width=0.85\textwidth]{./Figures/overlap_correlation_cropped.pdf}
+	\includegraphics[width=0.85\textwidth]{./Manuscript/Figures/overlap_correlation_cropped.pdf}
 	\caption{The \% overlap and Pearson correlation coefficients between LIP area and the actual ice-extent record are shown with circles.
@@
 \begin{center}
-	\includegraphics[width=0.9\textwidth]{./Figures/Continent_Areas.pdf}
+	\includegraphics[width=0.9\textwidth]{./Manuscript/Figures/Continent_Areas.pdf}
 	\caption{\textbf{A)} Total continental area through time.
@@
 
-\bibliographystyle{gsabull}
-\bibliography{References}
+\bibliographystyle{./Manuscript/gsabull.bst}
+\bibliography{./Manuscript/References.bib}
+
+\nolinenumbers
+\printindex
 
